### 2020-11-04 inst/jma_knot.tex
manuscript for Journal of Mathematics and the Arts

% ------------------------------------------------------------------------
% bjourdoc.tex for birkjour.cls*******************************************
% ------------------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{birkjour}
%%%Optional but convenient to use is the package ``cite''. If you do not want to use, remark the next line by placing the percent sign % in front of it:
\usepackage[noadjust]{cite}
\usepackage{amsmath}
%
%
% THEOREM Environments (Examples)-----------------------------------------
%
 \newtheorem{thm}{Theorem}[section]
 \newtheorem{cor}[thm]{Corollary}
 \newtheorem{lem}[thm]{Lemma}
 \newtheorem{prop}[thm]{Proposition}
 \theoremstyle{definition}
 \newtheorem{defn}[thm]{Definition}
 \theoremstyle{remark}
 \newtheorem{rem}[thm]{Remark}
\newtheorem{comment}[thm]{Comment}
 \newtheorem*{ex}{Example}
 \numberwithin{equation}{section}

\newcommand{\ed}{\end{document}}
\begin{document}

%-------------------------------------------------------------------------
% editorial commands: to be inserted by the editorial office
%
%\firstpage{1} \volume{228} \Copyrightyear{2004} \DOI{003-0001}
%
%
%\seriesextra{Just an add-on}
%\seriesextraline{This is the Concrete Title of this Book\br H.E. R and S.T.C. W, Eds.}
%
% for journals:
%
%\firstpage{1}
%\issuenumber{1}
%\Volumeandyear{1 (2004)}
%\Copyrightyear{2004}
%\DOI{003-xxxx-y}
%\Signet
%\commby{inhouse}
%\submitted{March 14, 2003}
%\received{March 16, 2000}
%\revised{June 1, 2000}
%\accepted{July 22, 2000}
%
%
%
%---------------------------------------------------------------------------
%Insert here the title, affiliations and abstract:
%

\title[Visually pleasing knot projections]{Visually pleasing knot projections}

%----------Author 1
\author[Hankin]{Robin K. S. Hankin}

\address{%
55 Wellesley Street East, Auckland 1010, New Zealand}

\email{hankin.robin@gmail.com}

%----------classification, keywords, date
\subjclass{Primary 99Z99; Secondary 00A00}

\keywords{Knot projections, Bezier curves, Multidimensional optimization}

\date{\today}
%----------additions
%%% ----------------------------------------------------------------------

\begin{abstract}
In this short article I introduce software which creates two
dimensional knot diagrams optimized for visual appearance.  Different
aspects of a knot's appearance are discussed and a framework for
objectively optimizing the visual appeal of a knot projection is
given.

In this short article I show how to create two dimensional knot
diagrams optimized for visual appearance using the R programming
language.  The package associated with this paper is a suite
of software for the creation of production-quality artwork of knot
diagrams, released under GPL2.
\end{abstract}

%%% ----------------------------------------------------------------------
\maketitle
%%% ----------------------------------------------------------------------
%\tableofcontents

\section{Introduction}

A {\em mathematical knot} is a smooth, unoriented embedding of a
circle~$\mathbb{S}^1$ into~$\mathbb{R}^3$~\cite{manturov2004}.  Two
knots are said to be equivalent if one can be continuously deformed in
to the other; if so, there is a
homeomorphism~$h\colon\mathbb{R}^3\longrightarrow\mathbb{R}^3$ which
takes one embedded circle to the other.

It is common to present a knot using diagrams such as
Figure~\ref{knot_table}, in which a two-dimensional projection of the
knot is given with broken lines indicating where one strand passes
over another.  Related work in the same language is given by Comoglio
and Rinaldi~\cite{comoglio2011,comoglio2012}; the visual depiction of
knot projections in the context of Celtic artwork is given by
Browne~\cite{browne2005,browne2006}.

\begin{figure}[h]
  \centering
  \includegraphics[width=10cm]{"1280px-Knot_table".png}
  \caption{A table of prime knots up \label{knot_table} to seven
    crossings, labels following~\cite{alexander1926}.  Image taken
    from~\cite{wikipedia_knot_theory}.}
\end{figure}

Consider Figure~\ref{knot_table} from an \ae sthetic perspective; the
diagrams are representative of those available under a free license.
However, these diagrams are not suitable for high-quality artwork such
as posters: they are not vectorized.  Many of the underlying knots
possess a line of symmetry (at least, the diagrams do if the breaks
are ignored), which is not present in the visual representation.
Also, several of the strands cross at acute angles.  The diagram for
knot~$7_3$, for example, contains ugly kinks and abrupt changes of
curvature.

Such considerations suggest that knot diagrams might be produced by
minimization of some objective function that quantifies the visual
inelegance of a knot diagram.  The precise nature of such an objective
function is, of course, a subjective choice but one might require the
following desiderata:

\begin{itemize}
\item Curvature to be as uniform as possible
\item Strands to cross at right angles
\item Crossing points to be separate from one another
\item Any symmetry present in the knot should be enforced exactly, and be visually apparent
\end{itemize}

Knot diagrams may be created using vectorized graphics software such
as inkscape~\cite{kirsanov2009}: one specifies a sequence of control
points, then interpolates between these points to create a knot
diagram with the appropriate topology (inkscape is a widely-used
system available under the GPL).  One way of smoothly interpolating
between specified points is Bezier curves~\cite{olsen2014}.  A Bezier
curve is a visually pleasing polynomial path that can be used to
specify the path of a knot projection.

The software discussed here~\cite{hankin2016} allows one to specify a
knot in terms of its Bezier control points within inkscape, import the
object into R~\cite{rcore2016}, and then to use numerical optimization
techniques to improve the visual appearance of the knot.

\section{The package in use}

In this section, I give workflows for creating two simple knots:
firstly~$7_6$, followed by the figure-of-eight knot~$4_1$ which
requires imposition of symmetry constraints.

The first step is to create a closed curve in inkscape that shows the
rough outline of the knot~(Figure~\ref{screenshot_inkscape_7_6} shows
a screenshot of {\tt 7\_6\_first\_draft.svg}, supplied with the
package).  Note that this file contains only the knot {\em path}; the
over and under information is to be added later.

Here, knot paths are required to have Bezier handles that are
symmetrically placed with regard to nodes.  Although radius of
curvature is not necessarily matched at either side of a node, visual
continuity of the strand is ensured.  In the diagram, strand crossing
points are far from nodes, as this ensures visual continuity of
strands at crossing points, especially the understrand.

\begin{figure}[h]
  \centering
    \includegraphics[width=10cm]{screenshot_inkscape_7_6.png}
\caption{Screenshot of inkscape\label{screenshot_inkscape_7_6} setup
  for knot~$7_6$.  Nodes are shown as squares, handles as small
  circles, symmetrically placed on either side of nodes}
\end{figure}

The knot shown in Figure~\ref{screenshot_inkscape_7_6} is clearly
suboptimal: even though the nodes are connected by Bezier curves which
are individually smooth, the path as a whole is ugly and unattractive
as its path has unsightly regions where the radius of curvature
changes abruptly.

Although it is possible in principle to improve the visual appearance
of the knot path by hand in inkscape, this is a surprisingly difficult
and frustrating task.  In order to remedy the flaws of the diagram
using an automated system, we first read the {\tt .svg} file into R
using the {\tt reader()} function; a typical R session follows:

\begin{verbatim}
> library(knotR)
> k76 <- reader(system.file("7_6_first_draft.svg",package="knotR"))
> head(k76)

              x         y
[1,]  -98.81963 339.81898
[2,] -223.87754 303.35366
[3,] -299.84521 121.06064
[4,] -236.36319  36.35340
[5,] -172.88118 -48.35384
[6,]  -92.86186 -69.78212
\end{verbatim}

Object~{\tt k76} is stored as an object of class {\tt inkscape}: a
two-column matrix with rows corresponding to the node and handle
positions of the inkscape path.  This representation has a certain
amount of redundancy as knot paths have handles which are
symmetrically placed with respect to nodes; also, the first node is
the same as the last for the loop is closed.  The package can coerce
inkscape objects to other forms, specifically {\tt minobj} objects,
which contain no redundancy (the position of each node, as well as one
of the handles, is stored); or {\tt controlpoints} objects, which
allow for easy construction of Bezier interpolation between nodes.

\begin{figure}[htbp]
  \begin{center}
\begin{verbatim}
> k76_rough <- reader(system.file("7_6_first_draft.svg",package="knotR"))
> knotplot2(k76_rough, seg=TRUE)
\end{verbatim}
\begin{centering}
\includegraphics{knot-7_6_rough}
\caption{The \emph{path} of (unoptimized) \label{7_6_rough}
  knot~$7_6$, showing Bezier handles as thin straight lines and
  circles.  The coloured circles have a radius  proportional to the
  curvature (that is, the reciprocal of the radius of curvature) along
  the strand; note large curvature at loop on left}
  \end{centering}
  \end{center}
\end{figure}

\begin{figure}[!tbp]
\begin{verbatim}
knotplot2(k7_6,text=TRUE,lwd=1,circ=0,rainbow=TRUE)
\end{verbatim}
 \centering
\includegraphics{knot-004}
\caption{Knot~$7_6$ with strands numbered \label{k76_strands} to guide
creation of over and under information} \end{figure}

To beautify it we need to specify a function of the path that
quantifies its displeasingness, and then minimize this objective
function using numerical methods.

Two examples of desiderata for such an objective function might be to
keep the strands crossing at right angles, and the overall bending
energy.  These are evaluated in the package by functions
{\tt total\_crossing\_angles()} and~{\tt total\_bending\_energy()}
respectively:

\begin{verbatim}
> b <- as.controlpoints(k76_rough)
> total_crossing_angles(b)

[1] 0.3145033

> total_bending_energy(b)

[1] 0.1276257
\end{verbatim}

The knots supplied in the package minimize a weighted sum of these and
other badnesses\footnote{Function {\tt badness()} includes various
``housekeeping'' badnesses which are used to make sure that the
minimum found by {\tt nlm()} is topologically identical to the
starting configuration.  Function
{\tt non\_crossing\_strand\_close\_approach\_badness()}, for example,
makes non-crossing strands ``repel'' one another so as not to
introduce spurious intersections.}, as evaluated by function
{\tt badness()}:

\begin{verbatim}
> badness(k76_rough)

[1] 6.16279
\end{verbatim}

This function may be minimized by numerical optimization:

\begin{verbatim}
> o <- nlm(badness, as.knotvec(k76_rough))
> k7_6 <- as.minobj(o$estimate)
> badness(k7_6)

  [1] 3.550152
\end{verbatim}

(the above takes about an hour of CPU time: it is optimizing a
function of~64 real variables, and the objective function takes a few
seconds to evaluate).  However, the result is much nicer
(Figure~\ref{7_6}).

\begin{figure}[!tbp]
\begin{verbatim}
knotplot2(k7_6)
\end{verbatim}
  \centering
\includegraphics{knot-k76_knotplot2}
\caption{Knot $7_6$, post-optimization\label{7_6}}
\end{figure}

To specify the senses of the knot's\ %---------------------------------'
crossings, we create an overunder object which is a two-column matrix:

\begin{verbatim}
> ou76 <- matrix(c(
+     12,01,
+     02,11,
+     07,03,
+     04,15,
+     16,06,
+     14,08,
+     10,13
+     ),byrow=TRUE,ncol=2)
\end{verbatim}

With reference to Figure~\ref{k76_strands}, each row of {\tt ou76}
corresponds to a crossing; the first element gives the overstrand and
the second the understrand; thus strand~12 passes over strand~1,
strand~2 passes over strand~11, and so on. The result is shown in
figure~\ref{76_overunder}.

\begin{figure}[!tbp]
\begin{verbatim}
knotplot(k7_6)
\end{verbatim}
  \centering
\includegraphics{knot-009}
\caption{Knot $7_6$, post-optimization with breaks indicating
  \label{76_overunder}  underpassing strands}
\end{figure}

\subsection{Symmetry}

Many of the knots in Figure~\ref{knot_table} have an axis of symmetry,
or possess rotational symmetry.  The package has the ability to impose
symmetry relations on knots, and to optimize the resulting symmetrical
knot.  Minimizing the badness is not entirely straightforward on
account of the induced redundancy, which is characterized using a
symmetry object specific to the knot under consideration.  However,
symmetry constraints do reduce the dimensionality of the optimization
problem.

I will consider the figure-of-eight knot~$4_1$ as an example.  Using
Figure~\ref{four_figure_8_knots}, top left, as reference, the
appropriate symmetry object is defined as follows:

\begin{verbatim}
> fig8 <- reader(system.file("4_1_first_draft.svg",package="knotR"))
> Mver8 <- matrix(c(
+     02,03,
+     09,07,
+     05,11,
+     10,06
+     ),ncol=2,byrow=TRUE)
> sym8 <- symmetry_object(fig8, Mver=Mver8, xver=8)
\end{verbatim}

(Matrix {\tt Mver8} specifies that nodes~2 and~3 are symmetric, as
are nodes~9 and~7, and so on; {\tt xver=8} forces node~8 to be on the
axis of symmetry).  The results are shown in
Figure~\ref{four_figure_8_knots}.

\begin{figure}[!tbp]
  \centering
\includegraphics[width=11cm]{knot-four_figure_eight_knots}
\caption{Figure eight knots drawn using different plotting methods.
  Top left, knot path with node numbers shown in order to facilitate
    \label{four_figure_8_knots} definition of the symmetry object; top
    right, result of symmetrizing the rough path; lower left, the
    optimized knot with imposed vertical symmetry, with curvature
    plotted; lower right, knot plotted with overstrand and understrand
    indicated using line breaks}
\end{figure}

\subsection{Rotational symmetry}

Consider knot~$5_1$. This knot has fivefold rotational symmetry, in
addition to a vertical line of symmetry.  The package includes
functionality to impose appropriate symmetry constraints. Using
Figure~\ref{k5_1_twoknots} as reference, we have:

\begin{figure}[htbp]
  \begin{center}
\begin{verbatim}
> knotplot2(k5_1,node=TRUE,width=FALSE)
\end{verbatim}
\includegraphics[width=11cm]{knot-k5_1}
\caption{Knot $5_1$ \label{k5_1_twoknots} shown with node numbers}
  \end{center}
\end{figure}

\begin{verbatim}
> sym51 <- symmetry_object(k5_1,
+                          Mver = cbind(11,13),
+                          xver = c(2,12),
+                          Mrot = rbind(
+                              c(12,04,16,08,20),
+                              c(13,05,17,09,01),
+                              c(11,03,15,07,19),
+                              c(02,14,06,18,10)
+                          ))
\end{verbatim}

Thus, using the same notation as before, nodes~11 and~13 are
symmetrical about the vertical axis, nodes~2 and~12 are on the
vertical axis.  The {\tt Mrot} argument specifies sets of nodes that
map to themselves under rotation.  The top line of {\tt Mrot}
indicates that nodes 12, 4, 16, 8, and~20 are concyclic.  An example
of a rotationally symmetric knot is given in Figure~\ref{orn20}.

\section{Conclusions and further work}

The {\tt knotR} package allows the user to create rough diagrams of
knots using the inkscape suite of software, and subsequently polish up
such diagrams in terms of a customizable objective function using
numerical optimization techniques.  Further work might include
functionality to deal with links.

\clearpage
\section{Gallery}

There now follows a selection of pleasing knot diagrams taken from
datasets provided with the package.

\begin{figure}[htbp]
  \begin{center}
\begin{verbatim}
par(mfcol=1:2)
knotplot(perko_A)
knotplot(perko_B)
\end{verbatim}
\includegraphics[width=11cm]{knot-perko_A_and_B}
\caption{Two representations of knot~$10_{125}$, known as the 
\label{perko_AB}  Perko Pair}
  \end{center}
\end{figure}

\begin{figure}[htbp]
  \begin{center}
\begin{verbatim}
> knotplot(ornamental20,gap=15)
\end{verbatim}
\includegraphics[width=11cm]{knot-ornamental}
\caption{An ornamental knot exhibiting fivefold
  rotational \label{orn20} symmetry; note the absence of mirror symmetry}
  \end{center}
\end{figure}

\begin{figure}[htbp]
  \begin{center}
\includegraphics[width=13cm]{knots_to_8crossings}
\caption{All \label{all8} prime knots with 8 or fewer crossings,
notation following Rolfsen}
  \end{center}
\end{figure}

\section*{References}

\bibliography{knot}
\bibliographystyle{unsrt}
\end{document}

% ------------------------------------------------------------------------

% ------------------------------------------------------------------------
\end{document}
% ------------------------------------------------------------------------

### 2020-11-04 inst/jma_knot.tex
Journal of Mathematics in the Arts submission kickoff
scale the diagrams so they fit nicely on the page
minor rewriting for JMA

@@
 homeomorphism~$h\colon\mathbb{R}^3\longrightarrow\mathbb{R}^3$ which
-takes one embedded circle to the other.
+takes one embedded circle to the other.  In the art world, knot theory
+has applications to the study of dance~\cite{khorami2020}, sculpture,
+\cite{bosch2010, widmark2020}, and many other branches of mathematical
+art~\cite{hart2008}.
 
@@
 knot is given with broken lines indicating where one strand passes
-over another.  Related work in the same language is given by Comoglio
-and Rinaldi~\cite{comoglio2011,comoglio2012}; the visual depiction of
-knot projections in the context of Celtic artwork is given by
-Browne~\cite{browne2005,browne2006}.
+over another.  Knot diagrams figure prominently in Celtic
+artwork~\cite{cromwell2008} and are discussed from a mathematical
+perspective by Browne~\cite{browne2005,browne2006}.
 
@@
 
-The software discussed here~\cite{hankin2016} allows one to specify a
+The software discussed here~\cite{hankin2020} allows one to specify a
 knot in terms of its Bezier control points within inkscape, import the
@@
 \begin{centering}
-\includegraphics{knot-7_6_rough}
+\includegraphics[scale = 0.9]{knot-7_6_rough}
 \caption{The \emph{path} of (unoptimized) \label{7_6_rough}
@@
  \centering
-\includegraphics{knot-004}
+\includegraphics[scale = 0.9]{knot-004}
 \caption{Knot~$7_6$ with strands numbered \label{k76_strands} to guide
@@
   \centering
-\includegraphics{knot-k76_knotplot2}
+\includegraphics[scale = 0.9]{knot-k76_knotplot2}
 \caption{Knot $7_6$, post-optimization\label{7_6}}
@@
   \centering
-\includegraphics{knot-009}
+\includegraphics[scale = 0.9]{knot-009}
 \caption{Knot $7_6$, post-optimization with breaks indicating
@@
   \begin{center}
-\includegraphics[width=13cm]{knots_to_8crossings}
+\includegraphics[width=13cm]{knots_to_8crossings}  % created by inst/knotter.R
 \caption{All \label{all8} prime knots with 8 or fewer crossings,

### 2020-11-06 inst/jma_knot.tex
tweaks

@@
 %----------classification, keywords, date
-\subjclass{Primary 99Z99; Secondary 00A00}
+\subjclass{Primary 57K10-XX; 32-XX}
 
@@
 \begin{abstract}
-In this short article I introduce software which creates two
-dimensional knot diagrams optimized for visual appearance.  Different
-aspects of a knot's appearance are discussed and a framework for
-objectively optimizing the visual appeal of a knot projection is
-given.
-
-In this short article I show how to create two dimensional knot
-diagrams optimized for visual appearance using the R programming
-language.  The package associated with this paper is a suite
-of software for the creation of production-quality artwork of knot
-diagrams, released under GPL2.
+In this short article I discuss the \ae sthetics of knot projections
+and introduce software which creates two dimensional knot diagrams
+optimized for visual appearance.  Different aspects of knot appearance
+are discussed and a framework for objectively optimizing the visual
+appeal of a knot projection is given.   I use the software to create 
+a wide range of knot diagrams.
 \end{abstract}
@@
 circle~$\mathbb{S}^1$ into~$\mathbb{R}^3$~\cite{manturov2004}.  Two
-knots are said to be equivalent if one can be continuously deformed in
-to the other; if so, there is a
+knots are said to be equivalent if the embedding of one can be
+continuously deformed to the other; if so, there is a
 homeomorphism~$h\colon\mathbb{R}^3\longrightarrow\mathbb{R}^3$ which
-takes one embedded circle to the other.  In the art world, knot theory
+takes one embedded circle to the other.  Knot theory
+
+In the art world, knot theory
 has applications to the study of dance~\cite{khorami2020}, sculpture,
@@
 
+One plausible technique for creating knot projections is to consider a
+two-dimensional projection of a knot's embedding
+$\mathbb{E}\in\mathbb{R}^3$.  However, this approach often results in
+poor visual appearance: cusps or other displeasing effects can occur.
+
 \section{The package in use}
@@
 Here, knot paths are required to have Bezier handles that are
-symmetrically placed with regard to nodes.  Although radius of
-curvature is not necessarily matched at either side of a node, visual
-continuity of the strand is ensured.  In the diagram, strand crossing
-points are far from nodes, as this ensures visual continuity of
-strands at crossing points, especially the understrand.
+symmetrically placed with regard to nodes (this is a user-settable
+option in inkscape).  Although radius of curvature is not necessarily
+matched at either side of a node, visual continuity of the strand is
+ensured.  In the diagram, strand crossing points are far from nodes,
+as this ensures visual continuity of strands at crossing points,
+especially the understrand.
 

### 2021-09-11 inst/jma_rebuttal.tex
work with a latex document for rebuttal instead, issue #15

Sep 08, 2021

Ref.:  Ms. No. TMAA-2020-0042
Visually pleasing knot projections
Journal of Mathematics and the Arts

Dear Robin Hankin,

Reviewers have now commented on your paper.  You will see that they
are advising that you revise your manuscript.  If you are prepared to
undertake the work required, I would be pleased to review a revision.

For your guidance, reviewers' comments are appended below.

If you decide to revise the work, please submit a list of changes or a
rebuttal against each point which is being raised when you submit the
revised manuscript.

Your revision is due by Dec 08, 2021.

To submit a revision, go to
https://rp.tandfonline.com/submission/flow?submissionId=209604076&step=1. If
you decide to revise the work, please submit a list of changes or a
rebuttal against each point which is being raised when you submit the
revised manuscript.

If you have any questions or technical issues, please contact the
journal's editorial office at SP-ingest@journals.tandf.co.uk.

Yours sincerely,

Susan Goldstine, Ph.D.
Associate Editor
Journal of Mathematics and the Arts

Comments from the Editors and Reviewers:

Reviewer #1:

This short article is a tutorial example of how to use Inkscape and
KnotR to beautify 2-dimensional knot diagrams found in the knot
tables.  The cost function used in optimizing a given diagram tries to
minimize the overall bending energy of the knot curve, preserve all
shown 2D symmetries, and render all strand crossings as close to 90
degrees as possible.  The paper does not discuss the underlying
mathematics by which this optimization is achieved, but relies on the
power built into the software packages mentioned.  Also, the user has
to characterize explicitly the desired symmetries; there is no "magic"
power that discovers the maximal possible symmetry that a given knot
can assume.  This exemplified by Figure 7, where the Figure-8-knot is
drawn with a single mirror axis, even though this same knot can
readily be drawn with two mirror lines; -- and also in Figure 9 where
a symmetrical and an asymmetrical rendering of the same knot is
presented.

This paper is clearly useful to someone who specifically wants to
clean up a given rendering of a particular knot, but it offers little
to readers who are not familiar with Inkscape and KnotR.

Detailed comments about the presentation:

Abstract: It should mention the software packages that are used in
this paper.

P1, line 46:  "Knot theory" -- ?  -- Is this an incomplete reference?

P2, line 43: "uniform" curvature is not possible, since many knots
have inflection points.  It is better to ask for 'smoothly changing'
curvature, and to limit maximal curvature.

P2, line 46: What does it mean for symmetry to be "present" in the
knot?

P3, line 31: Define "visual continuity".  Are you referring to tangent
continuity, G1-, or G2-continuity?

P3, line 33:  The "understrand" has not been designated yet.

Figure 2:  Specify what order Bezier curves are being used.

P4, line 18: A gentle introduction to "R" would help to make the
following easier to understand.

P4, lines 27-32: What is the meaning of this? - Six points defining
two handle pairs?  (Why the extreme precision with 8 digits?)

Figure 3: What is the color assignment? - one color for each Bezier
segment?  How many circles are drawn for each segment?  (Say that
solid blotches result from densely overlapping circles.)

Figure 4: Why is this required?  -- Does the user have to specify the
type of crossing via an "overunder" object?  -- (Lines 53-62?)

P10, line 17: "... a vertical line of symmetry." - Better to say that
this figure has D_5 symmetry with five mirror lines.

P11, line 26: It seems that all the functionality needed to deal with
links is already present in the software packages mentioned, as long
as one can let the program go around more than one single closed loop.

Fig. 10: Why is the absence of mirror symmetry pointed out here?

Fig. 11: Why is this table shown?  It is not mentioned in the main
text.  The figures are "better" than in Figure 1, but still not
optimal in the spirit of this paper.  "Rolfsen" needs a reference.

References to some of the background material are somewhat random.
E.g.  For JMA readers, "The Knot Book" by Colin Adams would be a good
introductory reference.

Ref [10]  Give a more up-to-date access date.

Ref [11]  Give URL of  Inkscape:  https://inkscape.org/

Ref [14]  Give URL of  R Core Team:  https://www.r-project.org/

>>>>>>>>>>>>>>

Reviewer #2: This lovely article addresses the interesting question of
how to produce pleasing and informative images of two dimensional knot
diagrams. In addition to the discussion it describes how to use a
combination of the software packages Inkscape and R to optimise knots.

At the moment, however the paper is a little too close to simply being
a software manual for the software, and needs in particular to
describe the techniques used by the software to create the optimised
images.

In addition there is a rather consistent use of absolute terms for
pleasing vs ugly knots. Aesthetics is a complex field, but generally
does not provide strong answers. While the choices made in this paper
are reasonable, they are simply stated, rather than including a
justification for why they are chosen. They certainly only apply to a
particular knid of knot diagram. This can be see in the final image
showing all the knots with up to 8 crossings. While some of the
resulting diagrams are beautiful, I particularly enjoyed 8_10 and
8_17, some are not so pleasing to me. In particular the algorithm
given likes to over-tighten when the line wraps round itself. This can
be seen particularly in 8_3, but also in 8_5, 8_6 and 8_11. This is
itself a personal preference not an absolute, but a discussion of how
to adjust the set-up of the software to make personal aesthetic
choices would strengthen the paper.

The paper also needs to do a far stronger job of placing itself within
the relevant literature. Simply showing that a single diagram from
wikipedia has weaknesses is not sufficient. The nineteenth century
series of papers "On Knots" by Peter Guthrie Tait in the Transactions
of the Royal Society of Edinburgh contain far superior images for
example. This is not the first paper to look at making pleasing knot
diagrams, the software knotplot for example offers many tools to do a
similar optimisation to the one discussed here, but is not cited,
Laura Taalman and Colin Abrams among others have multiple papers
addressing the creation of different knot representations and
optimising their form. More broadly the idea of using optimisation in
mathematical art has been used widely, the work of Robert Bosch and
Craig Kaplan springing immediately to mind.

For these reasons the current paper is not really strong enough for
publication. In order to get strong enough I would propose three major
pieces of work:

1) Conduct a thorough literature review of work on the creation of
knot diagrams and their aesthetics and place this paper into that
context as well as the broader context of optimisation in art.

2) Include a discussion of why the aesthetic choices made for the
optimisation are pleasing as well as how to adjust the methodology in
order to make different personal choices. It might even be interesting
to develop some distinct styles of diagrams (by changing the "badness"
function, and use the methods to create optimised images for each
style.

3) Give basic information about the specific methods being used for
the optimisation rather than just using the function given in R as a
black box.

With these changes the paper could be an interesting one, discussing
the context of knot diagrams more broadly and providing practical help
to help create sets for oneself.

__________________________________________________
In compliance with data protection regulations, you may request that we remove your personal registration details at any time.  (Use the following URL: https://www.editorialmanager.com/tmaa/login.asp?a=r). Please contact the publication office if you have any questions.

### 2021-09-11 inst/jma_rebuttal.tex
work on issue #16

@@
-Sep 08, 2021
+\documentclass[12pt]{article}
+\usepackage{xcolor}
+\begin{document}
 
-Ref.:  Ms. No. TMAA-2020-0042
-Visually pleasing knot projections
-Journal of Mathematics and the Arts
+\section*{Rebuttal to second editorial review: overall response}
 
-Dear Robin Hankin,
+(Below, the reviewer's comments are in black, and my replies to the
+issues are in \textcolor{blue}{blue}.  I have indicated changes to the
+manuscript where appropriate.
 
-Reviewers have now commented on your paper.  You will see that they
-are advising that you revise your manuscript.  If you are prepared to
-undertake the work required, I would be pleased to review a revision.
+Short story: I have accommodated all the comments with rewording.  The
+resulting document is, I believe, stronger and more scholarly than
+before and I recommend it to you.
 
-For your guidance, reviewers' comments are appended below.
-
-If you decide to revise the work, please submit a list of changes or a
-rebuttal against each point which is being raised when you submit the
-revised manuscript.
-
-Your revision is due by Dec 08, 2021.
-
-To submit a revision, go to
-https://rp.tandfonline.com/submission/flow?submissionId=209604076&step=1. If
-you decide to revise the work, please submit a list of changes or a
-rebuttal against each point which is being raised when you submit the
-revised manuscript.
-
-If you have any questions or technical issues, please contact the
-journal's editorial office at SP-ingest@journals.tandf.co.uk.
-
-Yours sincerely,
-
-Susan Goldstine, Ph.D.
-Associate Editor
-Journal of Mathematics and the Arts
-
-Comments from the Editors and Reviewers:
-
-Reviewer #1:
+\section*{Detailed rebuttal: Reviewer \#1}
 
@@
 power built into the software packages mentioned.  Also, the user has
-to characterize explicitly the desired symmetries; there is no "magic"
+to characterize explicitly the desired symmetries; there is no ``magic"
 power that discovers the maximal possible symmetry that a given knot
@@
 presented.
+
+\textcolor{blue}{This is a good characterisation of the submission,
+  but the comment that there is no ``magic" power ascertaining
+  maximal possible symmetry that a given knot may assume is a good
+  one, and one that I had not explicitly realised.  I have added a
+  brief discussion of this issue to the caption of Figure 9 that
+  discusses the Perko pair:
+  \begin{description}
+    \item[OLD text] ``Two representations of knot~$10_{125}$, known as the 
+     Perko Pair"
+    \item[NEW text] ``Two representations of knot~$10_{125}$, known as the 
+      Perko Pair.   The software requires the user to specify the symmetry
+      (mirror or rotational) of a knot projecton and has no notion of
+      topological invariance of a knot"
+      \end{description}
+}
 
@@
 
-P1, line 46:  "Knot theory" -- ?  -- Is this an incomplete reference?
+P1, line 46:  ``Knot theory" -- ?  -- Is this an incomplete reference?
 
-P2, line 43: "uniform" curvature is not possible, since many knots
+P2, line 43: ``uniform" curvature is not possible, since many knots
 have inflection points.  It is better to ask for 'smoothly changing'
@@
 
-P2, line 46: What does it mean for symmetry to be "present" in the
+P2, line 46: What does it mean for symmetry to be ``present" in the
 knot?
 
-P3, line 31: Define "visual continuity".  Are you referring to tangent
-continuity, G1-, or G2-continuity?
+P3, line 31: Define ``visual continuity".  Are you referring to tangent
+continuity, $G1-$, or $G2-$ continuity?
 
-P3, line 33:  The "understrand" has not been designated yet.
+P3, line 33:  The ``understrand" has not been designated yet.
 
@@
 
-P4, line 18: A gentle introduction to "R" would help to make the
+P4, line 18: A gentle introduction to ``R" would help to make the
 following easier to understand.
@@
 
-P10, line 17: "... a vertical line of symmetry." - Better to say that
-this figure has D_5 symmetry with five mirror lines.
+P10, line 17: ``... a vertical line of symmetry." - Better to say that
+this figure has $D_5$ symmetry with five mirror lines.
 
@@
 Fig. 11: Why is this table shown?  It is not mentioned in the main
-text.  The figures are "better" than in Figure 1, but still not
-optimal in the spirit of this paper.  "Rolfsen" needs a reference.
+text.  The figures are ``better" than in Figure 1, but still not
+optimal in the spirit of this paper.  ``Rolfsen" needs a reference.
 
 References to some of the background material are somewhat random.
-E.g.  For JMA readers, "The Knot Book" by Colin Adams would be a good
+E.g.  For JMA readers, ``The Knot Book" by Colin Adams would be a good
 introductory reference.
@@
 
-Ref [11]  Give URL of  Inkscape:  https://inkscape.org/
+Ref [11]  Give URL of  Inkscape:  {\tt https://inkscape.org/}
 
-Ref [14]  Give URL of  R Core Team:  https://www.r-project.org/
+Ref [14]  Give URL of  R Core Team:  {\tt https://www.r-project.org/}
 
->>>>>>>>>>>>>>
+\section*{Detailed rebuttal: Reviewer \#2}
 
-Reviewer #2: This lovely article addresses the interesting question of
+This lovely article addresses the interesting question of
 how to produce pleasing and informative images of two dimensional knot
@@
 showing all the knots with up to 8 crossings. While some of the
-resulting diagrams are beautiful, I particularly enjoyed 8_10 and
-8_17, some are not so pleasing to me. In particular the algorithm
-given likes to over-tighten when the line wraps round itself. This can
-be seen particularly in 8_3, but also in 8_5, 8_6 and 8_11. This is
-itself a personal preference not an absolute, but a discussion of how
-to adjust the set-up of the software to make personal aesthetic
-choices would strengthen the paper.
+resulting diagrams are beautiful, I particularly enjoyed {\tt 8\_10}
+and {\tt 8\_17}, some are not so pleasing to me. In particular the
+algorithm given likes to over-tighten when the line wraps round
+itself. This can be seen particularly in {\tt 8\_3}, but also in {\tt
+  8\_5}, {\tt 8\_6} and {\tt 8\_11}. This is itself a personal
+preference not an absolute, but a discussion of how to adjust the
+set-up of the software to make personal aesthetic choices would
+strengthen the paper.
 
@@
 wikipedia has weaknesses is not sufficient. The nineteenth century
-series of papers "On Knots" by Peter Guthrie Tait in the Transactions
+series of papers ``On Knots" by Peter Guthrie Tait in the Transactions
 of the Royal Society of Edinburgh contain far superior images for
@@
 order to make different personal choices. It might even be interesting
-to develop some distinct styles of diagrams (by changing the "badness"
+to develop some distinct styles of diagrams (by changing the ``badness"
 function, and use the methods to create optimised images for each
@@
 
-__________________________________________________
-In compliance with data protection regulations, you may request that we remove your personal registration details at any time.
-(Use the following URL: https://www.editorialmanager.com/tmaa/login.asp?a=r).
-Please contact the publication office if you have any questions.
+\end{document}

### 2021-09-11 inst/jma_knot.tex
work on issue #16
fixes #18

@@
 and introduce software which creates two dimensional knot diagrams
-optimized for visual appearance.  Different aspects of knot appearance
-are discussed and a framework for objectively optimizing the visual
-appeal of a knot projection is given.   I use the software to create 
-a wide range of knot diagrams.
+optimized for visual appearance.  The software is in the form of a
+documented and self-contained suite of functionality written in the R
+programming language (that is, a ``package"): {\tt knotR}.  The
+package leverages the graphical capabilities of the popular vectorised
+graphics software {\tt inkscape}.  Different aspects of knot
+appearance are discussed and a framework for objectively optimizing
+the visual appeal of a knot projection is given.  I use the software
+to create a wide range of knot diagrams.
 \end{abstract}
@@
 \caption{Two representations of knot~$10_{125}$, known as the 
-\label{perko_AB}  Perko Pair}
+  \label{perko_AB}  Perko Pair.  The software requires the user to
+  specify the symmetry (mirror or rotational) of a knot projecton and
+  has no notion of topological invariance of a knot}
   \end{center}

### 2021-09-12 inst/jma_rebuttal.tex
fixes #18
fixes #19
fixes #20
fixes #37

@@
 
+\textcolor{blue}{Done.  The abstract is now completely rewritten.}
+
 P1, line 46:  ``Knot theory" -- ?  -- Is this an incomplete reference?
+
+\textcolor{blue}{Corrected, this was a typo}
 
@@
 curvature, and to limit maximal curvature.
+
+\textcolor{blue}{Corrected:
+  \begin{description}
+  \item[OLD text] Curvature to be as uniform as possible
+\item[NEW text] Curvature to be as smoothly changing as possible, with
+  limited maximal curvature
+  \end{description}
+}
 
@@
 
+\textcolor{blue}{This is indeed an insightful comment.  My original
+  view was that there was a single set of weightings for the various
+  badnesses, that most people would agree upon; and my task as
+  software engineer was to find the projection that optimizes these
+  standard weightings.\\ \\ Following this comment, I realise with a
+  shock that my weightings do not reflect some universal ideal.
+  Rather, my choice of weightings are as subjective as any other
+  aspect of the software and, as such, are peculiar to me.  This is
+  obvious in retrospect but was not clear to me at all until just
+  now.\\ \\ One of the benefits of the {\tt knotR} software is that it
+  allows this kind of conversation to occur; I have added a discussion
+  to the manuscript, just before the conclusions section.  There, I
+  present projections of $8_3$ and $8_5$ but with different penalties
+  for angle crossing terms.  We see the results and can compare like
+  with like, perhaps coming to a deeper understanding of pleasingness
+  of projections in the process.\\ \\}
+
 The paper also needs to do a far stronger job of placing itself within

### 2021-09-12 inst/jma_knot.tex
fixes #20
fixes #37

@@
 homeomorphism~$h\colon\mathbb{R}^3\longrightarrow\mathbb{R}^3$ which
-takes one embedded circle to the other.  Knot theory
-
-In the art world, knot theory
+takes one embedded circle to the other.  In the art world, knot theory
 has applications to the study of dance~\cite{khorami2020}, sculpture,
@@
 \begin{itemize}
-\item Curvature to be as uniform as possible
+\item Curvature to be as smoothly changing as possible, with limited maximal curvature
 \item Strands to cross at right angles
@@
 minimum found by {\tt nlm()} is topologically identical to the
-starting configuration.  Function
-{\tt non\_crossing\_strand\_close\_approach\_badness()}, for example,
+starting configuration.  Function {\tt
+  non\_crossing\_strand\_close\_approach\_badness()}, for example,
 makes non-crossing strands ``repel'' one another so as not to
-introduce spurious intersections.}, as evaluated by function
-{\tt badness()}:
+introduce spurious intersections.}.  The weightings for the various
+badnesses are, of course, subjective; but the system discussed here
+allows the user to change the weightings used and compare results.  I
+present a short discussion in section~\ref{section_cca} below.
+Numeically, the badnesses are evaluated by function {\tt badness()}:
 
@@
 
+\subsection{Subjective \label{section_cca} choice for weighting}
+
+One of the benefits of the {\tt knotR} software is that it allows the
+relative importance of the different aspects of appearance to be
+assessed.  In the package, the badness weightings may be altered
+easily, and in this way we can investigate the visual impact of the
+badness components.  The default weightings were originally chosen as
+a bland compromise, but it is easy to place greater weight on one or
+other compoent.  Figure~\ref{figure_cca} shows the same knot optimized
+using different weightings; we see the effect of imposing right-angled
+strand crossing angles on the remainder of the knot.  Readers will
+have to judge for themselves which version they prefer.
+
+\begin{figure}[htbp]
+  \begin{center}
+\includegraphics[width=13cm]{compare_crossing_angles}  % created by
+                                                       % inst/knotter.R
+\caption{Comparison between optimized knot projections for $8_3$
+  \label{figure_cca} using different  badness weightings.  On the
+  left we see the default knot, and on the right we see the result of
+  increasing the angle-crossing penalty by a factor of 100,
+  effectively forcing all strand crossing angles to be $\pi/2$.  Which
+  one is preferable is, of course, a subjective choice}
+  \end{center}
+\end{figure}
+
 \section{Conclusions and further work}

### 2021-09-14 inst/jma_rebuttal.tex
fixes issue #27

@@
 solid blotches result from densely overlapping circles.)
+
+\textcolor{blue}{The original caption was lacking detail.
+  \begin{description}
+  \item[OLD text] ``The \emph{path} of (unoptimized) knot~$7_6$,
+    showing Bezier handles as thin straight lines and circles.  The
+    coloured circles have a radius proportional to the curvature (that
+    is, the reciprocal of the radius of curvature) along the strand;
+    note large curvature at loop on left"
+  \item[NEW text] ``The \emph{path} of (unoptimized) knot~$7_6$,
+    showing Bezier handles as thin straight lines and circles.  The
+    coloured circles have a radius proportional to the curvature (that
+    is, the reciprocal of the radius of curvature) along the strand.
+    Colouring is arbitrary, one color for each Bezier segment; solid
+    blotches result from densely overlapping circles.  Note large
+    curvature at loop on left"
+  \end{description}
+  Each segment has one circle per control point (which defaults to
+  $n=100$) but this is user-settable.
+}
 

### 2021-09-14 inst/jma_knot.tex
fixes issue #27

@@
 \end{verbatim}
+
 \begin{centering}
@@
   curvature (that is, the reciprocal of the radius of curvature) along
-  the strand; note large curvature at loop on left}
+  the strand.  Colouring is arbitrary, one color for each Bezier
+  segment; solid blotches result from densely overlapping circles.
+  Note large curvature at loop on left}
   \end{centering}

### 2021-09-14 inst/jma_knot.tex
fixes #31
fixes #26

@@
 two-column matrix with rows corresponding to the node and handle
-positions of the inkscape path.  This representation has a certain
-amount of redundancy as knot paths have handles which are
-symmetrically placed with respect to nodes; also, the first node is
-the same as the last for the loop is closed.  The package can coerce
-inkscape objects to other forms, specifically {\tt minobj} objects,
-which contain no redundancy (the position of each node, as well as one
-of the handles, is stored); or {\tt controlpoints} objects, which
-allow for easy construction of Bezier interpolation between nodes.
+positions of the inkscape path.  Above we see only the first six lines
+of the object.  Inkscape representation has a certain amount of
+redundancy as knot paths have handles which are symmetrically placed
+with respect to nodes; also, the first node is the same as the last
+for the loop is closed.  The package can coerce inkscape objects to
+other forms, specifically {\tt minobj} objects, which contain no
+redundancy (the position of each node, as well as one of the handles,
+is stored); or {\tt controlpoints} objects, which allow for easy
+construction of Bezier interpolation between nodes.
 
@@
 
-Consider knot~$5_1$. This knot has fivefold rotational symmetry, in
-addition to a vertical line of symmetry.  The package includes
-functionality to impose appropriate symmetry constraints. Using
-Figure~\ref{k5_1_twoknots} as reference, we have:
+Consider knot~$5_1$. This knot has $D_5$ symmetry with five mirror
+lines.  The package includes functionality to impose appropriate
+symmetry constraints. Using Figure~\ref{k5_1_twoknots} as reference,
+we have:
 

### 2021-09-14 inst/jma_knot.tex
fixes #31
issue #32

@@
 \caption{An ornamental knot exhibiting fivefold
-  rotational \label{orn20} symmetry; note the absence of mirror symmetry}
+  rotational \label{orn20} symmetry $C_5$; note the absence of mirror
+  symmetry which would impose $D_5$}
   \end{center}
@@
 \caption{All \label{all8} prime knots with 8 or fewer crossings,
-notation following Rolfsen}
+notation following Rolfsen~\cite{rolfsen2003}}
   \end{center}

### 2021-09-14 inst/jma_knot.tex
typo

@@
   \label{perko_AB}  Perko Pair.  The software requires the user to
-  specify the symmetry (mirror or rotational) of a knot projecton and
+  specify the symmetry (mirror or rotational) of a knot projection and
   has no notion of topological invariance of a knot}

### 2021-09-15 inst/jma_rebuttal.tex
fixes #31
fixes #26
issue #32
fixes #21

@@
 
+\textcolor{blue}{
+  \begin{description}
+  \item[OLD text] Any symmetry present in the knot should be enforced exactly, and be visually apparent
+\item[NEW text] Any symmetry desired in the knot should be enforced exactly, and be visually apparent
+  \end{description}
+}
+
 P3, line 31: Define ``visual continuity".  Are you referring to tangent
@@
 two handle pairs?  (Why the extreme precision with 8 digits?)
+
+\textcolor{blue}{Only the first six lines are shown for brevity, I have
+  changed the text accordingly: ``Above we see only the first six lines
+of the object"}
 
@@
 
+\textcolor{blue}{Done.}
+
 P11, line 26: It seems that all the functionality needed to deal with
@@
 optimal in the spirit of this paper.  ``Rolfsen" needs a reference.
+
+\textcolor{blue}{Rolfsen now cited.}
 

### 2021-09-16 inst/jma_knot.tex
fixes #21
fixes #34

@@
 \item Crossing points to be separate from one another
-\item Any symmetry present in the knot should be enforced exactly, and be visually apparent
+\item Any symmetry desired in the knot should be enforced exactly, and be visually apparent
 \end{itemize}
@@
 Knot diagrams may be created using vectorized graphics software such
-as inkscape~\cite{kirsanov2009}: one specifies a sequence of control
+as inkscape~\cite{Inkscape}: one specifies a sequence of control
 points, then interpolates between these points to create a knot
@@
 knot in terms of its Bezier control points within inkscape, import the
-object into R~\cite{rcore2016}, and then to use numerical optimization
+object into R~\cite{rcore2021}, and then to use numerical optimization
 techniques to improve the visual appearance of the knot.

### 2021-09-16 inst/jma_knot.tex
fixes #29

@@
 Many of the knots in Figure~\ref{knot_table} have an axis of symmetry,
-or possess rotational symmetry.  The package has the ability to impose
+or possess rotational symmetry; some have both and thus have a
+dihedral symmetry group.  The package has the ability to impose these
 symmetry relations on knots, and to optimize the resulting symmetrical

### 2021-09-16 inst/jma_rebuttal.tex
fixes #34
fixes #30

@@
 
+\textcolor{blue}{I have spent many many hours pondering how to deal
+  with links.  As the referee says, much of the required functionality
+  is (in principle) already present in the software.  Each component
+  would have its own badness, and in addition there would be
+  $n(n-1)/2$ inter-component interaction terms measuring features such
+  as closeness between non-intersecting strands of different
+  components.  Several problems stood out: firstly, the problem of
+  inter-component points.  Currently this is a (symmetrical) Boolean
+  matrix of strands with rows and columns corresponding to strands,
+  and entries corresponding to whether that strand pair intersects or
+  not.  The link generalization would be a symmetric matrix of
+  intersection matrices: an object with four indices and entry
+  (i,j,k,l) indicating whether strand j of component i intersects
+  strand k of component l.  I found this object to be unweildy to work
+  with and difficult to manipulate.  The second problem was the
+  imposition of symmetry.  Sometimes one might desire that some
+  components of a link have a particular kind of symmetry and others
+  to have a different kind (or no!) symmetry.  One unexpectedly
+  difficult problem was that one might have a subset of links that
+  individually possess no symmetry but collectively possess mirror-,
+  rotational-, or indeed dihedral- symmetry.  Thistlethwaite's L4a1 is
+  problematic: one would expect the two components to be identical
+  except for a 90-degree rotation, and implementing this in the
+  context of the package I found impossible. L6a3 is similarly
+  difficult.  We would need new badnesses too.  Consider L6a4; the
+  smaller component should be not only as circular as possible (?),
+  but also one might desire the overall length to be smaller than that
+  of the other component.\\ My overall view was that dealing with
+  links would be possible in principle but, due to a number of rather
+  poorly delineated reasons (which are not really suitable for
+  inclusion in a scholarly publication), an order of magnitude harder
+  than the single-component links considered in the submission.  I'm
+  not saying it's impossible, but just very difficult and, for me,
+  firmly in the category of ``further work''.  I hope that's OK.}
+
 Fig. 10: Why is the absence of mirror symmetry pointed out here?
@@
 Ref [14]  Give URL of  R Core Team:  {\tt https://www.r-project.org/}
+
+\textcolor{blue}{All amended as suggested}
 

### 2021-09-24 inst/jma_rebuttal.tex
fixes #39

@@
 knot diagrams and their aesthetics and place this paper into that
-context as well as the broader context of optimisation in art.
+context as well as the broader context of optimisation in art. % issue~#39
+
+\textcolor{blue}{I have added a literature review as suggested that
+  discusses knot diagrams from early Celtic artwork through to
+  mathematica.\\ \\ The referee makes the perfectly reasonable
+  suggestion of discussing the broader context of optimization of art.
+  I will address this comment colloquially.  It is obvious to me, and
+  clearly to the referee as well, that numerical optimization
+  techniques are well-suited to ``rounding off rough edges'' from a
+  wide range of applications; and further that numerical optimization
+  will produce ``nice looking'' results.  Is is also reasonable to
+  assume that such techniques will have been used in the past by
+  graphical artists.  But I can find no evidence of such activity,
+  after an extensive online search and consulting with two librarians
+  (mathematics and art subject specialists).  Perhaps we are using the
+  wrong search terms?  I would gladly take suggestions on previous
+  work to cite.  I have added a sentence to the effect that this seems
+  to be new to the ``further work'' section.}
 

### 2021-09-24 inst/jma_knot.tex
fixes #33
fixes #39

@@
 A {\em mathematical knot} is a smooth, unoriented embedding of a
-circle~$\mathbb{S}^1$ into~$\mathbb{R}^3$~\cite{manturov2004}.  Two
+circle~$\mathbb{S}^1$ into~$\mathbb{R}^3$~\cite{manturov2004,adams2004}.  Two
 knots are said to be equivalent if the embedding of one can be
@@
 knot is given with broken lines indicating where one strand passes
-over another.  Knot diagrams figure prominently in Celtic
-artwork~\cite{cromwell2008} and are discussed from a mathematical
-perspective by Browne~\cite{browne2005,browne2006}.
+over another.  Knot diagrams have a long history:
+Przytycki~\cite{przytycki1998} discusses an intricate braid used as a
+seal sign in around 1700\,BCE; they figure prominently in {\em The
+  Book of Kells} and the Lindisfarne gospels, graphic art produced
+around 800\,CE~\cite{bain1973}.  Such artwork makes extensive use of
+symmetry~\cite{cromwell2008}.
+
+Knot diagrams are ubiquitous in the modern discipline of mathematical
+knot theory.  In the nineteeth century we see Tait~\cite{tait1884}
+publishing an early systematic study of knots including a beautiful
+plate of elegantly hand-drawn cursive knot diagrams.  The tradition
+continues into the twentieth century with Rolfsen~\cite{rolfsen1976}
+presenting an appendix containing many knot and link projections.
+Rolfsen's diagrams are used widely today: Adams~\cite{adams2004}
+reproduces his work extensively, as do many more recent works.  Online
+resources are becoming more widely available, with one prominent
+example being the {\em Knot Atlas}~\cite{knot_atlas} presenting
+extensive knot tables following Rolfsen~\cite{rolfsen1976}---for knots
+up to~10 crossings---and Thistlethwaite~\cite{hoste1998}, for knots
+of~11 crossings.  Symbolic software such as Mathematica and Maple can
+be used to produce representations of knots, typically in the form of
+nonstandard embeddings of the torus into~$\mathcal{R}^3$; but as
+discussed below this approach is suboptimal from the perspective of producing a
+visually appealing projection.
 
@@
 numerical optimization techniques.  Further work might include
-functionality to deal with links.
+functionality to deal with links.  In the broader context of
+optimisation in art, we observe that numerical optimisation techniques
+can produce aesthetically pleasing results, an observation that might
+find uptake by graphic artists.
 
@@
 \caption{All \label{all8} prime knots with 8 or fewer crossings,
-notation following Rolfsen~\cite{rolfsen2003}}
+notation following Rolfsen~\cite{rolfsen1976}}
   \end{center}

### 2021-09-24 inst/jma_knot.tex
fixes #23

@@
 matched at either side of a node, visual continuity of the strand is
-ensured.  In the diagram, strand crossing points are far from nodes,
-as this ensures visual continuity of strands at crossing points,
-especially the understrand.
+thereby ensured near the node.  At a crossing point it is important
+that both the unbroken and broken strands (designated ``overstrand''
+and ``understrand'' respectively) are smooth.  We would like strand
+crossing points to be far from nodes, as this ensures visual
+continuity of strands at crossing points, especially the understrand.
 

### 2021-09-24 inst/jma_rebuttal.tex
fixes #23
fixes #24
fixes #17

@@
 
+\textcolor{blue}{In any software, one makes use of previously written
+  work.  One rarely writes an operating system from scratch, for
+  example.  The knotR package leverages the design capabilities of
+  inkscape in, I think, a useful and informative way; and it further
+  uses the numerical abilities of R in an efficient and natural way.}%issue17
+
 Detailed comments about the presentation:
@@
 
+\textcolor{blue}{fixed}%issue#23
+
 Figure 2:  Specify what order Bezier curves are being used.
+
+\textcolor{blue}{fixed}%issue#24
 

### 2021-09-24 inst/jma_knot.tex
fixes #24
fixes #22

@@
 curve is a visually pleasing polynomial path that can be used to
-specify the path of a knot projection.
+specify the path of a knot projection; here cubic Bezier curves are
+used.
 
@@
 symmetrically placed with regard to nodes (this is a user-settable
-option in inkscape).  Although radius of curvature is not necessarily
-matched at either side of a node, visual continuity of the strand is
-thereby ensured near the node.  At a crossing point it is important
-that both the unbroken and broken strands (designated ``overstrand''
-and ``understrand'' respectively) are smooth.  We would like strand
-crossing points to be far from nodes, as this ensures visual
-continuity of strands at crossing points, especially the understrand.
+option in inkscape).  One consquence of this design choice is that
+radius of curvature is not matched exactly across a node; the path is
+not G2 smooth in CAD terminology.  However, if we can ensure that the
+radius of curvature changes only by a small amount, visual
+continuity---in the sense of absence of visible interruptions to a
+smoothly evolving strand---is preserved.  At a crossing point it is
+important that both the unbroken and broken strands (designated
+``overstrand'' and ``understrand'' respectively) are smooth.  We would
+like strand crossing points to be far from nodes, as this ensures
+visual continuity of strands at crossing points, especially the
+understrand.
 

### 2021-09-24 inst/jma_rebuttal.tex
fixes #22
fixes #25

@@
 
+\textcolor{blue}{%--------------------------------------------issue22
+  The path does not have G2 continuity because the radius of curvature
+  has a discontiniuity at Bezier nodes.  But the change of curvature
+  should be small, I have added a short discussion to clarify the
+  issue.}
+
 P3, line 33:  The ``understrand" has not been designated yet.
@@
 following easier to understand.
+
+\textcolor{blue}{Fixed, a short introductory passage and a reference
+  to R core now added}%issue#25
 

### 2021-09-24 inst/jma_knot.tex
fixes #25

@@
 
-In this section, I give workflows for creating two simple knots:
-firstly~$7_6$, followed by the figure-of-eight knot~$4_1$ which
-requires imposition of symmetry constraints.
+The software presented here is written in the ``R'' programming
+language~\cite{rcore2021}, which is usually associated with
+statistical analysis and data processing.  However, R possesses a
+number of features such as object-oriented programming and
+high-dimensional optimization which make it suitable for producing
+knot diagrams.  In this section, I give workflows for creating two
+simple knots: firstly~$7_6$, followed by the figure-of-eight
+knot~$4_1$ which requires imposition of symmetry constraints.
 

### 2021-09-26 inst/jma_knot.tex
fixes #39

@@
   Book of Kells} and the Lindisfarne gospels, graphic art produced
-around 800\,CE~\cite{bain1973}.  Such artwork makes extensive use of
-symmetry~\cite{cromwell2008}.
+around 800\,CE~\cite{bain1973}.  Such Celtic artwork makes extensive
+use of symmetry~\cite{cromwell2008}, and their visual appeal persists
+to the present day~\cite{antonsen2021}.  More recent examples would
+include bobbin lace~\cite{irvine2020}, in which many threads are
+braided together to form a loose fabric.  The enduring popularity of
+such fibre arts attests to the aesthetic of knots and braids in the
+modern world
 
@@
 up to~10 crossings---and Thistlethwaite~\cite{hoste1998}, for knots
-of~11 crossings.  Symbolic software such as Mathematica and Maple can
-be used to produce representations of knots, typically in the form of
-nonstandard embeddings of the torus into~$\mathcal{R}^3$; but as
-discussed below this approach is suboptimal from the perspective of producing a
+of~11 crossings.
+
+\subsection{Computer-generated knot visualisations}
+
+Symbolic software such as Mathematica and Maple can be used to produce
+representations of knots, typically in the form of nonstandard
+embeddings of the torus into~$\mathbb{R}^3$; but as discussed below
+this approach is suboptimal from the perspective of producing a
 visually appealing projection.
+
+Visual appeal of knot projections appears to be important to
+contemporary mathematicians.  Recent work by Taalman (writing as
+mathgrrl)~\cite{taalman2019}, for example, shows that large amounts of
+effort have been expended by serious mathematicians making
+representations of knots attractive.  Robert Bosch~\cite{bosch2010},
+taking a different approach to knot representations, presents artwork
+that utilises high-dimensional computer optimization in a similar way
+to that presented here.
+
+No discussion of computer-generated representations of knots would be
+complete without the tour de force that is~{\em
+  KnotPlot}~\cite{scharein1997,scharein1998} which includes extensive
+functionality for visualising knots.  Much of the visualization
+considers (nonstandard) embeddings of a torus in $\mathbb{R}^3$; the
+configuration is ``relaxed'' in such a way as to produce a smoother
+output.  {\em KnotPlot} has two major display modes: any knot may be
+displayed in either a ``beads and sticks" mode or in a ``smooth tubes"
+mode.  The software does render knot diagrams but the author states
+[p77] that ``knots with higher crossing numbers (greater than six or
+seven) will not produce good 2D projections when fully relaxed"; he
+goes on to say that this phenomenon merits further study, and admits
+that some of the diagrams are ``a bit messy".
+
+%One of the early works using this approach is the 1932 German classic
+%Anschauliche Geometrie by David Hilbert and S. Cohn-Vossen, available
+%in English translation as Geometry and the Imagination [HCV52].  This
+%book is remarkable for the quality of insight and also for its
+%incredible diagrams, mostly hand drawn and (obviously) predating the
+%computer era.
+
+\subsubsection{Knot theory in R}
+
+The R programming language~\cite{rcore2021}, while usually associated
+with statistical analysis and data processing, is also a suitable tool
+for producing visually attractive illustrations.  Generative art [that
+  is, use of an algorithm, with or without a random component, in a
+  creative process] is possible in R~\cite{brunner2021}.  More
+specficially to knot theory, the {\tt Rknots}
+package\cite{comoglio2011,comoglio2016} presents software for folded
+protein structures, and includes 3D renderings of commonly encountered
+knotted strands.
+
+\section{The knotR package: Rationale}
 
@@
 package).  Note that this file contains only the knot {\em path}; the
-over and under information is to be added later.
+over and under information is to be added later.  Irvine {\em
+  et. al.}~\cite{irvine2020} use the terms {\em drawing} and {\em
+  braid word} for analogous concepts in their work on bobbin lace
+patterns.
 
@@
 
-Many of the knots in Figure~\ref{knot_table} have an axis of symmetry,
-or possess rotational symmetry; some have both and thus have a
-dihedral symmetry group.  The package has the ability to impose these
-symmetry relations on knots, and to optimize the resulting symmetrical
-knot.  Minimizing the badness is not entirely straightforward on
-account of the induced redundancy, which is characterized using a
-symmetry object specific to the knot under consideration.  However,
-symmetry constraints do reduce the dimensionality of the optimization
-problem.
+If a knot diagram can be drawn with a particular symmetry, one common
+response is to demand that this symmetry be respected exactly.
+Bosch~\cite{bosch2010}, for example, considered rotational symmetry to
+be sufficiently important to impose five-fold rotatioal symmetry in
+one of his works; similar comments apply to mirror symmetry.  Many of
+the knots in Figure~\ref{knot_table} have an axis of symmetry, or
+possess rotational symmetry; some have both and thus have a dihedral
+symmetry group.  The package implements Bosch-type symmetry: one can
+impose these symmetry relations on knots, and optimize the resulting
+symmetrical knot.
+
+Minimizing the badness is not entirely straightforward on account of
+the induced redundancy, which is characterized using a symmetry object
+specific to the knot under consideration.  However, symmetry
+constraints do reduce the dimensionality of the optimization problem.
 

### 2021-09-26 inst/jma_rebuttal.tex
fixes #41

@@
   wide range of applications; and further that numerical optimization
-  will produce ``nice looking'' results.  Is is also reasonable to
-  assume that such techniques will have been used in the past by
-  graphical artists.  But I can find no evidence of such activity,
-  after an extensive online search and consulting with two librarians
-  (mathematics and art subject specialists).  Perhaps we are using the
-  wrong search terms?  I would gladly take suggestions on previous
-  work to cite.  I have added a sentence to the effect that this seems
-  to be new to the ``further work'' section.}
+  will produce ``nice looking'' results.  I have cited some of the
+  more prominent workers in the field, whose work is closest to 2D
+  knot projections.}
 
@@
 
+\textcolor{blue}{Brief discussion added}
+
 With these changes the paper could be an interesting one, discussing

### 2021-09-26 inst/jma_knot.tex
fixes #41

@@
 quantifies its displeasingness, and then minimize this objective
-function using numerical methods.
+function using numerical methods.  In the package, minimization is
+performed using either Nelder-Mead~\cite{nelder1965} or a Newton-type
+method~\cite{dennis1983} (gradient information is not available).  The
+optimization typically proceeds over $\mathbb{R}^{50}$ and in the
+context of the package, the methods appear to be broadly comparable.
 

### 2021-09-26 inst/jma_knot.tex
fixes #36

@@
 Also, several of the strands cross at acute angles.  The diagram for
-knot~$7_3$, for example, contains ugly kinks and abrupt changes of
-curvature.
+knot~$7_3$, for example, contains kinks and abrupt changes of
+curvature which distract from the underlying topological form.
 
@@
 suboptimal: even though the nodes are connected by Bezier curves which
-are individually smooth, the path as a whole is ugly and unattractive
-as its path has unsightly regions where the radius of curvature
-changes abruptly.
+are individually smooth, the path as a whole is visually disjointed as
+it has places where the radius of curvature changes abruptly.
 
@@
 function of~64 real variables, and the objective function takes a few
-seconds to evaluate).  However, the result is much nicer
-(Figure~\ref{7_6}).
+seconds to evaluate).  However, the result is visually smoother and
+thus arguably more attractive (Figure~\ref{7_6}).
 

### 2021-09-26 inst/jma_knot.tex
fixes #32

@@
 There now follows a selection of pleasing knot diagrams taken from
-datasets provided with the package.
+datasets provided with the package, illustrating the relative ease
+with which knot diagrams may be created and optimized.
+Figure~\ref{perko_AB} shows two forms of a 10-crossing knot once
+erroneously considered to be topologically distinct;
+figure~\ref{orn20} shows an ornamnental knot with exact five-fold
+symmetry (but not dihedral symmetry); and figure~\ref{all8} shows all
+prime knots with eight or fewer crossings, following
+Rolfsen~\cite{rolfsen1976}.  The package includes all knots to nine
+crossings.
 

### 2021-09-26 inst/jma_rebuttal.tex
fixes #36
fixes #32
fixes #28

@@
 Figure 4: Why is this required?  -- Does the user have to specify the
-type of crossing via an "overunder" object?  -- (Lines 53-62?)
+type of crossing via an "overunder" object?  -- (Lines 53-62?)%issue28
+
+\textcolor{blue}{The {\tt overunderobject} specifies which of two
+  strands at a crossing is the overstrand and which the understand.  I
+  have added a definition of this and a brief discussion in the figure
+  caption.}
 
@@
 
-\textcolor{blue}{Rolfsen now cited.}
+%issue32
+
+\textcolor{blue}{Rolfsen now cited, and a discussion of the table now
+  given to place it in context.}
 
@@
 pleasing vs ugly knots. Aesthetics is a complex field, but generally
-does not provide strong answers. While the choices made in this paper
+does not provide strong answers.
+
+\textcolor{blue}{Absolute terms now rephrased more carefully.}%issue#36
+
+While the choices made in this paper
 are reasonable, they are simply stated, rather than including a

### 2021-09-26 inst/jma_knot.tex
fixes #28

@@
 \includegraphics[scale = 0.9]{knot-004}
-\caption{Knot~$7_6$ with strands numbered \label{k76_strands} to guide
-creation of over and under information} \end{figure}
+\caption{Knot~$7_6$ with strands numbered \label{k76_strands} so that
+  the sense of the crossings can be established.  For example, strand
+  7 passes over strand 3.  In the package, this fact is specified by
+  the {\tt overunder} object, as discussed in the text, having a row
+  {\tt 7 3} (and not {\tt 3 7} which would be the crossing having the
+  opposite sense)} \end{figure}
 

### 2021-09-26 inst/jma_rebuttal.tex
fixes #35

@@
 describe the techniques used by the software to create the optimised
-images.
+images.%issue35
+
+\textcolor{blue}{I think this is a perfectly reasonable comment for
+  the original manuscript but the revision is a very much improved
+  document.}
 

### 2021-09-27 inst/jma_rebuttal.tex
fixes #16

@@
       topological invariance of a knot"
-      \end{description}
+  \end{description}%issue16
+  The referee observes that $4_1$ has a symmetrical rendering with two
+  mirror lines and this is available in the package, as object {\tt
+    k4\_1a}.  I have added a description of this to the manuscript.
 }
@@
 describe the techniques used by the software to create the optimised
-images.%issue35
+images.%issue35<
 

### 2021-09-27 inst/jma_knot.tex
fixes #16
typos

@@
 such fibre arts attests to the aesthetic of knots and braids in the
-modern world
+modern world.
 
@@
 Rolfsen's diagrams are used widely today: Adams~\cite{adams2004}
-reproduces his work extensively, as do many more recent works.  Online
+reproduces his work extensively, as do many more recent works.
+
+\subsection{Computer-generated knot visualisations}
+
+Online
 resources are becoming more widely available, with one prominent
@@
 of~11 crossings.
-
-\subsection{Computer-generated knot visualisations}
 
@@
 [p77] that ``knots with higher crossing numbers (greater than six or
-seven) will not produce good 2D projections when fully relaxed"; he
-goes on to say that this phenomenon merits further study, and admits
-that some of the diagrams are ``a bit messy".
+seven) will not produce good 2D projections when fully relaxed", and
+admits that some of the diagrams are ``a bit messy".
 
@@
   creative process] is possible in R~\cite{brunner2021}.  More
-specficially to knot theory, the {\tt Rknots}
+specificially to knot theory, the {\tt Rknots}
 package\cite{comoglio2011,comoglio2016} presents software for folded
@@
 specify the path of a knot projection; here cubic Bezier curves are
-used.
-
-The software discussed here~\cite{hankin2020} allows one to specify a
-knot in terms of its Bezier control points within inkscape, import the
-object into R~\cite{rcore2021}, and then to use numerical optimization
-techniques to improve the visual appearance of the knot.
+used.  The software discussed here~\cite{hankin2020} allows one to
+specify a knot in terms of its Bezier control points within inkscape,
+import the object into R~\cite{rcore2021}, and then to use numerical
+optimization techniques to improve the visual appearance of the knot.
 
@@
 present a short discussion in section~\ref{section_cca} below.
-Numeically, the badnesses are evaluated by function {\tt badness()}:
+Numerically, the badnesses are evaluated by function {\tt badness()}:
 
@@
 Bosch~\cite{bosch2010}, for example, considered rotational symmetry to
-be sufficiently important to impose five-fold rotatioal symmetry in
+be sufficiently important to impose five-fold rotational symmetry in
 one of his works; similar comments apply to mirror symmetry.  Many of
@@
 
-(Matrix {\tt Mver8} specifies that nodes~2 and~3 are symmetric, as
-are nodes~9 and~7, and so on; {\tt xver=8} forces node~8 to be on the
-axis of symmetry).  The results are shown in
-Figure~\ref{four_figure_8_knots}.
+(Matrix {\tt Mver8} specifies that nodes~2 and~3 are symmetric, as are
+nodes~9 and~7, and so on; {\tt xver=8} forces node~8 to be on the axis
+of symmetry).  The results are shown in
+Figure~\ref{four_figure_8_knots}.  Note that $4_1$ may be rendered in
+a form that has two mirror lines, and this is available in the package
+as object {\tt k4\_1a}.
 
@@
 a bland compromise, but it is easy to place greater weight on one or
-other compoent.  Figure~\ref{figure_cca} shows the same knot optimized
+other component.  Figure~\ref{figure_cca} shows the same knot optimized
 using different weightings; we see the effect of imposing right-angled
@@
 erroneously considered to be topologically distinct;
-figure~\ref{orn20} shows an ornamnental knot with exact five-fold
+figure~\ref{orn20} shows an ornamenntal knot with exact five-fold
 symmetry (but not dihedral symmetry); and figure~\ref{all8} shows all
@@
 
-\section*{References}
-
 \bibliography{knot}

### 2021-09-28 inst/jma_rebuttal.tex
rewording and typos

@@
   components.  Several problems stood out: firstly, the problem of
-  inter-component points.  Currently this is a (symmetrical) Boolean
-  matrix of strands with rows and columns corresponding to strands,
-  and entries corresponding to whether that strand pair intersects or
-  not.  The link generalization would be a symmetric matrix of
-  intersection matrices: an object with four indices and entry
-  (i,j,k,l) indicating whether strand j of component i intersects
-  strand k of component l.  I found this object to be unweildy to work
-  with and difficult to manipulate.  The second problem was the
-  imposition of symmetry.  Sometimes one might desire that some
-  components of a link have a particular kind of symmetry and others
-  to have a different kind (or no!) symmetry.  One unexpectedly
-  difficult problem was that one might have a subset of links that
-  individually possess no symmetry but collectively possess mirror-,
-  rotational-, or indeed dihedral- symmetry.  Thistlethwaite's L4a1 is
-  problematic: one would expect the two components to be identical
-  except for a 90-degree rotation, and implementing this in the
-  context of the package I found impossible. L6a3 is similarly
-  difficult.  We would need new badnesses too.  Consider L6a4; the
-  smaller component should be not only as circular as possible (?),
-  but also one might desire the overall length to be smaller than that
-  of the other component.\\ My overall view was that dealing with
-  links would be possible in principle but, due to a number of rather
-  poorly delineated reasons (which are not really suitable for
-  inclusion in a scholarly publication), an order of magnitude harder
-  than the single-component links considered in the submission.  I'm
-  not saying it's impossible, but just very difficult and, for me,
-  firmly in the category of ``further work''.  I hope that's OK.}
+  inter-component intersection points.  Currently this is a
+  (symmetrical) Boolean matrix with rows and columns corresponding to
+  strands, and entry $(i,j)$ being whether or not strand $i$
+  intersects with strand $j$.  The link generalization would be a
+  symmetric matrix of intersection matrices: an object with four
+  indices and entry $(i,j,k,l)$ indicating whether strand $j$ of
+  component $i$ intersects strand $k$ of component $l$.  I found this
+  object to be unwieldy to work with and difficult to manipulate.  The
+  second problem was the imposition of symmetry.  Sometimes one might
+  desire that some components of a link have a particular kind of
+  symmetry and others to have a different kind (or no!) symmetry.  One
+  unexpectedly difficult problem was that one might have a subset of
+  links that individually possess no symmetry but collectively possess
+  mirror-, rotational-, or indeed dihedral- symmetry.
+  Thistlethwaite's L4a1 is problematic: one would expect the two
+  components to be identical except for a 90-degree rotation, and I
+  was unable to implement this in the context of the package. L6a3 is
+  similarly difficult.  We would need new badnesses too.  Consider
+  L6a4; the smaller component should be not only as circular as
+  possible (?), but also one might desire the overall length to be
+  smaller than that of the other component.\\ My considered view is
+  that dealing with links would be possible in principle but, due to a
+  number of non-interesting technical reasons (which are not really
+  suitable for inclusion in a scholarly publication), an order of
+  magnitude harder than the single-component links considered in the
+  submission.  I'm not saying it's impossible, but just very difficult
+  and, for me, firmly in the category of ``further work''.  I hope
+  that's OK.}
 

### 2021-09-28 inst/jma_knot.tex
minor tweaks

@@
 over another.  Knot diagrams have a long history:
-Przytycki~\cite{przytycki1998} discusses an intricate braid used as a
-seal sign in around 1700\,BCE; they figure prominently in {\em The
-  Book of Kells} and the Lindisfarne gospels, graphic art produced
-around 800\,CE~\cite{bain1973}.  Such Celtic artwork makes extensive
-use of symmetry~\cite{cromwell2008}, and their visual appeal persists
-to the present day~\cite{antonsen2021}.  More recent examples would
-include bobbin lace~\cite{irvine2020}, in which many threads are
-braided together to form a loose fabric.  The enduring popularity of
-such fibre arts attests to the aesthetic of knots and braids in the
-modern world.
+Przytycki~\cite{przytycki1998} discusses an intricate braid drawn
+around 1700\,BC; they figure prominently in {\em The Book of Kells}
+and the Lindisfarne gospels, graphic art produced around
+800\,AD~\cite{bain1973}.  Such artwork makes extensive use of
+symmetry~\cite{cromwell2008}, and its visual appeal persists to the
+present day~\cite{antonsen2021}.  More recent examples would include
+bobbin lace~\cite{irvine2020}, in which many threads are braided
+together to form a loose fabric.  The enduring popularity of such
+fibre arts attests to the aesthetic of knots and braids in the modern
+world.
 
@@
 
-Online
-resources are becoming more widely available, with one prominent
-example being the {\em Knot Atlas}~\cite{knot_atlas} presenting
-extensive knot tables following Rolfsen~\cite{rolfsen1976}---for knots
-up to~10 crossings---and Thistlethwaite~\cite{hoste1998}, for knots
-of~11 crossings.
+Online resources are becoming more widely available, with one
+prominent example being the {\em Knot Atlas}~\cite{knot_atlas}
+presenting extensive knot tables following
+Rolfsen~\cite{rolfsen1976}---for knots up to~10 crossings---and
+Thistlethwaite~\cite{hoste1998}, for knots of~11 crossings.
 
@@
 this approach is suboptimal from the perspective of producing a
-visually appealing projection.
+visually pleasing projection.
 
@@
 effort have been expended by serious mathematicians making
-representations of knots attractive.  Robert Bosch~\cite{bosch2010},
-taking a different approach to knot representations, presents artwork
-that utilises high-dimensional computer optimization in a similar way
-to that presented here.
+representations of knots attractive.  Bosch~\cite{bosch2010}, taking a
+different approach to knot representations, presents artwork that
+utilises high-dimensional computer optimization, as here.
 
 No discussion of computer-generated representations of knots would be
-complete without the tour de force that is~{\em
+complete without mentioning the tour de force that is~{\em
   KnotPlot}~\cite{scharein1997,scharein1998} which includes extensive
@@
 with statistical analysis and data processing, is also a suitable tool
-for producing visually attractive illustrations.  Generative art [that
-  is, use of an algorithm, with or without a random component, in a
+for producing visually attractive images.  Generative art [that is,
+  use of an algorithm, with or without a random component, in a
   creative process] is possible in R~\cite{brunner2021}.  More

### 2021-09-28 inst/jma_rebuttal.tex
further to #31
minor tweaks

@@
 
+\textcolor{blue}{Some knots have dihedral symmetry but not this one.
+  Discussion added.}
+
 Fig. 11: Why is this table shown?  It is not mentioned in the main
@@
 \textcolor{blue}{I think this is a perfectly reasonable comment for
-  the original manuscript but the revision is a very much improved
-  document.}
+  the original manuscript.  I believe the revision is a very much
+  improved document.}
 
@@
   software engineer was to find the projection that optimizes these
-  standard weightings.\\ \\ Following this comment, I realise with a
-  shock that my weightings do not reflect some universal ideal.
-  Rather, my choice of weightings are as subjective as any other
-  aspect of the software and, as such, are peculiar to me.  This is
-  obvious in retrospect but was not clear to me at all until just
-  now.\\ \\ One of the benefits of the {\tt knotR} software is that it
-  allows this kind of conversation to occur; I have added a discussion
-  to the manuscript, just before the conclusions section.  There, I
-  present projections of $8_3$ and $8_5$ but with different penalties
+  standard weightings.\\ \\ Reflecting on the referee's comment, I
+  realise with a shock that my weightings do not reflect some
+  universal ideal.  Rather, my choice of weightings are as subjective
+  as any other aspect of the software and, as such, are peculiar to
+  me.  This is obvious in retrospect but was not clear to me at all
+  until just now.\\ \\ One of the benefits of the {\tt knotR} software
+  is that it allows this kind of conversation to occur; I have added a
+  discussion to the manuscript, just before the conclusions section.
+  There, I present projections of $8_3$ but with different penalties
   for angle crossing terms.  We see the results and can compare like
@@
 
+\textcolor{blue}{I have included a figure that compares a knot with
+  two different weightings}
+
 3) Give basic information about the specific methods being used for
@@
 
-\textcolor{blue}{Brief discussion added}
+\textcolor{blue}{Brief discussion added: the package includes
+  Nelder-Mead by default but the user has the option to use a
+  Newton-type method.}
 

### 2021-09-28 inst/jma_knot.tex
spelling check

@@
 Knot diagrams are ubiquitous in the modern discipline of mathematical
-knot theory.  In the nineteeth century we see Tait~\cite{tait1884}
+knot theory.  In the nineteenth century we see Tait~\cite{tait1884}
 publishing an early systematic study of knots including a beautiful
@@
   creative process] is possible in R~\cite{brunner2021}.  More
-specificially to knot theory, the {\tt Rknots}
+specifically to knot theory, the {\tt Rknots}
 package\cite{comoglio2011,comoglio2016} presents software for folded
@@
 symmetrically placed with regard to nodes (this is a user-settable
-option in inkscape).  One consquence of this design choice is that
+option in inkscape).  One consequence of this design choice is that
 radius of curvature is not matched exactly across a node; the path is
@@
 erroneously considered to be topologically distinct;
-figure~\ref{orn20} shows an ornamenntal knot with exact five-fold
+figure~\ref{orn20} shows an ornamental knot with exact five-fold
 symmetry (but not dihedral symmetry); and figure~\ref{all8} shows all

### 2022-06-01 inst/jma_rebuttal_review3.tex
work on issue #15

\documentclass[12pt]{article}
\usepackage{xcolor}
\begin{document}

(Below, the reviewer's comments are in black, and my replies to the
issues are in \textcolor{blue}{blue}.  I have indicated changes to the
manuscript where appropriate.

\section*{Rebuttal to second third review: overall response}

Short story: the reviewers make reasonable and constructivde comments.
I have accommodated the comments with  with rewording.  The resulting
document is, I believe, stronger and more scholarly than before and I
recommend it to you.

\section*{Detailed rebuttal: Reviewer \#1}

Comments from the Editors and Reviewers:

Thank you for this revised manuscript.  My apologies for contributing
to the slow turnaround of these reviews.

I have given this version the official status of ``Minor Revision" to
reflect the fact that we all agree that this draft is much closer to
publication-ready. In particular, I believe that the remaining
reviewer concerns can be addressed with moderate rewriting.  However,
some of the reviewer feedback falls closer to ``Major Revision," so I
wanted to acknowledge that a few of the remaining critiques are beyond
surface level.

As you will see in the detailed feedback below, one of the reviewers
still has substantive concerns about the overall framing of your
discussion and how directly it addresses the mathematical and
aesthetic choices that went into the software package.  I suspect that
you can address the concerns in the first few review paragraphs by
being more specific about the constraints of your implementation, the
level of flexibility you hoped to provide for the user, and the
alternate possibilities for future work, but since I don't know what
technical decisions went into your coding, I might be misjudging
exactly what is suitable.

I look forward to receiving your further revision, and will make sure
the editorial process goes more quickly from this point forward.

Best wishes,

Susan Goldstine, Ph.D.
Associate Editor
Journal of Mathematics and the Arts

Reviewer \#1: The paper has improved in terms of its overall
presentation in the context of knot renderings, and many detailed
fixes have been made as suggested by the reviewers.

However, the paper still has the major weakness of being too much a
somewhat ad-hoc technical manual on ``How-to-use-R and Inkplot to get
some decent knot plots" by using the software package [27].

The paper is lacking good discussions of the key issues and
explanations of the choices made, specifically: What makes a knot plot
attractive, followed by a thorough discussion of how to best set up a
knot representation and an optimization procedure to achieve the
desired qualities?

The problem starts with the basic curve representation:

Why use Bezier curves?

Why not use some other spline curve that guarantees G2-continuity?

And, if forced by the chosen tool to use Bezier curves, why not place
nodes at the crossing points, (and one or two more for large loopy
lobes between subsequent crossing points)?  -- This would allow the
user to place the crossing points far enough apart to start with, and
this would make it easy to set the Bezier handles to force right-angle
crossings.

Also, the implementation in the ``badness function" seems far from
optimal.  ``Total crossing angles()" does not address a key problem
directly: If one wants to avoid highly acute crossing angles, then
this can be addressed in a much more sensitive way with a function
such as 1/ sqr( cross.angle ).  This would go to infinity in the worst
case!  It also corresponds more directly to the way that excessive
curvature is avoided by minimizing ``bending energy," which is the
integral over the square of curvature -- or: 1/ sqr( curve-radius).
These are some ``math" issues that readers of JMA would be interested
in.

Some more detailed comments:

Page 3 line 42: ``...knots possess a line of symmetry (at least, the
diagrams do if the breaks are ignored), which ..."  -$>>$ ``...knots
possess a C2 rotation axis, which ..."

Page 4 lines 22-25: ``One plausible technique for creating knot
projections is to consider a two-dimensional projection of a knot's
embedding E in R3. However, this approach often results in poor visual
appearance: cusps or other displeasing effects can occur."  -$>>$ True!
- But what is the alternative, if the user has an entangled physical
knot in her hands?

Page 4 line 45: Define ``nodes" !

Page 4 line 48: ``However, if we can ensure that the radius of
curvature changes only by a small amount,..."  -$>>$ HOW can the user do
this?

Page 4 lines 50-52: No need to differentiate here between over- and
under-strands.  I assume they are treated the same in the optimization
process.

Page 4 lines 53-54: ``We would like strand crossing points to be far
from nodes,..."  -- Why?  The curves are also visually continuous across
the nodes!

Page 5 lines 46-51: Why are there SEVEN significant decimal digits?
Does the graphic screen have 10 million pixels per line?

Figure 3, 5, 7: Perhaps fewer than 100 circles per Bezier curve would
be better.

Figure 4.  Knot 76 with strands numbered... -- The main use here seems
to be to see very clearly the extent of every Bezier curve.
Over-/under-crossings seem to become relevant only for the final
rendering of the knot curve with the gaps included.

Page 7 line 52: ``The optimization typically proceeds over R 50" - What
is the meaning of this?

Page 7 line 56: ``...function total crossing angles()" The function is
not explicitly defined in the paper, but it sounds clearly suboptimal.
One bad acute-angle crossing can easily ```hide" in a sum of all angles.
It would be better to individually penalize acute angles.  A suitable
term might be: 1/ sqr( cross.angle ).  This would go to infinity in
the worst case!

Page 8 line 20, and footnote: The ``devil" and the ``interesting math"
is in the details!  The footnote is not sufficient to explain the
trade-offs that are being made in the optimization process.

Page 9, line 62: What is ``Bosch-type symmetry:..." ?  Please explain.
$>>$ If all nodes fulfill the specified symmetry constraints (as shown
on page 11, lines 15-22 and on lines 35-43), but the connection
diagram does not completely adhere to the specified symmetry - What
happens? - Does the package check this and let the user know about the
discrepancy?

Page 10 lines 56-58: ``Minimizing the badness is not entirely
straightforward..."  -- What is the problem?  I think that if all the
points are moved jointly within the specified symmetry constraints, an
optimized symmetrical solution will result.

{The above two comments suggest, that it might be advantageous to use
  some hierarchy in the representation of symmetrical knots: Only
  specify the curve for one unique fragment of the knot curve, and
  compose the complete curve by instancing that fragment with
  mirroring and rotation as required to construct the complete knot!
  The optimization would then only have to be done for this unique
  fragment (with proper end-conditions).  If this same hierarchy is
  also used in the graphical display of the knot curve, the user could
  easily make adjustments to just a few nodes, while always
  automatically obtaining the desired symmetrical result.}

Page 10 lines 55: ``In the package, the badness weightings may be
altered easily,..."  -- What is the user interface for this?  Are
there sliders in the display that allow interactive adjustments of the
weights?

Figure 9: The fact that ``angle-crossing penalty" had to be enhanced by
a factor of 100 to achieve a more desirable (in my opinion) result,
illustrates the fact that the "total crossing angles()"-function is
not a good choice for optimization.

I am missing a description how the two breaks at every crossing point
are being graphically implemented.

Conclusions:

RE: ``Further work might include functionality to deal
with links."  $>>$ Readers may think that this would be a trivial
extension.  Please add a paragraph like the one you wrote in response
to Reviewer 1, to explain why the extension to links may be more
difficult than one might think.

RE: ``In the broader context of optimization in art, we observe that
numerical optimization techniques can produce aesthetically pleasing
results, an observation that might find uptake by graphic artists."
$>>$ This only holds with the proviso that we can think of a good
mathematical formulations to capture the various features that capture
aesthetics appropriately.  - This then becomes the real ``art."

Gallery:

Knot diagrams 83 through 86 in Figure 12 stand out with their helices
too tightly wound.  The two criteria on page 4, lines 54 and 55 do not
seem to be sufficiently enforced.  I suggest adjusting these figures
in the spirit of Figure 9b.  This may happen automatically with a
better ``badness function" that penalizes non-orthogonal crossing
angles more directly.

[27]:  Incomplete reference.  Perhaps give a URL.

Reviewer \#2: Firstly I would like to apologise for my late response on
this. Reviewing this edited paper had not got onto my radar.

Secondly all the points in my original report have now been
addressed. I do feel that the literature review could have gone a
little bit further beyond my stated suggestions, but it is now
sufficient to give the paper more context.

\end{document}

### 2022-06-02 inst/jma_rebuttal_review3.tex
start work on #47 and #48

@@
 desired qualities?
+
+\textcolor{blue}{ I appreciate the sentiment, which I take to be a
+  comment on the large amount of programming-oriented detail in the
+  submission.  This reviewer evidently regards the knots in the
+  submission as ``decent'', which I take to be a great compliment.  It
+  is often difficult to know the correct balance between overly
+  technical descriptions [which indeed often overlaps with the aims of
+    a technical manual] and gushing conceptual big-picture thinking.
+  However, I note that many of the comments below are essentially
+  requests for clarifications of minor technical details of the
+  implementation: and this would indicate that perhaps the balance
+  should be shifted back toward the technical manual end of the
+  spectrum.  I want the editor and reviewer to know that in addressing
+  the concerns below I have tried hard to answer not just ``what'' but
+  also the ``why'' of the implementation.  \\ \\ Nevertheless, the cue
+  above is a good one.  The two questions: (a) {\em what makes a knot
+    plot attractive?}, and (b) {\em how best ... to achieve this?}
+  are indeed profound, and were not addressed directly in the old
+  manuscript.  In the revision I have prominently stated the questions
+  in the introduction, and given a provocative, if brief, outline of
+  my view.}
 

### 2022-06-02 inst/jma_rebuttal_review3.tex
more reflections on issue #47

@@
 
-\textcolor{blue}{ I appreciate the sentiment, which I take to be a
+\textcolor{blue}{I appreciate the sentiment, which I take to be a
   comment on the large amount of programming-oriented detail in the
@@
   should be shifted back toward the technical manual end of the
-  spectrum.  I want the editor and reviewer to know that in addressing
-  the concerns below I have tried hard to answer not just ``what'' but
+  spectrum.  The other observation I would make here is that the
+  software package [27] is intended to be part of the submission (it
+  is released under the GPL and is thus available for public use) and
+  with this understanding, even a downbeat technical manual would have
+  scholarly value in conjunction with the new software it describes.
+  I want the editor and reviewer to know that in addressing the
+  concerns below I have tried hard to answer not just ``what'' but
   also the ``why'' of the implementation.  \\ \\ Nevertheless, the cue

### 2022-06-06 inst/jma_rebuttal_review3.tex
fixes #58

@@
 
+\textcolor{blue}{I have added a figure of a trefoil knot that
+  illustrates the reason why nodes' being at crossing points is
+  undesirable.  In short, at the node of a Bezier curve, the radius of
+  curvature undergoes a visually jarring discontinuity which looks
+  bad, especially the understrand, which is particularly susecptible
+  to such considerations because of the break.}
+
 Page 5 lines 46-51: Why are there SEVEN significant decimal digits?

### 2022-06-06 inst/jma_knot.tex
fixes #58

@@
 visual continuity of strands at crossing points, especially the
-understrand.
+understrand (see Figure~\ref{trefoil_circles} for an example and
+discussion).
+
+\begin{figure}[h]
+  \centering
+    \includegraphics[width=10cm]{trefoil_circles_post_inkscape.pdf}
+\caption{A trefoil knot \label{trefoil_circles} constructed from arcs
+  of circles (elastica) constrained to cross at right angles; the
+  strands exert a couple but no force on one another.  Mirror symmetry
+  and threefold rotational symmetry follow.  Elementary trigonometry
+  shows that the radii of the inner and outer loops are
+  $\operatorname{cosec}(\pi/12)=\sqrt{6}+\sqrt{2}$ and
+  $\sec(\pi/12)=\sqrt{6}-\sqrt{2}$ respectively.  The curvature thus
+  has a displeasing discontinuity at a node which interrupts the
+  visual continuity there, especially the understrand}
+\end{figure}
 

### 2022-06-07 inst/jma_rebuttal_review3.tex
more work

--- 2022-06-02 inst/jma_rebuttal_review3.tex
@@
 
+\textcolor{blue}{The curves certainly are visually continuous across
+  the nodes---but only in the optimized knots.  Figure 4, for example,
+  illustrates an unoptimized knot and we see that the Bezier nodes can
+  exhibit ugly features.  The motivation for keeping Bezier nodes far from crossing points 
+I have added a
+  figure of a trefoil knot that illustrates the reason why nodes'
+  being at crossing points is undesirable.  In short, at the node of a
+  Bezier curve, the radius of curvature undergoes a visually jarring
+  discontinuity which looks bad, especially the understrand, which is
+  particularly susecptible to such considerations because of the
+  break.\\ \\ Also (refer to figure 5, which shows 7_6 with numbered
+  strands), we can see that if a Bezier
+
+}
+
 Page 5 lines 46-51: Why are there SEVEN significant decimal digits?

### 2022-06-07 inst/jma_knot.tex
more work
further work on #58

@@
   strands exert a couple but no force on one another.  Mirror symmetry
-  and threefold rotational symmetry follow.  Elementary trigonometry
-  shows that the radii of the inner and outer loops are
+  and threefold rotational symmetry are imposed.  Elementary
+  trigonometry shows that the radii of the inner and outer loops are
   $\operatorname{cosec}(\pi/12)=\sqrt{6}+\sqrt{2}$ and
@@
   visual continuity there, especially the understrand}
+\end{figure}
+
+\begin{figure}[h]
+  \centering  
+  % To create file "two_trefoil_knots_post_inkscape.pdf", run
+  % knotter.R, which makes (among other things) two_trefoil_knots.pdf.
+  % Edit this in inkscape [make a mirror image, then change the
+  % colours and finally use the raise and lower tools to put
+  % everything in the right order]
+    \includegraphics[width=10cm]{two_trefoil_knots_post_inkscape.pdf}
+    \caption{Two trefoil knots \label{two_trefoil_knots}.  Left,
+      optimized as per the {\tt knotR} package techniques documented
+      here; right, constructed from arcs of circles constrained to
+      cross at right angles.  The strands exert a couple but no force
+      on one another.  Mirror symmetry and threefold rotational
+      symmetry are imposed.  Elementary trigonometry shows that the
+      radii of the inner and outer loops are
+      $\operatorname{cosec}(\pi/12)=\sqrt{6}+\sqrt{2}$ and
+      $\sec(\pi/12)=\sqrt{6}-\sqrt{2}$ respectively.  The curvature
+      thus has a displeasing discontinuity at a node which interrupts
+      the visual continuity there, especially the understrand}
 \end{figure}

### 2022-06-07 inst/jma_rebuttal_review3.tex
fixes #58

--- 2022-06-02 inst/jma_rebuttal_review3.tex
@@
 
+\textcolor{blue}{The curves certainly are visually continuous across
+  the nodes---but only in the optimized knots.  The motivation for
+  keeping Bezier nodes far from crossing points is twofold.  Firstly,
+  as one traces a curve visually, Bezier nodes cannot possibly avoid
+  introducing {\em some} form of disjointness in the path; and such
+  disjointness is particularly disruptive if on the understrand of a
+  node, as the visual continuity is interrupted.  I have added a
+  figure (including two different diagrams of a trefoil knot) to the
+  Gallery that illustrates the reason why nodes' being at crossing
+  points is undesirable.  In short, at the node of a Bezier curve, the
+  radius of curvature undergoes a visually jarring discontinuity which
+  looks bad, especially the understrand, which is particularly
+  susecptible to such considerations because of the break.  I do
+  appreciate that many viewers will find the distinction invisible or
+  inconsequential; but my informal polling of family and colleagues
+  reveals that, once sensitised to the issue, they cannot unsee the
+  kink.  \\ \\ The second reason is more mundane.  With reference to
+  figure 7 (the four figure-of-eight knots diagram), we can see a
+  vertical line of symmetry on which there are two crossing points.
+  Imagine there was a node at one of these crossing points.  There
+  would have to be {\em two} nodes at the same point, one for the
+  overstrand and one for the understrand; the software would somehow
+  have to impose vertical mirror symmetry on the diagram and this
+  would necessitate that the knot class contain yet another symmetry
+  object, alongside the 12 existing ones.  These considerations become
+  yet more onerous when considering rotational symmetry as in Figure 8
+  (which shows $5_1$).  If there were a node at, say, the intersection
+  of strand 18-19 and 9-10 then somehow the two nodes would have to
+  embody not only mirror symmetry (with respect to 5-6/14-15) but also
+  fivefold rotational symmetry.  Simply put, this is a nightmare.  The
+  software documentation contains a substantial discussion of the
+  issue, but I do not believe that the manuscript would benefit from
+  what amounts to little more than me grumbling about implementation
+  difficulties.}
+
 Page 5 lines 46-51: Why are there SEVEN significant decimal digits?

### 2022-06-07 inst/jma_knot.tex
fixes #58

@@
 visual continuity of strands at crossing points, especially the
-understrand (see Figure~\ref{trefoil_circles} for an example and
+understrand (see Figure~\ref{two_trefoil_knots} for an example and
 discussion).
-
-\begin{figure}[h]
-  \centering
-    \includegraphics[width=10cm]{trefoil_circles_post_inkscape.pdf}
-\caption{A trefoil knot \label{trefoil_circles} constructed from arcs
-  of circles (elastica) constrained to cross at right angles; the
-  strands exert a couple but no force on one another.  Mirror symmetry
-  and threefold rotational symmetry are imposed.  Elementary
-  trigonometry shows that the radii of the inner and outer loops are
-  $\operatorname{cosec}(\pi/12)=\sqrt{6}+\sqrt{2}$ and
-  $\sec(\pi/12)=\sqrt{6}-\sqrt{2}$ respectively.  The curvature thus
-  has a displeasing discontinuity at a node which interrupts the
-  visual continuity there, especially the understrand}
-\end{figure}
-
-\begin{figure}[h]
-  \centering  
-  % To create file "two_trefoil_knots_post_inkscape.pdf", run
-  % knotter.R, which makes (among other things) two_trefoil_knots.pdf.
-  % Edit this in inkscape [make a mirror image, then change the
-  % colours and finally use the raise and lower tools to put
-  % everything in the right order]
-    \includegraphics[width=10cm]{two_trefoil_knots_post_inkscape.pdf}
-    \caption{Two trefoil knots \label{two_trefoil_knots}.  Left,
-      optimized as per the {\tt knotR} package techniques documented
-      here; right, constructed from arcs of circles constrained to
-      cross at right angles.  The strands exert a couple but no force
-      on one another.  Mirror symmetry and threefold rotational
-      symmetry are imposed.  Elementary trigonometry shows that the
-      radii of the inner and outer loops are
-      $\operatorname{cosec}(\pi/12)=\sqrt{6}+\sqrt{2}$ and
-      $\sec(\pi/12)=\sqrt{6}-\sqrt{2}$ respectively.  The curvature
-      thus has a displeasing discontinuity at a node which interrupts
-      the visual continuity there, especially the understrand}
-\end{figure}
 
@@
 with which knot diagrams may be created and optimized.
-Figure~\ref{perko_AB} shows two forms of a 10-crossing knot once
-erroneously considered to be topologically distinct;
-figure~\ref{orn20} shows an ornamental knot with exact five-fold
-symmetry (but not dihedral symmetry); and figure~\ref{all8} shows all
-prime knots with eight or fewer crossings, following
-Rolfsen~\cite{rolfsen1976}.  The package includes all knots to nine
-crossings.
+Figure~\ref{two_trefoil_knots} shows and compares two plausible
+schemes for rendering a trefoil.  Figure~\ref{perko_AB} shows two
+forms of a 10-crossing knot once erroneously considered to be
+topologically distinct; figure~\ref{orn20} shows an ornamental knot
+with exact five-fold symmetry (but not dihedral symmetry); and
+figure~\ref{all8} shows all prime knots with eight or fewer crossings,
+following Rolfsen~\cite{rolfsen1976}.  The package includes all knots
+to nine crossings.
+
+\begin{figure}[h]
+  \centering  
+  % To create file "two_trefoil_knots_post_inkscape.pdf", run
+  % knotter.R, which makes (among other things) two_trefoil_knots.pdf.
+  % Edit this in inkscape [make a mirror image, then change the
+  % colours and finally use the raise and lower tools to put
+  % everything in the right order]
+    \includegraphics[width=10cm]{two_trefoil_knots_post_inkscape.pdf}
+    \caption{Two trefoil \label{two_trefoil_knots} knots.  Left,
+      optimized as per the {\tt knotR} package techniques documented
+      here; right, constructed from arcs of circles constrained to
+      cross at right angles.  The strands exert a couple but no force
+      on one another.  Mirror symmetry and threefold rotational
+      symmetry are imposed.  Elementary trigonometry shows that the
+      radii of the inner and outer loops are
+      $\operatorname{cosec}(\pi/12)=\sqrt{6}+\sqrt{2}$ and
+      $\sec(\pi/12)=\sqrt{6}-\sqrt{2}$ respectively.  The curvature
+      thus has a displeasing discontinuity at a node which interrupts
+      the visual continuity there, especially the understrand}
+\end{figure}
 

### 2022-06-07 inst/jma_rebuttal_review3.tex
fixes #59

@@
 
+\textcolor{blue}{This is the default print method used by R (which
+  internally uses IEC-60559 arithmetic, 14 significant figures).  Any
+  attempt by me to change the print method would, I believe, be
+  confusing and unnecessary.  The point of the table was to show that
+  a knot path could be represented by a few dozen floating-point
+  numbers.}
+
 Figure 3, 5, 7: Perhaps fewer than 100 circles per Bezier curve would

### 2022-06-07 inst/jma_rebuttal_review3.tex
fixes #57
fixes #74
fixes #65

@@
 
+\textcolor{blue}{The referee is correct to say that the over and under
+  strand are treated the same by the optimization routine, but my
+  intent was to introduce and define the words themselves (rather than
+  to differentiate between them).}
+
 Page 4 lines 53-54: ``We would like strand crossing points to be far
@@
 Page 9, line 62: What is ``Bosch-type symmetry:..." ?  Please explain.
+
+\textcolor{blue}{Bosch-type symmetry refers to Bosch (2010) which was
+  cited in the previous sentence.\\ \\ old text: The package
+  implements Bosch-type symmetry\ldots \\ \\ new text: The package
+  implements symmetry in a similar way to that of Bosch (2010)\ldots}
+
 $>>$ If all nodes fulfill the specified symmetry constraints (as shown
@@
 
+\textcolor{blue}{Canonical URL now added}
+
 Reviewer \#2: Firstly I would like to apologise for my late response on

### 2022-06-07 inst/jma_knot.tex
fixes #65

@@
 possess rotational symmetry; some have both and thus have a dihedral
-symmetry group.  The package implements Bosch-type symmetry: one can
-impose these symmetry relations on knots, and optimize the resulting
-symmetrical knot.
+symmetry group.  The package implements symmetry in a similar way to
+that of Bosch~\cite{bosch2010}: one can impose these symmetry
+relations on knots, and optimize the resulting symmetrical knot.
 

### 2022-06-08 inst/jma_rebuttal_review3.tex
fixes #55
fixes #71

@@
 Page 4 line 45: Define ``nodes" !
+
+\textcolor{blue}{``Nodes'' is a standard word in the context of Bezier
+  curves. I have not used the word in any other sense in the
+  manuscript and I think the text is clear as it stands.}
 
@@
 
+\textcolor{blue}{Done, new subsection ``some reflections on links''
+  at the end, under ``conclusions and further work''}
+
 RE: ``In the broader context of optimization in art, we observe that

### 2022-06-08 inst/jma_knot.tex
fixes #71

@@
 such diagrams in terms of a customizable objective function using
-numerical optimization techniques.  Further work might include
-functionality to deal with links.  In the broader context of
-optimisation in art, we observe that numerical optimisation techniques
-can produce aesthetically pleasing results, an observation that might
-find uptake by graphic artists.
+numerical optimization techniques.
+
+In the broader context of optimisation in art, we observe that
+numerical optimisation techniques can produce aesthetically pleasing
+results, an observation that might find uptake by graphic artists.
+
+\subsection{Some reflections on links}
+
+Further work might include functionality to deal with links.  Much of
+the required functionality is (in principle) already present in the
+software.  Each component would have its own badness, and in addition
+there would be $n(n-1)/2$ inter-component interaction terms measuring
+features such as closeness between non-intersecting strands of
+different components.  However, the implementation of such
+functionality is not straightforward and several problems stand out.
+Firstly, dealing with inter-component intersection points is
+difficult.  Currently, crossing points are represented as a
+(symmetrical) Boolean matrix with rows and columns corresponding to
+strands, and entry $(i,j)$ being whether strand $i$ intersects with
+strand $j$.  The link generalization would be a symmetric matrix with
+entries that are themselves intersection matrices: an object with four
+indices and entry $(i,j,k,l)$ indicating whether strand $j$ of
+component $i$ intersects strand $l$ of component $k$.  Such objects
+are unwieldy to work with and are typically difficult to manipulate.
+Secondly, the imposition of symmetry is not straightforward in the
+case of links: one might desire that certain components of a link have
+a particular kind of symmetry and others to have a different kind (or
+no!) symmetry.  Further, one unexpectedly difficult problem was
+dealing with a subset of components that individually possess no
+symmetry but collectively possess mirror (or rotational) symmetry.
+Even the simplest links, as in Figure~\ref{thistlethwaite}, can be
+problematic.  Consider Thistlethwaite's L4a1, for example: one would
+expect the two components to be identical except for a 90-degree
+rotation, and implementing this in the context of the package does not
+seem to be at all easy.  In general, new badnesses would be needed.
+Further, in L7a1, the smaller component should be not only as circular
+as possible (?), but also one might desire the overall length to be
+smaller than that of the other component.  It would be reasonable to
+conclude that dealing with links might be possible in principle but,
+due to a number of technical and mathematical reasons, considerably
+harder than the single-component links considered here.
+
+\begin{figure}[htbp]
+  \begin{center}
+\includegraphics[width=7cm]{L4a1_L7a1}  % made by exporting L4a1_L7a1.svg as PDF
+\caption{Two links \label{thistlethwaite} taken from the
+  Thistlethwaite link table~\cite{hoste1998}.  These are produced
+  directly with inkscape and are not optimized}
+  \end{center}
+\end{figure}
 

### 2022-06-08 inst/jma_rebuttal_review3.tex
fixes #49

@@
 Why use Bezier curves?
+
+\textcolor{blue}{Most of the answer is in the manuscript: ``A Bezier
+  curve is a visually pleasing polynomial path that can be used to
+  specify the path of a knot projection; here cubic Bezier curves are
+  used''.  However, Bezier curves are a natural choice for a number of
+  reasons and I have added the following sentence: ``Bezier curves are
+  a natural choice for working with mathematical knots for several
+  reasons: they are familiar to many graphical workers; they have a
+  natural and intuitive control system (usually called ``handles''),
+  and are implemented in many graphical software suites''.  I do not
+  wish to suggest that Bezier curves are the only parametrizations for
+  working with knots; but they are a natural starting point, and
+  AFAICS they have no serious disadvantages in this context, nor is
+  there any obvious superior method.}
 

### 2022-06-08 inst/jma_rebuttal_review3.tex
further work on #49

@@
 Why not use some other spline curve that guarantees G2-continuity?
+
+\textcolor{blue}{I don't have a good answer to that, other than to
+  point out that the focus of the manuscript was to investigate the
+  visual appearance of knots using mathematical optimization routines.
+  I do not wish to rule out using other G2-continuous splines, and it
+  might be interesting some day to consider (e.g.) B-splines.  But as
+  I say above, Bezier curves are a good starting point.}
 

### 2022-06-08 inst/jma_rebuttal_review3.tex
fixes #54

@@
 knot in her hands?
+
+\textcolor{blue}{Being able to render a nice optimized diagram
+  corresponding to a physical knot that one has in one's hands is a
+  reasonable wish.  But not one that I can fulfil.  Indeed, the
+  package is predicated upon the notion that the diagram is the
+  ``real'' object, and the (3D) knot itself is irrelevant.  Thus much
+  of the package {\em prevents} the optimization routine from making
+  Reidermeister moves [that is, changes that affect the topology of
+    the diagram but not that of the 3D knot].  The operational reason
+  for this is that the user specifies a particular diagram which the
+  software then optimizes for appearance.  The manuscript includes two
+  diagrams of $10_{125}$, the Perko pair, which show that sometimes
+  one wants to preserve the topology of the diagram.}
 

### 2022-06-08 inst/jma_rebuttal_review3.tex
fixes #56

@@
 
+\textcolor{blue}{It is not the user that stops large changes, it is
+  the optimization routine and in particular the rather eclectic set
+  of badness functions.  Incidentally, the ``keep nodes far from
+  crossing points'' badness function is instrumental in stopping large
+  changes of this type from occurring.\\ \\ OLD TEXT: However, if we
+  can ensure that the radius of curvature changes only by a small
+  amount, visual continuity---in the sense of absence of visible
+  interruptions to a smoothly evolving strand---is
+  preserved. \\ \\ NEW TEXT: However, because the optimization routine
+  ensures that the radius of curvature changes only by a small
+  amount, visual continuity---in the sense of absence of visible
+  interruptions to a smoothly evolving strand---is
+  preserved.\\ \\
+}
+
 Page 4 lines 50-52: No need to differentiate here between over- and

### 2022-06-08 inst/jma_knot.tex
fixes #49
fixes #56

@@
 specify the path of a knot projection; here cubic Bezier curves are
-used.  The software discussed here~\cite{hankin2020} allows one to
-specify a knot in terms of its Bezier control points within inkscape,
-import the object into R~\cite{rcore2021}, and then to use numerical
-optimization techniques to improve the visual appearance of the knot.
+used.  Bezier curves are a natural choice for working with
+mathematical knots for several reasons: they are familiar to many
+graphical workers; they have a natural and intuitive control system
+(usually called ``handles''), and are implemented in many graphical
+software suites.  The software discussed here~\cite{hankin2020} allows
+one to specify a knot in terms of its Bezier control points within
+inkscape, import the object into R~\cite{rcore2021}, and then to use
+numerical optimization techniques to improve the visual appearance of
+the knot.
 
@@
 radius of curvature is not matched exactly across a node; the path is
-not G2 smooth in CAD terminology.  However, if we can ensure that the
-radius of curvature changes only by a small amount, visual
-continuity---in the sense of absence of visible interruptions to a
-smoothly evolving strand---is preserved.  At a crossing point it is
-important that both the unbroken and broken strands (designated
-``overstrand'' and ``understrand'' respectively) are smooth.  We would
-like strand crossing points to be far from nodes, as this ensures
-visual continuity of strands at crossing points, especially the
-understrand (see Figure~\ref{two_trefoil_knots} for an example and
-discussion).
+not G2 smooth in CAD terminology.  However, because the optimization
+routine ensures that the radius of curvature changes only by a small
+amount, visual continuity---in the sense of absence of visible
+interruptions to a smoothly evolving strand---is preserved.  At a
+crossing point it is important that both the unbroken and broken
+strands (designated ``overstrand'' and ``understrand'' respectively)
+are smooth.  We would like strand crossing points to be far from
+nodes, as this ensures visual continuity of strands at crossing
+points, especially the understrand (see Figure~\ref{two_trefoil_knots}
+for an example and discussion).
 

### 2022-06-08 inst/jma_rebuttal_review3.tex
fixes #52
fixes #66

@@
 in.
+
+\textcolor{blue}{Function {\tt total\_crossing\_angles()} was poorly
+  named.  With crossing angles $\theta_i$, $i=1,2,\ldots k$ the
+  function actually returns $\sum_{i=1}^k\left(\cos\theta_i\right)^2$.
+  \\ \\}
 
@@
 
+\textcolor{blue}{The problem is reconstructing the knot from the
+  minimal set of of nodes and the various types of imposed symmetry.
+  This is not difficult exactly, but not entirely straightforward
+  either.\\ \\ Also, optimizing high-dimensional functions is rarely
+  straightforward: the search space is large; in my field of
+  computational ecology we talk of the ``curse of dimensionality''.
+  One has to contend with a wide variety of frustrating issues such as
+  local extrema, expensive function evaluation, and not least the
+  inadequacies of Euclidean distance as a metric for the domain.
+  Dimensionally cursed problems such as that addressed by {\tt knotR}
+  are sometimes literally intractable.  My considered view is that
+  Nelder-Mead simplex technique used in the package is unreasonably
+  effective.\\ \\ }
+
 {The above two comments suggest, that it might be advantageous to use

### 2022-06-08 inst/jma_rebuttal_review3.tex
fixes #67
fixes #63
fixes #72

@@
 
+\textcolor{blue}{Function {\tt total\_crossing\_angles()} was poorly
+  named.  With crossing angles $\theta_i$, $i=1,2,\ldots k$ the
+  function actually returns $\sum_{i=1}^k\left(\cos\theta_i\right)^2$.
+  \\ \\}
+
 Page 8 line 20, and footnote: The ``devil" and the ``interesting math"
@@
 
+\textcolor{blue}{This is exactly what I have done!  The referee gives
+  an accurate summary of the code's methodology.\\ \\}
+
 Page 10 lines 55: ``In the package, the badness weightings may be
@@
 
+\textcolor{blue}{ An insightful observation.\\ \\ OLD TEXT In the
+  broader context of optimisation in art, we observe that numerical
+  optimisation techniques can produce aesthetically pleasing results,
+  an observation that might find uptake by graphic artists.\\ \\ NEW
+  TEXT In the broader context of optimisation in art, we observe that
+  numerical optimisation techniques can produce aesthetically pleasing
+  results, an observation that might find uptake by graphic artists.
+  Indeed, one could argue that real art lies in the capturing of
+  mathematical formulations of aesthetics.  \\ \\ }
+
 Gallery:

### 2022-06-09 inst/jma_rebuttal_review3.tex
use standard knot names with subscripts
fixes #73

@@
 
-Knot diagrams 83 through 86 in Figure 12 stand out with their helices
+Knot diagrams $8_3$ through $8_6$ in Figure 12 stand out with their helices
 too tightly wound.  The two criteria on page 4, lines 54 and 55 do not
@@
 angles more directly.
+
+\textcolor{blue}{The referee is correct to say that changing the
+  badness function would result in figures less like 9a and more like
+  9b.  But I would take issue with the phrase ``too tightly wound''.
+  I have polled my colleauges who appear to be evenly split between
+  preferring 9a and 9b.  The comparison appears to elicit strong
+  opinions here at AUT.  Proponents of 9b point to the broader winding
+  along the central helix, while those of 9a point to the more liquid
+  and languid appearance, and in particular emphasise the absence of
+  retrograde curvature at the bottom.  I have added a brief discussion
+  of these issues to the caption of figure 12 [which is now figure 14
+    of the resubmission].  I would observe that the {\tt knotR}
+  package makes such discussions possible.\\ \\}
 

### 2022-06-09 inst/jma_knot.tex
fixes #72
fixes #73

@@
 results, an observation that might find uptake by graphic artists.
+Indeed, one could argue that real art lies in the capturing of
+mathematical formulations of aesthetics.
 
@@
 \caption{All \label{all8} prime knots with 8 or fewer crossings,
-notation following Rolfsen~\cite{rolfsen1976}}
+  notation following Rolfsen~\cite{rolfsen1976}.  Here, the penalty
+  for crossing angles departing from $90^\circ$ is relatively light
+  and so knot $8_3$, for example, matches the left diagram of
+  Figure~\ref{figure_cca}, rather than the right}
   \end{center}

### 2022-06-09 inst/jma_rebuttal_review3.tex
fixes #68
fixes #69

@@
 
+\textcolor{blue}{I have not implemented sliders.  The weightings may
+  be changed in function {\tt badness()}, and documentation (including
+  examples) is given at {\tt badness.Rd}.\\ \\}
+
 Figure 9: The fact that ``angle-crossing penalty" had to be enhanced by
@@
 not a good choice for optimization.
+
+\textcolor{blue}{As discussed above, {\tt total\_crossing\_angles()}
+  was a somewhat misleading name.  The factor of 100 ensured that even
+  tiny deviations from exactly $90^\circ$ were heavily penalised.}
 

### 2022-06-09 inst/jma_rebuttal_review3.tex
fixes #51

@@
 crossings.
+
+\textcolor{blue}{Most of this issue is discussed below, at the comment
+  for page 4, lines 53-54.  But here I would observe that it is not at
+  all easy to manipulate Bezier handles on two distinct nodes [which
+    happen to coincide spatially] to result in their two respective
+  lines being orthogonal.  Observe that the overstrand and understrand
+  of a crossing point may have (for good reasons!) very different
+  curvatures.}
 

### 2022-06-09 inst/jma_knot.tex
fixes #61

@@
   {\tt 7 3} (and not {\tt 3 7} which would be the crossing having the
-  opposite sense)} \end{figure}
+  opposite sense).  Crossings become relevant only for the final
+  rendering of the knot curve with the gaps included}
+\end{figure}
 

### 2022-06-09 inst/jma_rebuttal_review3.tex
fixes #61
fixes #62
fixes #70

@@
   kink.  \\ \\ The second reason is more mundane.  With reference to
-  figure 7 (the four figure-of-eight knots diagram), we can see a
+  figure 8 (the four figure-of-eight knots diagram), we can see a
   vertical line of symmetry on which there are two crossing points.
@@
   object, alongside the 12 existing ones.  These considerations become
-  yet more onerous when considering rotational symmetry as in Figure 8
+  yet more onerous when considering rotational symmetry as in Figure 9
   (which shows $5_1$).  If there were a node at, say, the intersection
@@
 
-Figure 4.  Knot 76 with strands numbered... -- The main use here seems
+Figure 4.  Knot $7_6$ with strands numbered... -- The main use here seems
 to be to see very clearly the extent of every Bezier curve.
@@
 
+\textcolor{blue}{Good point, I have added a sentence to this effect to
+  the caption of figure 4.\\ \\}
+
 Page 7 line 52: ``The optimization typically proceeds over R 50" - What
 is the meaning of this?
+
+\textcolor{blue}{It is a 50-dimensional optimization, optimizing a
+  function with fifty parameters (which correspond to different
+  positions of Bezier nodes and handles).\\ \\}
 
@@
 
+\textcolor{blue}{New figure added (Figure 6) which illustrates the implementation}
+
 Conclusions:
@@
   retrograde curvature at the bottom.  I have added a brief discussion
-  of these issues to the caption of figure 12 [which is now figure 14
+  of these issues to the caption of figure 12 [which is now figure 15
     of the resubmission].  I would observe that the {\tt knotR}

### 2022-06-09 inst/jma_knot.tex
fixes #70

@@
 the second the understrand; thus strand~12 passes over strand~1,
-strand~2 passes over strand~11, and so on. The result is shown in
-figure~\ref{76_overunder}.
+strand~2 passes over strand~11, and so on.  The graphical
+implementation is indicated in
+Figure~\ref{overundergraphicalimplementation}, and the complete knot
+is shown in figure~\ref{76_overunder}.
+
+\begin{figure}[!tbp] % file overundergraphicalimplementation.pdf is
+                     % created in file inst/overunder.R
+  \centering
+\includegraphics[width=4in]{overundergraphicalimplementation}
+\caption{Diagram \label{overundergraphicalimplementation} showing the
+  graphical implementation of the line breaks distinguishing
+  overstrand from understrand at a crossing point.  Top, details of
+  technique used, with mask coloured red.  Bottom, same diagram but
+  with mask coloured white, illustrating mechanism for production of
+  publication-quality knots}
+\end{figure}
 

### 2022-06-09 inst/jma_rebuttal_review3.tex
typos

@@
 
-Short story: the reviewers make reasonable and constructivde comments.
+Short story: the reviewers make reasonable and constructive comments.
 I have accommodated the comments with  with rewording.  The resulting
@@
   looks bad, especially the understrand, which is particularly
-  susecptible to such considerations because of the break.  I do
+  susceptible to such considerations because of the break.  I do
   appreciate that many viewers will find the distinction invisible or
@@
   9b.  But I would take issue with the phrase ``too tightly wound''.
-  I have polled my colleauges who appear to be evenly split between
+  I have polled my colleagues who appear to be evenly split between
   preferring 9a and 9b.  The comparison appears to elicit strong

### 2022-06-09 inst/jma_rebuttal_review3.tex
fixes #65
newlines to mark end of my blue comments

@@
   in the introduction, and given a provocative, if brief, outline of
-  my view.}
+  my view.\\ \\}
 
@@
   AFAICS they have no serious disadvantages in this context, nor is
-  there any obvious superior method.}
+  there any obvious superior method.\\ \\}
 
@@
   might be interesting some day to consider (e.g.) B-splines.  But as
-  I say above, Bezier curves are a good starting point.}
+  I say above, Bezier curves are a good starting point.\\ \\}
 
@@
   of a crossing point may have (for good reasons!) very different
-  curvatures.}
+  curvatures.\\ \\}
 
@@
   diagrams of $10_{125}$, the Perko pair, which show that sometimes
-  one wants to preserve the topology of the diagram.}
+  one wants to preserve the topology of the diagram.\\ \\}
 
@@
   curves. I have not used the word in any other sense in the
-  manuscript and I think the text is clear as it stands.}
+  manuscript and I think the text is clear as it stands.\\ \\}
 
@@
   intent was to introduce and define the words themselves (rather than
-  to differentiate between them).}
+  to differentiate between them).\\ \\}
 
@@
   what amounts to little more than me grumbling about implementation
-  difficulties.}
+  difficulties.\\ \\}
 
@@
   a knot path could be represented by a few dozen floating-point
-  numbers.}
+  numbers.\\ \\}
 
@@
   implements Bosch-type symmetry\ldots \\ \\ new text: The package
-  implements symmetry in a similar way to that of Bosch (2010)\ldots}
+  implements symmetry in a similar way to that of Bosch
+  (2010)\ldots\\ \\}
 
@@
 discrepancy?
+
+\textcolor{blue}{This situation cannot occur: symmetric nodes imply a
+  symmetric diagram.\\ \\}
 
@@
   was a somewhat misleading name.  The factor of 100 ensured that even
-  tiny deviations from exactly $90^\circ$ were heavily penalised.}
+  tiny deviations from exactly $90^\circ$ were heavily penalised.\\ \\}
 
@@
 
-\textcolor{blue}{New figure added (Figure 6) which illustrates the implementation}
+\textcolor{blue}{New figure added (Figure 6) which illustrates the
+  implementation\\ \\}
 
@@
 \textcolor{blue}{Done, new subsection ``some reflections on links''
-  at the end, under ``conclusions and further work''}
+  at the end, under ``conclusions and further work''\\ \\}
 
@@
 
-\textcolor{blue}{Canonical URL now added}
+\textcolor{blue}{Canonical URL now added\\ \\}
 

### 2022-06-09 inst/bjourdoc.tex
further work on issue #68

% ------------------------------------------------------------------------
% bjourdoc.tex for birkjour.cls*******************************************
% ------------------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{birkjour}
%%%Optional but convenient to use is the package ``cite''. If you do not want to use, remark the next line by placing the percent sign % in front of it:
\usepackage[noadjust]{cite}
%
%
% THEOREM Environments (Examples)-----------------------------------------
%
 \newtheorem{thm}{Theorem}[section]
 \newtheorem{cor}[thm]{Corollary}
 \newtheorem{lem}[thm]{Lemma}
 \newtheorem{prop}[thm]{Proposition}
 \theoremstyle{definition}
 \newtheorem{defn}[thm]{Definition}
 \theoremstyle{remark}
 \newtheorem{rem}[thm]{Remark}
\newtheorem{comment}[thm]{Comment}
 \newtheorem*{ex}{Example}
 \numberwithin{equation}{section}

\newcommand{\ed}{\end{document}}
\begin{document}

%-------------------------------------------------------------------------
% editorial commands: to be inserted by the editorial office
%
%\firstpage{1} \volume{228} \Copyrightyear{2004} \DOI{003-0001}
%
%
%\seriesextra{Just an add-on}
%\seriesextraline{This is the Concrete Title of this Book\br H.E. R and S.T.C. W, Eds.}
%
% for journals:
%
%\firstpage{1}
%\issuenumber{1}
%\Volumeandyear{1 (2004)}
%\Copyrightyear{2004}
%\DOI{003-xxxx-y}
%\Signet
%\commby{inhouse}
%\submitted{March 14, 2003}
%\received{March 16, 2000}
%\revised{June 1, 2000}
%\accepted{July 22, 2000}
%
%
%
%---------------------------------------------------------------------------
%Insert here the title, affiliations and abstract:
%

\title[Visually pleasing knot projections]{Visually pleasing knot projections}

%----------Author 1
\author[Hankin]{Robin K. S. Hankin}

\address{%
55 Wellesley Street East, Auckland 1010, New Zealand}

\email{hankin.robin@gmail.com}

%----------classification, keywords, date
\subjclass{Primary 99Z99; Secondary 00A00}

\keywords{Class file, HTML, journal}

\date{\today}
%----------additions
%%% ----------------------------------------------------------------------

\begin{abstract}
In this short article I introduce software which creates two
dimensional knot diagrams optimized for visual appearance.  Different
aspects of a knot's appearance are discussed and a framework for
objectively optimizing the visual appeal of a knot projection is
given.
\end{abstract}

%%% ----------------------------------------------------------------------
\maketitle
%%% ----------------------------------------------------------------------
%\tableofcontents
\section{Document Preamble}
Start the article with the command

\begin{verbatim}\documentclass{birkjour}\end{verbatim}

After that, needed macro packages and new commands can be inserted
as in every \LaTeX\ or \AmS-\LaTeX\ document. Don't use commands
that change the page layout (like
\verb+\textwidth, \oddsidemargin+
etc.) or fonts.\bigskip

\section{Frontmatter}
The command
\begin{verbatim}\begin{document}\end{verbatim}
starts -- as always -- the article.

\subsection{Author Data}

Afterwards, insert title, author(s) and affiliation(s), as in the source file to this document,
\verb+bjourdoc.tex+. E.g.,
\begin{verbatim}
\title[An Example for birkjour]
 {An Example for the Usage of the\\ birkjour Class File}
%----------Author 1
\author[Birkh\"auser]{Birkh\"{a}user Publishing Ltd.}
\address{%
Viaduktstr. 42\\
P.O. Box 133\\
CH 4010 Basel\\
Switzerland}
\email{info@birkhauser.ch}
\end{verbatim}
For each author the commands \verb+\author+, \verb+\address+ and \verb+\email+ should be used separately. See the last page of this document for the typesetting layout of the above addresses.

\subsection{Abstract, Thanks, Key Words, MSC}

The \verb+abstract+ environment typesets the abstract:
\begin{verbatim}
\begin{abstract}
The aim of this work is to provide the contributors to edited
books with an easy-to-use and flexible class file compatible
with \LaTeX\ and \AmS-\LaTeX.
\end{abstract}
\end{verbatim}
In addition, the Mathematical Subject Codes, some key words and thanks can be given:
\begin{verbatim}
\thanks{This work was completed with the support of our
\TeX-pert.}
\subjclass{Primary 99Z99; Secondary 00A00}
\keywords{Class file, journal}
\end{verbatim}
Finally, \verb+\maketitle+ typesets the title.

\section{Mainmatter}
Now type the article using the usual \LaTeX\ and (if you need them)
\AmS-\LaTeX\ commands.

We gratefully appreciate if the text does
not contain \verb+\overfull+ and/or \verb+\underfull+ boxes, if equations do not exceed the indicated width, if hyphenations have been checked, and if the hierarchical structure of your article is clear. Please avoid caps and underlines.

Just to give examples of a few typical environments:

\begin{defn}
This serves as environment for definitions. Note that the text appears not in italics.
\end{defn}

\begin{equation}\label{testequation}
\text{This is a sample equation: } c^2=a^2+b^2
\end{equation}

The above equation received the label \verb+testequation+.

\begin{thm}[Main Theorem]
In contrast to definitions, theorems appear typeset in italics as
it has become more or less standard in most textbooks and
monographs. Equations can be cited using the \verb+\eqref+ command which
automatically adds brackets: \verb+\eqref{testequation}+ results in \eqref{testequation}.
\end{thm}

\begin{proof}
A special environment is predefined: the \textit{proof} environment. Please use
\begin{verbatim}\begin{proof}\end{verbatim}
proof of the statement
\begin{verbatim}\end{proof}\end{verbatim}
for typesetting your proofs. The end-of-proof symbol $\Box$ will be added automatically.
\end{proof}

There are two known problems with the placement of the end-of-proof sign:

\begin{enumerate}
  \item if your proof ends with a\ \ s i n g l e\ \ displayed line, the end-of-proof sign would
be placed in the line below; if you want to avoid this, write your line in the form
\begin{verbatim}$$displayed math line \eqno\qedhere$$\end{verbatim}
which results in

\begin{proof}
$$displayed math line \eqno\qedhere$$
\end{proof}
\item if your proof ends with an aligned displayed environment, the command
\verb+\tag*{\qed}+ can be used to place the end-of-proof sign properly:
\begin{verbatim}
\begin{align*}
\alpha&=\beta+\gamma\\
&=\delta+\epsilon\tag*{\qed}
\end{align*}
\end{verbatim}
results in
\begin{align*}
\alpha&=\beta+\gamma\\
&=\delta+\epsilon\tag*{\qed}
\end{align*}
\end{enumerate}
Please try to avoid using the obsolete \verb+\eqnarray+ environment. This environment has several bugs
and has been replaced by the more flexible \AmS\ environments \verb+align, gather, multline, split+.

\begin{rem}
Additional comments are being typeset without boldfaced entrance
word as they may be minor important.
\end{rem}

\begin{ex}
For some constructs, even no number is required.
\end{ex}

Displayed equations may be numbered like the following one:
\begin{equation}
\sqrt{1-\sin^2(x)}=|\cos(x)|.
\end{equation}

\subsection{Here is a Sample Subsection}

Just needed because next thing is

\subsubsection{Here is a Sample for a Subsubsection}

One more sample will follow which clearly shows the difference between subsubsection deeper nested lists:

\paragraph{Here is a Sample for a Paragraph}

As you observe, paragraphs do not have numbers and start new lines after the heading, by default.

\subsection{Indentation}
Though indentation to indicate a new paragraph is welcome, please
do not use indentation when the new paragraph is already marked by
an extra vertical space, as for example in the case of the first
paragraph following a heading (this is standard in this class), or
after using commands like
\verb+\smallskip, \medskip, \bigskip+ etc.

\subsection{Figures}

Please use whenever possible figures in EPS format (encapsulated postscript). Then, you can include the figure with the command

\begin{verbatim}\includegraphics{figure.eps}\end{verbatim}

It is sometimes difficult to place insertions at an exact location
in the final form of the article. Therefore, all figures and tables
should be numbered and you should refer to these numbers within
the text. Please avoid formulations like ``the following
figure\dots".

\subsection{Your Own Macros}

If you prefer to use your own macros within your document(s), either define them in your preamble or, if they are collected in a separate \texttt{.tex} file, please don't forget to upload  that file together with the source file for the manuscript. We will need all these files to produce the final layout. For technical comments on submitting your manuscript files through the Author's Account, please see the next section.

\section{Technical Comments on Submitting Manuscript Files Through 
the Editorial Manager (EM) for AACA} 
\label{sec:technicalcomments}

The purpose of this section is to provide a few hints to Authors how to properly submit manuscript files in order to avoid errors when AACA's EM system produces a pdf file with a new or revised submission. For the sake of our example, our manuscript file is called \texttt{manuscript.tex}.

\begin{comment}
Before you attempt to upload your manuscript files, please make sure that when you typeset your document on your local system, no errors are being reported in \texttt{manuscript.log} file. In particular, that all your citation commands \verb!\cite{...}! create correct reference numbers like \cite{paper}, \cite{latex}, \cite{paper,latex,scharler}, etc. instead of \texttt{[?]}.
\end{comment}

\begin{comment}
While AACA does not require that original submissions are typeset in \LaTeX\ using \texttt{birkjour.cls} class file, or, typeset in \LaTeX\ at all, e.g., they could be typed in MS Word, when the manuscript is accepted for publication, its final version will have to be typeset by the author in \LaTeX\ using \texttt{birkjour.cls}. 
\end{comment}

\begin{comment}
To avoid time delays after the manuscript is accepted, it is faster to submit the original (or, at least its revision, if requested by the Editor) typeset in \LaTeX\ with \texttt{birkjour.cls} as it will later shorten the time before the accepted paper gets published. Also, referees often comment on improper \LaTeX\ usage or features not particular to the \texttt{birkjour.cls} style. This then takes more time for them to submit their reviews.
\end{comment}

\begin{comment}
A single place where time-consuming and error-prone changes often need to be made, is in the References. If the References are not initially formatted according to  MathPhysRef format shown in a file 
\begin{center}
\texttt{Key\_Style\_Points\_MathPhysRef.pdf}
\end{center}
included in \texttt{birkjour.zip}, more time later usually will be needed for the Editorial Office to return Page Proofs of the accepted paper to its Author. See also   \cite{keystylepoints} for more examples how the References need to be formatted.

\begin{comment}
Please be aware that when your paper is accepted, you will receive \textbf{only one set of Page Proofs to correct}. Your Proofs will be emailed to you as a single HTML file with a request to make corrections in that file.

One difficulty with making corrections in that file is that the file will not show pages and line numbers in your paper, and, the margins will be different than those in your paper \LaTeX\ formatted paper.

That is, the HTML file will look like a web page.

\begin{center}
\framebox{%
\begin{minipage}{0.9\linewidth}
It will be a bit difficult to locate places needing correction in the HTML file as these places are usually identified by referees and editors by the page and the line numbers in your \LaTeX\ formatted paper.    
\end{minipage}}
\end{center}

Using built-in icon-driven editorial commands, you will need to make your corrections in the HTML file: text you choose to delete will appear crossed with a red line, while text you add will appear highlighted in yellow.

There is a short learning curve how to use the HTML editor. Correcting or changing mathematical formulas could be really awkward, while modifying text is not that difficult.

However, it is best to make all necessary changes before submitting your final revision through your Author's Account.

Finally, when you submit marked-up HTML file via email back to the Springer Editorial office, you will then receive your last set of Page Proofs 
as a \LaTeX\ formatted \texttt{.pdf} file only to verify that your requested corrections have been inserted properly. \textbf{However, no new corrections will be allowed at that time.}   
\end{comment}

%\vspace*{5ex}
\begin{center}
\framebox{%
\begin{minipage}{0.9\linewidth}
In short, it is a good idea to carefully prepare your manuscript in \LaTeX\ with the required \texttt{birkjour.cls}, including the references in MathPhysRef format, prior to submitting it as it later shortens time from acceptance to when the paper appears OnLineFirst.    
\end{minipage}}
\end{center}
\end{comment}
%\vspace*{5ex}

\noindent
Here are a few common errors that may appear in \texttt{manuscript.pdf} created by the EM:

\begin{itemize}
\item[(a)] Some unwanted additional files get appended at the end, or, the 
\texttt{.pdf} file created by the EM contains two copies of the manuscript because, most likely, you have also uploaded \texttt{manuscript.pdf}.\newline

\noindent
\textbf{Solution:} Do not upload any log files, auxiliary files, etc., that is, files with extensions \texttt{.aux, .bbl, .blg, .log} created by your typesetting system. In particular, do not upload \texttt{manuscript.pdf}.

\item[(b)] Citation of some references appears like [?].\newline

\noindent
\textbf{Solution:} Make sure that all \texttt{cite} commands in \texttt{manuscript.tex} produce correct reference numbers on your local system first, and then upload once again your corrected \texttt{manuscript.tex} file.

\item[(c)] The EM is unable to create \texttt{manuscript.pdf} because some style files are missing.\newline

\noindent
\textbf{Solution:} Upload all needed style files. If you are using Scientific Word, use its ``portable \LaTeX''  format option for saving your 
\texttt{.tex} document and then upload the \texttt{.tex} file.
\end{itemize}

Here are a few practical hints how to upload only the needed manuscript files.

\begin{enumerate}

\item Login into your Author's account at
      \begin{center} 
      \texttt{https://www.editorialmanager.com/aaca/default.aspx}
			\end{center}
      \begin{itemize}
			\item To submit a new manuscript, click on \underline{Submit New Manuscript}.      
			\item To submit a revision to a manuscript already in the system, click on 
			      \underline{Submissions Needing Revision} and in the ``Action Menu'' click 
						on \underline{Revise Submission} and pick the submission you wish to revise.
			\end{itemize}
\item Select ``Article Type'': pick ``Original Research'' or, from the drop-down menu, pick a name of Topical Collection. Proceed.

\item If you are entering:
\begin{itemize}
\item A new manuscript, type in title, authors' names, abstract, keywords separated by ;	(it is best to copy and paste text from your \LaTeX\ document). Answer a questionnaire that appears on the page\footnote{If your manuscript is submitted already to another journal, be aware that AACA's policy is not to consider such manuscripts for  possible publication. Only if you answer ``No'', you can proceed.}.

\item A revision of an already submitted manuscript, you will be able to reuse previously entered title, authors' names, abstract, and keywords, unless they have changed.
\end{itemize}

\item Remember that the EM system does not search on the web for needed style files or the class file: you must upload them all to the system. In particular, you must upload \texttt{birkjour.cls} and any additional style file used, e.g., \texttt{cite.sty}.

\item The order in which files are uploaded and then are listed in the system does matter: always first upload your \texttt{manuscript.tex} file\footnote{If, by mistake, you upload another file first, or, you upload files not in the order described below, you will be later able to re-order the uploaded files. The reordering procedure is described below.}. If you upload it first, it will appear as Number~1 in a submission window as
\begin{verbatim}
1. manuscript.tex
\end{verbatim}
Select ``manuscript" as the file type in the window next to the file name.

\item DO NOT UPLOAD any other files that your local \LaTeX\ system has created such as 
\begin{center}
\texttt{manuscript.aux}, \texttt{manuscript.log}, 
\texttt{manuscript.bbl}, \texttt{manuscript.blg}, \texttt{manuscript.pdf}
\end{center}
since they are not needed by the EM to create \texttt{manuscript.pdf}. In particular, DO NOT upload \texttt{manuscript.pdf} either.

\begin{comment}
If you upload \texttt{manuscript.log} or any other text/ASCII file, it will be appended to your paper inside a \texttt{.pdf} file that the EM will create. Thus, do not upload this file or any other text/ASCII file, ONLY upload \texttt{manuscript.tex}.
\end{comment}

\begin{comment}
Sometimes authors store definitions of \LaTeX\ macros in an additional \texttt{.tex} file, call it \texttt{macros.tex}, which is later read into the manuscript with a command \verb+\input{macros}+. If so, the file 
\texttt{macros.tex} needs to be uploaded and listed in the system window as Number~2 file.
\end{comment}

\item After uploading \texttt{manuscript.tex} as Number~1 file (or, after uploading \texttt{macros.tex} as Number~2 file), upload all style files with extension \texttt{.sty} needed to typeset your manuscript. For example,

\begin{itemize}
\item \texttt{tikz-cd.sty} for making commutative diagrams,
\item \texttt{cite.sty} for more nicely arranged references,
\end{itemize}
and so on.

\item Lastly, upload \texttt{birkjour.cls} style file.

\item Here is an example of how that window in the EM could look:

\begin{verbatim}
1. manuscript.tex
2. tikz-cd.sty
3. cite.sty
4. birkjour.cls
\end{verbatim}

\begin{comment} 
If your order of files is different than the one shown above, there is a way of reordering these files: in a small window you need to type the order in which files must appear. Then, the system will reorder your files according to the supplied order.

This reordering would need to be done if you make some changes in your 
\texttt{manuscript.tex file}: you would then need to delete the old \texttt{manuscript.tex} in the system by using buttons in the window (see the next comment), upload new \texttt{manuscript.tex}, and then reorder all files making sure that \texttt{manuscript.tex} is listed first. This is because files that have been uploaded, automatically are listed in the order of the upload.
\end{comment}

\begin{comment}
If by mistake, you have uploaded not needed files, to remove them, follow these steps:  
\begin{enumerate}
\item check off each file to be removed: there is a small box on the same level as the file name,

\item once all files to be removed have been checked, click on ``remove selected files''.
\end{enumerate}
\end{comment}

\item It pays off not to load in your paper preamble \LaTeX\ packages you do not actually use in your paper as then you might have to upload also a style file for each package. For example, the command
\begin{verbatim}
\usepackage{cite}
\end{verbatim}
requires that in your paper directory you also have file \texttt{cite.sty}. If not, the paper won't process correctly. Thus, remark or eliminate commands in your preamble that load packages you don't actually use in the paper. For example, AACA's EM had a problem with this command
\begin{verbatim}
\usepackage[utf8]{inputenc}
\end{verbatim}
but since I really did not need that package, remarking it out in the preamble of my paper fixed the problem.

\item  When AACA's EM system creates \texttt{manuscript.pdf} file and errors occur, it creates a log file that looks like a log file from any \LaTeX\ typesetting system, e.g., \texttt{TeXnicCenter}. That file can be looked at to see why the error has occurred (if you know how to read it). 
\end{enumerate}

From my experience, a bad looking unacceptable \texttt{manuscript.pdf} file is created by the system if the above points are not followed. In that case, do not accept it and have the system recreate it by first making sure that the above points have been followed.

%If you need any help or have any questions, contact the second author of this paper.

\section{Backmatter}

At the end of the document, the affiliation(s) will be typeset
automatically. For this it is necessary that you used the \verb+\address+ command for including your affiliation, as explained above.

\subsection{References}

Please note that references must appear in the MathPhysRef format shown in the file 
\begin{center}
\texttt{Key\_Style\_Points\_MathPhysRef.pdf}
\end{center}
included in \texttt{birkjour.zip}. See also   \cite{keystylepoints} for more examples how the References need to be formatted. When the references are properly formatted already in the original manuscript file, this saves plenty of time later when the manuscript is accepted and page proofs are created. In particular, the references need to be alphabetized by the last name of the first author.

Please cite your references with the command \verb+\cite{...}+ typed as, for example, \verb+\cite{latex,scharler}+ which produces \cite{latex,scharler}.

All references need to be cited in the text: if a reference is not cited, it should be removed from the References.

A package called \verb+cite+ conveniently manages references especially when the author cites ranges of references. Command
\begin{verbatim}
\usepackage[noadjust]{cite}
\end{verbatim}
placed in the Preamble loads the package provided a file \verb+cite.sty+ 
(included in \texttt{bjourdoc.zip}) is placed in the same directory as the manuscript file.  Then, the command \verb+\cite{scharler,paper,latex}+ produces 
\cite{scharler,paper,latex} instead of $[3,\,1,\,2]$.

Also, \texttt{cite} puts labels of the cited references in the ascending order. Typing \verb+\cite{latex,paper}+ produces \cite{latex,paper} when the package is used instead of $[2,\,1]$ when the package is not used.

If you do use the package \verb+cite+, do not forget to upload the file \verb+cite.sty+ with your manuscript as explained in Section~\ref{sec:technicalcomments}.

At \texttt{https://link.springer.com/journal/6}, the Latest Articles appearing in AACA are posted.  Please note that at the top of each article page, there is a link \underline{Cite this article} which provides a correct way to cite the given article. For example, on the page\\

\texttt{https://link.springer.com/article/10.1007/s00006-019-1037-1},\\

\noindent
we find an article entitled ``Quadratic Split Quaternion Polynomials: Factorization and Geometry'' which should be cited as shown in \cite{scharler}.

The DOI numbers on the web become links which lead to the web pages with posted articles. Thus, it is recommended that they be used as they enable searches and viewing abstracts on the web.

% ------------------------------------------------------------------------

\subsection*{Acknowledgment}
Many thanks to our \TeX-pert for developing this class file.

%Paper:
%Balli, S., Chand, S.: Transmission angle in mechanisms. Mech. Mach. Theory 37(2), 175–195 (2002)
%Book:
%Bayro-Corrochano, E.: Geometric Computing: for Wavelet Transforms, Robot Vision, Learning, Control and Action. Springer Publishing Company Inc., Berlin (2010)

\begin{thebibliography}{1}
%%%1
\bibitem{paper} 
Bekar,~M. and Yayl{\i},~Y.: Involutions in dual-split quaternions. Adv. Appl. Clifford Algebras \textbf{26}(2), 553--571 (2006). 
\texttt{https://doi.org/10.1007/s00006-015-0624-z}
%%%2
\bibitem{latex} 
Gr\"atzer,~G.: Math into \LaTeX. 3rd edn., Birkh\"auser, Boston (2000)
%%%3
\bibitem{scharler}
Scharler,~D.~F., Siegele,~J. \& Schr\"{o}cker,~H.~P.: Quadratic split quaternion polynomials: Factorization and geometry. Adv. Appl. Clifford Algebras (2020) 30: 11. \texttt{https://doi.org/10.1007/s00006-019-1037-1}
%%%4
\bibitem{keystylepoints}
\texttt{\small https://www.springer.com/cda/content/document/cda\_downloaddocument/ \newline Key\_Style\_Points\_Aug2012.pdf?SGWID=0-0-45-1340009-0}
%%%5
\bibitem{test} Test,~A.~B.~C.: On a test. J. of Testing \textbf{88}, 100--120 (2000)
\end{thebibliography}

% ------------------------------------------------------------------------
\end{document}
% ------------------------------------------------------------------------

### 2022-06-09 inst/jma_rebuttal_review3.tex
further work on issue #68

@@
 
-\textcolor{blue}{I have not implemented sliders.  The weightings may
-  be changed in function {\tt badness()}, and documentation (including
-  examples) is given at {\tt badness.Rd}.\\ \\}
+\textcolor{blue}{I have not implemented sliders.  A typical
+  optimization takes a couple of hours, a big one [e.g. some of the
+    pretzel knots in the package] may take a few weeks.  With a slider
+  interface, one would adjust a slider's position and have to wait
+  several days to see the effect.  Weightings may be changed in
+  function {\tt badness()}, and documentation (including examples) is
+  given at {\tt badness.Rd}.\\ \\}
 

### 2022-06-09 inst/jma_rebuttal_review3.tex
fixes #53

@@
 possess a C2 rotation axis, which ..."
+
+\textcolor{blue}{I do not believe that describing the symmetry in
+  question as a C2 rotation axis is appropriate.  A C2 axis refers to
+  a three-dimensional object which is unchanged by a $180^\circ$
+  rotation about a particular axis in 3D space.  I am considering the
+  knot path here, which is a plane figure.  Referring to a C2 axis
+  might suggest to readers that I am considering a 3D figure [e.g. the
+    result of ``lifting'' the strand slightly at a crossing point].
+  But this particular 3D figure might not possess a C2 axis, a
+  potentially confusing (and anyway irrelevant) fact.  On balance, I
+  believe that mirror symmetry is the appropriate concept here, but
+  the original sentence was poorly worded.\\ \\
+OLD TEXT: ...possess a line of symmetry (at least if...ignored)\\ \\
+NEW TEXT: ...possess mirror symmetry (at least if...ignored)\\ \\
+}
 

### 2022-06-09 inst/jma_knot.tex
fixes #53

@@
 as posters: they are not vectorized.  Many of the underlying knots
-possess a line of symmetry (at least, the diagrams do if the breaks
+possess mirror symmetry (at least, the diagrams do if the breaks
 are ignored), which is not present in the visual representation.

### 2022-06-09 inst/jma_knot.tex
pointers to storage locations or creation methods for image files

@@
   \centering
-  \includegraphics[width=10cm]{"1280px-Knot_table".png}
+  \includegraphics[width=10cm]{"1280px-Knot_table".png}  % canonical version in vignettes/ directory
   \caption{A table of prime knots up \label{knot_table} to seven
@@
   \centering
-    \includegraphics[width=10cm]{screenshot_inkscape_7_6.png}
+    \includegraphics[width=10cm]{screenshot_inkscape_7_6.png} % held in vignettes/ directory
 \caption{Screenshot of inkscape\label{screenshot_inkscape_7_6} setup
@@
   \centering
-\includegraphics[scale = 0.9]{knot-k76_knotplot2}
+\includegraphics[scale = 0.9]{knot-k76_knotplot2} % in vignettes/ directory
 \caption{Knot $7_6$, post-optimization\label{7_6}}
@@
   \centering
-\includegraphics[scale = 0.9]{knot-009}
+\includegraphics[scale = 0.9]{knot-009}% in vignettes/ directory
 \caption{Knot $7_6$, post-optimization with breaks indicating
@@
   \centering
-\includegraphics[width=11cm]{knot-four_figure_eight_knots}
+\includegraphics[width=11cm]{knot-four_figure_eight_knots}  % in vignettes/ directory
 \caption{Figure eight knots drawn using different plotting methods.
@@
 \end{verbatim}
-\includegraphics[width=11cm]{knot-k5_1}
+\includegraphics[width=11cm]{knot-k5_1}  % in vignettes/ directory
 \caption{Knot $5_1$ \label{k5_1_twoknots} shown with node numbers}
@@
   \centering  
-  % To create file "two_trefoil_knots_post_inkscape.pdf", run
-  % knotter.R, which makes (among other things) two_trefoil_knots.pdf.
-  % Edit this in inkscape [make a mirror image, then change the
-  % colours and finally use the raise and lower tools to put
-  % everything in the right order]
+  % To create file "two_trefoil_knots_post_inkscape.pdf" (which is
+  % stored on the repo), run knotter.R, which makes (among other
+  % things) two_trefoil_knots.pdf.  Edit this in inkscape [make a
+  % mirror image, then change the colours and finally use the raise
+  % and lower tools to put everything in the right order]
     \includegraphics[width=10cm]{two_trefoil_knots_post_inkscape.pdf}
@@
 \end{verbatim}
-\includegraphics[width=11cm]{knot-perko_A_and_B}
+\includegraphics[width=11cm]{knot-perko_A_and_B}  % in vignettes/ directory
 \caption{Two representations of knot~$10_{125}$, known as the 
@@
 \end{verbatim}
-\includegraphics[width=11cm]{knot-ornamental}
+\includegraphics[width=11cm]{knot-ornamental} % in vignettes/ directory
 \caption{An ornamental knot exhibiting fivefold

### 2022-06-10 inst/jma_knot.tex
fixes #64

@@
 minimum found by {\tt nlm()} is topologically identical to the
-starting configuration.  Function {\tt
-  non\_crossing\_strand\_close\_approach\_badness()}, for example,
+starting configuration (a list is given in the Appendix).  Function
+{\tt non\_crossing\_strand\_close\_approach\_badness()}, for example,
 makes non-crossing strands ``repel'' one another so as not to
@@
 \bibliographystyle{unsrt}
+
+\clearpage
+\section*{Appendix}
+
+There follows a more detailed description and motivation for various
+metrics used to evaluate knot diagrams in the package.
+
+\begin{description}
+\item[{\tt total\_crossing\_potential\_energy()}] The idea is to make
+  the crossing points attract one another as if they are
+  gravitationally active objects; metric measured by their mutual
+  potential energy.  If crossing point $i$ is at position ${\mathbf
+    p}_i$ we have $\sum_{i\neq j}\left|{\mathbf p}_i-{\mathbf
+    p}_j\right|^{2}$.  The idea is to encourage more compact
+  configurations rather than long skinny ones.
+\item[{\tt total\_crossing\_angles()}] Here we wish to punish
+  configurations with strands that cross at acute angles.  If crossing
+  point $i$ has crossing angle $\theta_i$ the metric is
+  $\sum_i\left(\cos\theta_i\right)^2$.
+\item[{\tt total\_bending\_energy()}] The idea is for the strands to
+  preferentially follow elastica curves, as in elementary (nonlinear)
+  beam theory.  From differential geometry, the Euler elastica
+  potential energy is $\int\kappa(s)^2\,ds$.  In the package, this is
+  calculated numerically for each strand and the total returned.
+\item[{\tt total\_string\_length()}] The length is just $\int ds$, with a
+  nominal value of 5000 (which value is about right for a printed
+  copy).
+\item[{\tt midpoint\_badness()}] The idea is to suppress strands with
+  crossing points too close to the endpoint nodes.  For each crossing
+  point $c_ij$ of strand $i$ with strand $j$, we calculate the arc
+  length $s=\int_{s=0}^{c_{ij}}\,ds$ from the node to each crossing
+  point and normalize with the strand length $\int\,ds$.  We reward
+  configurations with normalized crossing point arc lengths that are
+  close to $1/2$.  Note that this particular metric is not independent
+  of the total crossing potential energy metric.
+\item[{\tt node\_crossing\_badness()}] Somewhat complementary to {\em
+  total crossing potential energy}, here we make the crossing points
+  repel one another, as though they are all charged objects (with the
+  same sign).  The idea is to spread out the nodes so no two nodes are
+  too close together.  If crossing point $i$ is at position ${\mathbf
+    p}_i$ we have $\sum_{i\neq j}\left|{\mathbf p}_i-{\mathbf
+    p}_j\right|^{-1}$.
+\item[{\tt curvature\_switching\_badness()}] Here the problem is knots
+  which should have either all their strands curving to the left, or
+  all to the right; the signed curvature should be everywhere the same
+  sign (the canonical example would be $8_{18}$ which in earlier
+  versions of the software had ugly changes of curvature).  We impose
+  a fixed penalty for any knot with more than one sign.
+\item[{\tt curvature\_consecutive\_segment\_switching\_badness()}] We
+  wish to discourage configurations in which curvature undergoes a
+  sudden change at a node.  The value for a single (Bezier) node is
+  simply $\left(\delta\kappa\right)^2$, and the penalty for a given
+  configuration is the sum of the values for each node.  See
+  Figure~\ref{two_trefoil_knots} for motivation and discussion.
+\item[{\tt non\_crossing\_strand\_close\_approach\_badness()}] Here
+  the idea is to prevent non-intersecting strands from approaching one
+  another too closely.  Quite apart from aesthetic considerations,
+  this metric prevents the creation of new spurious nodes that would
+  disrupt the optimization routines.
+\end{description}
+
 \end{document}

### 2022-06-10 inst/jma_knot.tex
slight rearrangement for issue #64

@@
   configurations rather than long skinny ones.
+\item[{\tt node\_crossing\_badness()}] Here we make the crossing
+  points {\em repel} one another, as though they are all charged
+  objects (with the same sign).  The idea is to spread out the nodes
+  so no two nodes are too close together.  If crossing point $i$ is at
+  position ${\mathbf p}_i$ we have $\sum_{i\neq j}\left|{\mathbf
+    p}_i-{\mathbf p}_j\right|^{-1}$.  The length scale for this
+  repulsion is smaller than that of the attractive term above.  The
+  overall potential thus resembles that sometimes seen in atomic
+  physics when considering the hydrogen molecule $H_2$: the two
+  protons attract one another when distant, but repel when closer.  In
+  the package, the intent is for this optimal distance to be a shallow
+  minimum and the optimization routine is free to adjust within a wide
+  range.
 \item[{\tt total\_crossing\_angles()}] Here we wish to punish
@@
   crossing points too close to the endpoint nodes.  For each crossing
-  point $c_ij$ of strand $i$ with strand $j$, we calculate the arc
-  length $s=\int_{s=0}^{c_{ij}}\,ds$ from the node to each crossing
+  point $c_{ij}$ of strand $i$ with strand $j$, we calculate the arc
+  length $\int_{s=0}^{c_{ij}}\,ds$ from the node to each crossing
   point and normalize with the strand length $\int\,ds$.  We reward
@@
   of the total crossing potential energy metric.
-\item[{\tt node\_crossing\_badness()}] Somewhat complementary to {\em
-  total crossing potential energy}, here we make the crossing points
-  repel one another, as though they are all charged objects (with the
-  same sign).  The idea is to spread out the nodes so no two nodes are
-  too close together.  If crossing point $i$ is at position ${\mathbf
-    p}_i$ we have $\sum_{i\neq j}\left|{\mathbf p}_i-{\mathbf
-    p}_j\right|^{-1}$.
 \item[{\tt curvature\_switching\_badness()}] Here the problem is knots

### 2022-06-10 inst/jma_rebuttal_review3.tex
fixes #64
fixes #60

@@
 
+\textcolor{blue}{Each coloured circle is linked to a Bezier point and
+  this is set to 100 by default.  This is technically user-settable
+  but to make it much fewer than 100 would seriously erode the quality
+  of the strand plot.\\ \\}
+
 Figure 4.  Knot $7_6$ with strands numbered... -- The main use here seems
@@
 trade-offs that are being made in the optimization process.
+\textcolor{blue}{I have added an appendix detailing the metrics used.
+  Although the functions are described mathematically in the software,
+  this appendix gives some insight into why the functions are
+  there\\ \\}
 

### 2022-06-15 inst/jma_knot.tex
fixes #48

@@
 knotted strands.
+
+\subsection{Visual knot diagrams: summary}
+
+Two key questions emerge.  Firstly, {\em What makes a knot plot
+  attractive?}, and secondly {\em How best to create knot plots with
+  the desired qualities?}  I address both questions, and attempt to
+capture the aesthetics of knot diagrams using mathematical
+formulations.  These formulations are of two types: hard constraints
+of different forms of symmetry; and quantification of various
+undesirable qualities of a knot diagram that may be minimized
+numerically.  I present software and workflows that not only produce
+pleasing knot diagrams, but also allow one to investigate the
+connection between the aesthetics of knot diagrams and their
+mathematics.
 

### 2022-06-15 inst/jma_rebuttal_review3.tex
fixes #75

@@
 
+\textcolor{blue}{This revision incorporates many referee's suggestions
+  into the literature review.  I would agree that the literature
+  review is tightly focussed on the aesthetics of knot diagrams.
+  Although it is clear that many workers place high value on pleasing
+  diagrams, few sources explicitly address the issue.  I would of
+  course be happy to cite any suggested references.}
+
 \end{document}

### 2022-06-15 inst/jma_rebuttal_review3.tex
more work on issue #52
minor tweaks

@@
 
+\textcolor{blue}{
 Short story: the reviewers make reasonable and constructive comments.
@@
 document is, I believe, stronger and more scholarly than before and I
-recommend it to you.
+recommend it to you.}
 
@@
   submission as ``decent'', which I take to be a great compliment.  It
-  is often difficult to know the correct balance between overly
-  technical descriptions [which indeed often overlaps with the aims of
-    a technical manual] and gushing conceptual big-picture thinking.
-  However, I note that many of the comments below are essentially
-  requests for clarifications of minor technical details of the
-  implementation: and this would indicate that perhaps the balance
-  should be shifted back toward the technical manual end of the
-  spectrum.  The other observation I would make here is that the
-  software package [27] is intended to be part of the submission (it
-  is released under the GPL and is thus available for public use) and
-  with this understanding, even a downbeat technical manual would have
-  scholarly value in conjunction with the new software it describes.
-  I want the editor and reviewer to know that in addressing the
-  concerns below I have tried hard to answer not just ``what'' but
-  also the ``why'' of the implementation.  \\ \\ Nevertheless, the cue
-  above is a good one.  The two questions: (a) {\em what makes a knot
-    plot attractive?}, and (b) {\em how best ... to achieve this?}
-  are indeed profound, and were not addressed directly in the old
-  manuscript.  In the revision I have prominently stated the questions
-  in the introduction, and given a provocative, if brief, outline of
-  my view.\\ \\}
+  is often difficult to know the correct balance between an overly
+  technical description [which indeed often overlaps with the aims of
+    a technical manual] and conceptual big-picture thinking.  However,
+  I note that many of the comments below are essentially requests for
+  clarifications of minor technical details of the implementation: and
+  this would indicate that perhaps the balance should be shifted back
+  toward the technical manual end of the spectrum.  The other
+  observation I would make here is that the software package [27] is
+  intended to be part of the submission (it is released under the GPL
+  and is thus available for public use) and with this understanding,
+  even a downbeat technical manual would have scholarly value in
+  conjunction with the new software it describes.  I want the editor
+  and reviewer to know that in addressing the concerns below I have
+  tried hard to answer not just ``what'' but also the ``why'' of the
+  implementation.  \\ \\ Nevertheless, the cue above is a good one.
+  The two questions: (a) {\em what makes a knot plot attractive?}, and
+  (b) {\em how best ... to achieve this?}  are indeed profound, and
+  were not addressed directly in the old manuscript.  In the revision
+  I have prominently stated the questions in the introduction (section
+  1.2), and given a provocative, if brief, outline of my view.\\ \\}
 
@@
   specify the path of a knot projection; here cubic Bezier curves are
-  used''.  However, Bezier curves are a natural choice for a number of
-  reasons and I have added the following sentence: ``Bezier curves are
-  a natural choice for working with mathematical knots for several
-  reasons: they are familiar to many graphical workers; they have a
-  natural and intuitive control system (usually called ``handles''),
-  and are implemented in many graphical software suites''.  I do not
-  wish to suggest that Bezier curves are the only parametrizations for
-  working with knots; but they are a natural starting point, and
-  AFAICS they have no serious disadvantages in this context, nor is
-  there any obvious superior method.\\ \\}
+  used''.  However, as observed, there are alternatives.  I have added
+  the following sentence: ``Bezier curves are a natural choice for
+  working with mathematical knots for several reasons: they are
+  familiar to many graphical workers; they have a natural and
+  intuitive control system (usually called ``handles''), and are
+  implemented in many graphical software suites''.  I do not wish to
+  suggest that Bezier curves are the only parametrizations for working
+  with knots; but they are a natural starting point, and AFAICS they
+  have no serious disadvantages in this context, nor is there any
+  obvious superior method.\\ \\}
 
@@
   function actually returns $\sum_{i=1}^k\left(\cos\theta_i\right)^2$.
-  \\ \\}
+  \\ \\ The function is now named {\tt
+    total\_crossing\_angle\_badness()}, which more accurately
+  describes its purpose.\\ \\}
 
@@
   might suggest to readers that I am considering a 3D figure [e.g. the
-    result of ``lifting'' the strand slightly at a crossing point].
-  But this particular 3D figure might not possess a C2 axis, a
-  potentially confusing (and anyway irrelevant) fact.  On balance, I
-  believe that mirror symmetry is the appropriate concept here, but
-  the original sentence was poorly worded.\\ \\
-OLD TEXT: ...possess a line of symmetry (at least if...ignored)\\ \\
-NEW TEXT: ...possess mirror symmetry (at least if...ignored)\\ \\
-}
+    result of taking a 2D figure and perturbing strands slightly at a
+    crossing point to avoid self-intersection].  But this particular 3D
+  figure might not possess a C2 axis, a potentially confusing (and
+  anyway irrelevant) fact.  On balance, I believe that mirror symmetry
+  is the appropriate concept here, but the original sentence was
+  poorly worded.\\ \\ OLD TEXT: ...possess a line of symmetry (at
+  least if...ignored)\\ \\ NEW TEXT: ...possess mirror symmetry (at
+  least if...ignored)\\ \\ }
 
@@
   package is predicated upon the notion that the diagram is the
-  ``real'' object, and the (3D) knot itself is irrelevant.  Thus much
-  of the package {\em prevents} the optimization routine from making
-  Reidermeister moves [that is, changes that affect the topology of
-    the diagram but not that of the 3D knot].  The operational reason
-  for this is that the user specifies a particular diagram which the
-  software then optimizes for appearance.  The manuscript includes two
-  diagrams of $10_{125}$, the Perko pair, which show that sometimes
-  one wants to preserve the topology of the diagram.\\ \\}
+  ``real'' object, and the (3D) knot itself is to some extent
+  irrelevant.  Thus much of the package {\em prevents} the
+  optimization routine from making Reidermeister moves [that is,
+    changes that affect the topology of the diagram but not that of
+    the 3D knot].  The operational reason for this is that the user
+  specifies a particular diagram which the software then optimizes for
+  appearance.  The manuscript includes two diagrams of $10_{125}$, the
+  Perko pair, which show that sometimes one wants to preserve the
+  topology of the diagram.\\ \\}
 
@@
 \textcolor{blue}{``Nodes'' is a standard word in the context of Bezier
-  curves. I have not used the word in any other sense in the
-  manuscript and I think the text is clear as it stands.\\ \\}
+  curves, defined in the caption of figure 2.  I have not used the
+  word in any other sense in the manuscript and I think the text is
+  clear as it stands.\\ \\}
 
@@
 \textcolor{blue}{Good point, I have added a sentence to this effect to
-  the caption of figure 4.\\ \\}
+  the caption of figure 4 which now ends with `` Crossings become
+  relevant only for the final rendering of the knot curve with the
+  gaps included''.  \\ \\}
 
@@
 \textcolor{blue}{The referee is correct to say that changing the
-  badness function would result in figures less like 9a and more like
-  9b.  But I would take issue with the phrase ``too tightly wound''.
-  I have polled my colleagues who appear to be evenly split between
-  preferring 9a and 9b.  The comparison appears to elicit strong
-  opinions here at AUT.  Proponents of 9b point to the broader winding
-  along the central helix, while those of 9a point to the more liquid
-  and languid appearance, and in particular emphasise the absence of
-  retrograde curvature at the bottom.  I have added a brief discussion
-  of these issues to the caption of figure 12 [which is now figure 15
-    of the resubmission].  I would observe that the {\tt knotR}
-  package makes such discussions possible.\\ \\}
+  badness function would result in diagrams less like Figure 10a and
+  more like 10b.  But I would take issue with the phrase ``too tightly
+  wound''.  I have polled my colleagues who appear to be evenly split
+  between preferring 10a and 10b.  The comparison appears to elicit
+  strong opinions here at AUT.  Proponents of 10b point to the broader
+  winding along the central helix, while those of 10a point to the
+  more liquid and languid appearance, and in particular emphasise the
+  absence of retrograde curvature at the bottom.  I have added a brief
+  discussion of these issues to the caption of figure 12 [which is now
+    figure 15 of the resubmission].  I would observe that the {\tt
+    knotR} package makes such discussions possible.\\ \\}
 

### 2022-06-15 inst/jma_knot.tex
minor tweaks

@@
 Two key questions emerge.  Firstly, {\em What makes a knot plot
-  attractive?}, and secondly {\em How best to create knot plots with
+  attractive?}, and secondly, {\em How best to create knot plots with
   the desired qualities?}  I address both questions, and attempt to
@@
 two-dimensional projection of a knot's embedding
-$\mathbb{E}\in\mathbb{R}^3$.  However, this approach often results in
+$\mathbb{E}\subset\mathbb{R}^3$.  However, this approach often results in
 poor visual appearance: cusps or other displeasing effects can occur.

### 2022-06-15 inst/jma_knot.tex
move link discussion to appendix B also more futher work

@@
 such diagrams in terms of a customizable objective function using
-numerical optimization techniques.
+numerical optimization techniques.  The software makes a large number
+of pleasing knot diagrams available to the community under the GPL.
+
+One suggestion is that the software might be used in preference tests.
+Subjects might be asked to state which of two knots [e.g. the two
+  shown in Figure~\ref{figure_cca}] they find more attractive.  This
+might allow us to make inferences about the mathematical basis for
+aesthetics of knot diagrams---or perhaps to quantify some features of
+subjective decision-making.
+
+Further work might include adding functionality to deal with links,
+although this is not straightforward, and a brief discussion is given
+in Appendix B.
 
@@
 mathematical formulations of aesthetics.
-
-\subsection{Some reflections on links}
-
-Further work might include functionality to deal with links.  Much of
-the required functionality is (in principle) already present in the
-software.  Each component would have its own badness, and in addition
-there would be $n(n-1)/2$ inter-component interaction terms measuring
-features such as closeness between non-intersecting strands of
-different components.  However, the implementation of such
-functionality is not straightforward and several problems stand out.
-Firstly, dealing with inter-component intersection points is
-difficult.  Currently, crossing points are represented as a
-(symmetrical) Boolean matrix with rows and columns corresponding to
-strands, and entry $(i,j)$ being whether strand $i$ intersects with
-strand $j$.  The link generalization would be a symmetric matrix with
-entries that are themselves intersection matrices: an object with four
-indices and entry $(i,j,k,l)$ indicating whether strand $j$ of
-component $i$ intersects strand $l$ of component $k$.  Such objects
-are unwieldy to work with and are typically difficult to manipulate.
-Secondly, the imposition of symmetry is not straightforward in the
-case of links: one might desire that certain components of a link have
-a particular kind of symmetry and others to have a different kind (or
-no!) symmetry.  Further, one unexpectedly difficult problem was
-dealing with a subset of components that individually possess no
-symmetry but collectively possess mirror (or rotational) symmetry.
-Even the simplest links, as in Figure~\ref{thistlethwaite}, can be
-problematic.  Consider Thistlethwaite's L4a1, for example: one would
-expect the two components to be identical except for a 90-degree
-rotation, and implementing this in the context of the package does not
-seem to be at all easy.  In general, new badnesses would be needed.
-Further, in L7a1, the smaller component should be not only as circular
-as possible (?), but also one might desire the overall length to be
-smaller than that of the other component.  It would be reasonable to
-conclude that dealing with links might be possible in principle but,
-due to a number of technical and mathematical reasons, considerably
-harder than the single-component links considered here.
-
-\begin{figure}[htbp]
-  \begin{center}
-\includegraphics[width=7cm]{L4a1_L7a1}  % made by exporting L4a1_L7a1.svg as PDF
-\caption{Two links \label{thistlethwaite} taken from the
-  Thistlethwaite link table~\cite{hoste1998}.  These are produced
-  directly with inkscape and are not optimized}
-  \end{center}
-\end{figure}
 
@@
 \clearpage
-\section*{Appendix}
+\section*{Appendix A: details and motivation for the badness functions}
 
@@
 
+\section*{Appendix B: some reflections on links}
+
+Further work might include functionality to deal with links.  Much of
+the required functionality is (in principle) already present in the
+software.  Each component would have its own badness, and in addition
+there would be $n(n-1)/2$ inter-component interaction terms measuring
+features such as closeness between non-intersecting strands of
+different components.  However, the implementation of such
+functionality is not straightforward and several problems stand out.
+Firstly, dealing with inter-component intersection points is
+difficult.  Currently, crossing points are represented as a
+(symmetrical) Boolean matrix with rows and columns corresponding to
+strands, and entry $(i,j)$ being whether strand $i$ intersects with
+strand $j$.  The link generalization would be a symmetric matrix with
+entries that are themselves intersection matrices: an object with four
+indices and entry $(i,j,k,l)$ indicating whether strand $j$ of
+component $i$ intersects strand $l$ of component $k$.  Such objects
+are unwieldy to work with and are typically difficult to manipulate.
+Secondly, the imposition of symmetry is not straightforward in the
+case of links: one might desire that certain components of a link have
+a particular kind of symmetry and others to have a different kind (or
+no!) symmetry.  Further, one unexpectedly difficult problem was
+dealing with a subset of components that individually possess no
+symmetry but collectively possess mirror (or rotational) symmetry.
+Even the simplest links, as in Figure~\ref{thistlethwaite}, can be
+problematic.  Consider Thistlethwaite's L4a1, for example: one would
+expect the two components to be identical except for a 90-degree
+rotation, and implementing this in the context of the package does not
+seem to be at all easy.  In general, new badnesses would be needed.
+Further, in L7a1, the smaller component should be not only as circular
+as possible (?), but also one might desire the overall length to be
+smaller than that of the other component.  It would be reasonable to
+conclude that dealing with links might be possible in principle but,
+due to a number of technical and mathematical reasons, considerably
+harder than the single-component links considered here.
+
+\begin{figure}[htbp]
+  \begin{center}
+\includegraphics[width=7cm]{L4a1_L7a1}  % made by exporting L4a1_L7a1.svg as PDF
+\caption{Two links \label{thistlethwaite} taken from the
+  Thistlethwaite link table~\cite{hoste1998}.  These are produced
+  directly with inkscape and are not optimized}
+  \end{center}
+\end{figure}
+
 \end{document}
-
-% ------------------------------------------------------------------------
-
-% ------------------------------------------------------------------------
-\end{document}
-% ------------------------------------------------------------------------

### 2023-01-30 inst/jma_rebuttal_2023_01_30.tex
start work on issue #15

\documentclass[12pt]{article}
\usepackage{xcolor}
\begin{document}

(Below, the reviewer's comments are in black, and my replies to the
issues are in \textcolor{blue}{blue}.  I have indicated changes to the
manuscript where appropriate.

\section*{Rebuttal to "Visually pleasing knot projections" by
  Robin K. S. Hankin (submitted to JMA, manuscript 209604076); review
  30 January 2023}

\textcolor{blue}{ Short story: }

\section*{Detailed rebuttal}

Comments from the Editors and Reviewers:

I appreciate and accept most of the comments made by the author in his
rebuttal and the corresponding fixes in the manuscript.  But I have a
remaining concern that much of this material is perfect for an on-line
tutorial on how to use package [27] to draw optimized knot diagrams,
but JMA is not the right place to publish such a tutorial.

The content that is most fitting for a JMA paper is a discussion of
what makes a knot diagram "beautiful"  and that discussion will
clearly be biased by the personal view of the author (perhaps enhanced
by a "user-study" among family and colleagues).  Thus lines 14 - 20 on
page 4 are of prime importance and could be fleshed out more.

The main contribution to JMA would then be a detailed discussion of
how the author has struggled to capture these goals in the form of
mathematical functions that can be optimized and will then approximate
these goals.  In this respect, the manuscript still has some
shortcomings.

They may be fixed in two ways: Enhance the detailed explanations of
the choices made in creating package [27]; and add a more extended
discussion of a few things that might be approached differently in a
future package ( -- which can also handle links).

In view of this, I recommend that Appendices A and B should be firmly
integrated into the main text.  The discussions therein relate rather
directly to what I see as the main contribution of a JMA paper on this
topic.

A few comments on some rebuttals:

RE: Why use Bezier curves? : A key point, which should be mentioned in
the justification for Bezier curves, is that they interpolate their
end-points. This gives users a convenient, more direct control of
where these curve segments will appear than what they would get from
G2-continuous B-splines, which are only approximating their control
points.

RE: Why not place nodes at the crossing points? : Yes, this will
require a different, slightly more sophisticated drawing interface;
but this may then pay off heavily in a simpler, more efficient
optimization process.

For each crossing point, the user would see an orthogonal "cross-hair"
that can be shifted and rotated as a whole.  In addition, there are
two parameters (handle-pairs) that control the "velocities" of the two
strands through the crossing.  The optimization routine would
fine-tune these velocities to achieve matching curvature for the two
sequentially connected Bezier curves.

And, yes, if a crossing point falls on a line of mirror symmetry, the
crosshair would automatically be angled at 45 degrees with respect to
this line.  And if the knot has n-fold rotational symmetry, n such
cross hairs would be placed at the vertices of a regular n-gon.  (All
this is not such a "nightmare" if it is properly considered from the
beginning.)

I don't expect that anything in the current system should be changed;
but this approach should definitely be considered in a future version
of such a package.

RE: Mismatched diagram and symmetry specifications: "This situation
cannot occur" -- If the user makes a mistake, will the current system
send a suitable error message?

RE: A hierarchical way of dealing with symmetry: "This is exactly what
I have done!  The referee gives an accurate summary of the code's
methodology."  - Place a corresponding, extended description into the
main text.

RE: A slider interface for changing some optimization parameters: It
is clear why this would not be practical in the current system, where
the optimization of a pretzel knot "may take a few weeks."  But this
also suggests that there is much room for improvement!  Perhaps some
badness functions are not really needed or are badly implemented.

RE: Link projections: In a future system, the planning for handling
links should come in right at the beginning.  It is not at all clear
why "Each component would have its own badness."

Many of the important badness functions (e.g. separations of nodes or
of non-crossing curve segments) are based purely on local geometry and
need not know what component of a link they refer to.  Perhaps the
overall length of a knot or of a link-component is not a good metric,
and should be replaced with some other functions of the length of
curve segments between crossings.

Some detailed comments on manuscript R2:

Page 6, line 20:  How many (x,y) pairs are there?

Page 6, Line 36: Why "R50"? - Are there a total of 50 parameters?
What are they?

Page 6, Line 39: "… the overall bending energy." ->> 'minimizing
overall bending energy.'

Page 6, Line 42:  What is "> b <" ?

Figures 3 and 5: Try running these figures with only 50 points (and
circles) for each Bezier segment.

Figure 4 caption: Much of the explanation of "7 3" versus "3 7" is
already in the main text.

Page 8, Line 60: "Why 64 real variables"?  What are they?  How does
this agree with "R50" on page 6 ?

Figure 7:  Where does the spiky artifact in the upper right lobe come from?

Page 12, Line 52: "...a rotationally symmetric" ->> " ... a knot with
pure cyclic rotational symmetry, rather than dihedral symmetry"

Figure 9: Why did the user choose to place three nodes on the outer
lobes, rather than only two? - Was this easier to express the desired
symmetry??

Page 16, Line 47: "The strands exert a couple but no force on one
another." ??  Lengthy discussions should be relegated to the main
text, rather than be in figure captions.

Figure 14: All the knots are nicer than the ones in Ref [24].  Good
results!

References: Several of them are incomplete and need an URL or a more
detailed description where the source can be found.

\end{document}

### 2023-01-31 inst/jma_rebuttal_2023_01_30.tex
fixes #91

@@
 
-Page 6, Line 42:  What is "> b <" ?
+Page 6, Line 42:  What is "{\tt > b <}" ?
+\\ \\
+\textcolor{blue}{This is just a detail of standard R syntax.  The
+  first ``{\tt > }'' is just the R prompt; letter {\tt b} is just a
+  variable; and the symbol ``{\tt <-}" usually has a gloss of
+  ``becomes".  Thus {\tt b <- 4} is read ``{\tt b} becomes 4". The
+  reason I need a variable is that I want to show the result of
+  applying two functions, {\tt total\_crossing\_angles()} and {\tt
+    total\_bending\_energy()}, to the same variable {\tt b}.}
 

### 2023-01-31 inst/jma_knot.tex
consistentification
fixes #90

@@
 programming language (that is, a ``package"): {\tt knotR}.  The
-package leverages the graphical capabilities of the popular vectorised
+package leverages the graphical capabilities of the popular vectorized
 graphics software {\tt inkscape}.  Different aspects of knot
@@
 Two examples of desiderata for such an objective function might be to
-keep the strands crossing at right angles, and the overall bending
+keep the strands crossing at right angles, and minimizing overall bending
 energy.  These are evaluated in the package by functions

### 2023-01-31 inst/jma_rebuttal_2023_01_30.tex
fixes #90
fixes #89

@@
 \usepackage{xcolor}
+\usepackage{amssymb}
 \begin{document}
@@
 
-Page 6, Line 36: Why "R50"? - Are there a total of 50 parameters?
-What are they?
+Page 6, Line 36: Why "$\mathbb{R}^{50}$"? - Are there a total of 50
+parameters?  What are they?  \textcolor{blue}{\\This was not clear and
+  I have rephased.  Old text: ``The optimization typically proceeds
+  over $\mathbb{R}^{50}$\ldots"`. New text: ``The optimization
+  proceeds over $\mathbb{R}^n$ where $n$ is the number of parameters
+  needed to specify the path; typically $n\simeq 50$ for the knots
+  presented here".\\ \\}
 
-Page 6, Line 39: "… the overall bending energy." ->> 'minimizing
+Page 6, Line 39: ``... the overall bending energy." {\tt ->>} 'minimizing
 overall bending energy.'
+\textcolor{blue}{\\Fixed\\ \\}
 

### 2023-01-31 inst/jma_knot.tex
fixes #89

@@
 method~\cite{dennis1983} (gradient information is not available).  The
-optimization typically proceeds over $\mathbb{R}^{50}$ and in the
-context of the package, the methods appear to be broadly comparable.
+optimization proceeds over $\mathbb{R}^n$ where $n$ is the number of
+parameters needed to specify the path; typically $n\simeq 50$ for the
+knots presented here.  In the context of the package, the methods
+appear to be broadly comparable.
 

### 2023-01-31 inst/jma_rebuttal_2023_01_30.tex
fixes #88

@@
 Page 6, line 20: How many (x,y) pairs are there?
+\textcolor{blue}{\\Object {\tt k76} has a total of 49 lines and I have
+  indicated this in the text.  Old text: ``Above we see only the first
+  six lines of the object".  New text: ``The object has 49 lines, of
+  which only the first six are shown".\\ \\}
 
-Page 6, Line 36: Why "$\mathbb{R}^{50}$"? - Are there a total of 50
+Page 6, Line 36: Why ``$\mathbb{R}^{50}$"? - Are there a total of 50
 parameters?  What are they?  \textcolor{blue}{\\This was not clear and
   I have rephased.  Old text: ``The optimization typically proceeds
-  over $\mathbb{R}^{50}$\ldots"`. New text: ``The optimization
+  over $\mathbb{R}^{50}$\ldots". New text: ``The optimization
   proceeds over $\mathbb{R}^n$ where $n$ is the number of parameters

### 2023-01-31 inst/jma_rebuttal_2023_01_30.tex
fixes #95
typo

@@
 \textcolor{blue}{This is just a detail of standard R syntax.  The
-  first ``{\tt > }'' is just the R prompt; letter {\tt b} is just a
+  first ``{\tt > }" is just the R prompt; letter {\tt b} is just a
   variable; and the symbol ``{\tt <-}" usually has a gloss of
@@
 
-Figure 7:  Where does the spiky artifact in the upper right lobe come from?
+Figure 7: Where does the spiky artifact in the upper right lobe come
+from?  \textcolor{blue}{\\I see nothing in figure 7 that could be so
+  described.  I expect this is a rendering issue, if so it will be
+  sorted out at proofreading stage.\\ \\}
 

### 2023-01-31 inst/jma_knot.tex
fixes #88
fixes #96

@@
 two-column matrix with rows corresponding to the node and handle
-positions of the inkscape path.  Above we see only the first six lines
-of the object.  Inkscape representation has a certain amount of
+positions of the inkscape path.   The object has 49 lines, of which only the  first six are shown.  Inkscape representation has a certain amount of
 redundancy as knot paths have handles which are symmetrically placed
@@
 indicates that nodes 12, 4, 16, 8, and~20 are concyclic.  An example
-of a rotationally symmetric knot is given in Figure~\ref{orn20}.
+of a knot with pure rotational symmetry, rather than dihedral
+symmetry, is given in Figure~\ref{orn20}.
 

### 2023-01-31 inst/jma_rebuttal_2023_01_30.tex
fixes #96
fixes #97

@@
 
-Page 12, Line 52: "...a rotationally symmetric" ->> " ... a knot with
+Page 12, Line 52: ``...a rotationally symmetric" ->> ``... a knot with
 pure cyclic rotational symmetry, rather than dihedral symmetry"
+  \textcolor{blue}{Done\\ \\}
 
@@
 lobes, rather than only two? - Was this easier to express the desired
-symmetry??
+symmetry??  \textcolor{blue}{Sort of.  Having more nodes allows for
+  finer and more sophisticated control over the path, and having too
+  few nodes results in a displeasing, unnatural flabby configuration.
+  On the other hand, having too many nodes results in the optimization
+  routine thrashing and failing to converge.  In this case it is
+  desirable to have an odd number of nodes: 12,4,16,8,20 are
+  constrained by the symmetry routine to be exactly on one of the five
+  lines of symmetry, and to intersect its line at exactly $90^\circ$.
+  Frankly the choice of number of nodes has been troublesome in the
+  development of the software.  Perhaps this issue is of interest to
+  the JMA readership and I have added a brief sentence discussing it
+  to the manuscript: ``Note that increasing the number of nodes
+  provides more control over the path at the expense of a more
+  difficult optimization process".  Also, at the discussion of Figure
+  9: ``from Figure 9 we see that these nodes are constrained to lie on
+  one of the five lines of symmetry, cutting it at $90^\circ$".}
 

### 2023-01-31 inst/jma_knot.tex
fixes #97

@@
 is stored); or {\tt controlpoints} objects, which allow for easy
-construction of Bezier interpolation between nodes.
+construction of Bezier interpolation between nodes.  Note that increasing the number of nodes provides more control over the path at the expense of a more difficult optimization process.
 
@@
 map to themselves under rotation.  The top line of {\tt Mrot}
-indicates that nodes 12, 4, 16, 8, and~20 are concyclic.  An example
-of a knot with pure rotational symmetry, rather than dihedral
-symmetry, is given in Figure~\ref{orn20}.
+indicates that nodes 12, 4, 16, 8, and~20 are concyclic and indeed
+from Figure~\ref{k5_1_twoknots} we see that these nodes are
+constrained to lie on one of the five lines of symmetry, cutting it at
+$90^\circ$.  An example of a knot with pure rotational symmetry,
+rather than dihedral symmetry, is given in Figure~\ref{orn20}.
 

### 2023-01-31 inst/jma_rebuttal_2023_01_30.tex
typos

@@
 
-Page 12, Line 52: ``...a rotationally symmetric" ->> ``... a knot with
+Page 12, Line 52: ``...a rotationally symmetric" should be ``... a knot with
 pure cyclic rotational symmetry, rather than dihedral symmetry"
@@
 lobes, rather than only two? - Was this easier to express the desired
-symmetry??  \textcolor{blue}{Sort of.  Having more nodes allows for
+symmetry??  \textcolor{blue}{\\Sort of.  Having more nodes allows for
   finer and more sophisticated control over the path, and having too
@@
   9: ``from Figure 9 we see that these nodes are constrained to lie on
-  one of the five lines of symmetry, cutting it at $90^\circ$".}
+  one of the five lines of symmetry, cutting it at $90^\circ$".\\ \\}
 
 Page 16, Line 47: "The strands exert a couple but no force on one
-another." ??  Lengthy discussions should be relegated to the main
+another."    Lengthy discussions should be relegated to the main
 text, rather than be in figure captions.

### 2023-01-31 inst/jma_rebuttal_2023_01_30.tex
fixes #98

@@
 
-Page 16, Line 47: "The strands exert a couple but no force on one
-another."    Lengthy discussions should be relegated to the main
-text, rather than be in figure captions.
+Page 16, Line 47: ``The strands exert a couple but no force on one
+another."  Lengthy discussions should be relegated to the main text,
+rather than be in figure captions \textcolor{blue}{\\Done.  This
+  discussion is relocated to the main text as suggested.\\ \\}
 

### 2023-01-31 inst/jma_knot.tex
fixes #98

@@
 Figure~\ref{two_trefoil_knots} shows and compares two plausible
-schemes for rendering a trefoil.  Figure~\ref{perko_AB} shows two
+schemes for rendering a trefoil: optimized as per the {\tt knotR}
+package techniques documented here; compared with a trefoil knot
+constructed from arcs of circles constrained to cross at right angles.
+In the circular-arc trefoil, the strands exert a couple but no force
+on one another.  Mirror symmetry and threefold rotational symmetry are
+imposed.  Elementary trigonometry shows that the radii of the inner
+and outer loops are $\operatorname{cosec}(\pi/12)=\sqrt{6}+\sqrt{2}$
+and $\sec(\pi/12)=\sqrt{6}-\sqrt{2}$ respectively.  We see the
+disadvantages of placing nodes at crossing points.
+
+Figure~\ref{perko_AB} shows two
 forms of a 10-crossing knot once erroneously considered to be
@@
       here; right, constructed from arcs of circles constrained to
-      cross at right angles.  The strands exert a couple but no force
-      on one another.  Mirror symmetry and threefold rotational
-      symmetry are imposed.  Elementary trigonometry shows that the
-      radii of the inner and outer loops are
-      $\operatorname{cosec}(\pi/12)=\sqrt{6}+\sqrt{2}$ and
-      $\sec(\pi/12)=\sqrt{6}-\sqrt{2}$ respectively.  The curvature
-      thus has a displeasing discontinuity at a node which interrupts
+      cross at right angles.    On the right, the curvature
+      has a displeasing discontinuity at a node which interrupts
       the visual continuity there, especially the understrand}

### 2023-01-31 inst/jma_knot.tex
fixes #93

@@
   the sense of the crossings can be established.  For example, strand
-  7 passes over strand 3.  In the package, this fact is specified by
-  the {\tt overunder} object, as discussed in the text, having a row
-  {\tt 7 3} (and not {\tt 3 7} which would be the crossing having the
-  opposite sense).  Crossings become relevant only for the final
+  7 should pass over strand 3 as per the {\tt overunder}
+  specification.  Crossings become relevant only for the final
   rendering of the knot curve with the gaps included}

### 2023-01-31 inst/jma_rebuttal_2023_01_30.tex
fixes #93
fixes #82

@@
 G2-continuous B-splines, which are only approximating their control
-points.
+points.\textcolor{blue}{\\Fixed: sentence added to manuscript\\ \\}
 
@@
 
-Page 6, Line 42:  What is "{\tt > b <}" ?
-\\ \\
-\textcolor{blue}{This is just a detail of standard R syntax.  The
-  first ``{\tt > }" is just the R prompt; letter {\tt b} is just a
-  variable; and the symbol ``{\tt <-}" usually has a gloss of
-  ``becomes".  Thus {\tt b <- 4} is read ``{\tt b} becomes 4". The
-  reason I need a variable is that I want to show the result of
-  applying two functions, {\tt total\_crossing\_angles()} and {\tt
-    total\_bending\_energy()}, to the same variable {\tt b}.}
+Page 6, Line 42: What is "{\tt > b <}" ?  \textcolor{blue}{\\ This is
+  just a detail of standard R syntax.  The first ``{\tt > }" is just
+  the R prompt; letter {\tt b} is just a variable; and the symbol
+  ``{\tt <-}" usually has a gloss of ``becomes".  Thus {\tt b <- 4} is
+  read ``{\tt b} becomes 4". The reason I need a variable is that I
+  want to show the result of applying two functions, {\tt
+    total\_crossing\_angles()} and {\tt total\_bending\_energy()}, to
+  the same variable {\tt b}.\\ \\}
 
@@
 Figure 4 caption: Much of the explanation of "7 3" versus "3 7" is
-already in the main text.
+already in the main text.  \textcolor{blue}{\\Fixed\\ \\}
 

### 2023-01-31 inst/jma_rebuttal_2023_01_30.tex
fixes #84

@@
 
-RE: Mismatched diagram and symmetry specifications: "This situation
+RE: Mismatched diagram and symmetry specifications: ``This situation
 cannot occur" -- If the user makes a mistake, will the current system
-send a suitable error message?
+send a suitable error message?\textcolor{blue}{\\For convenience, I
+  reproduce the original exchange below:\\ \\} {\tiny Reviewer's
+  comment, round 3: ``If all nodes fulfill the specified symmetry
+  constraints (as shown on page 11, lines 15-22 and on lines 35-43),
+  but the connection diagram does not completely adhere to the
+  specified symmetry - What happens? - Does the package check this and
+  let the user know about the discrepancy?"}  \textcolor{blue}{\tiny
+  My original reply: ``This situation cannot occur: symmetric nodes
+  imply a symmetric diagram".\\ \\} \textcolor{blue}{A more nuanced
+  reply would be to observe that in the event of a mismatch between
+  diagram and symmetry specifications, the symmetry specifications
+  take priority and the resulting symmetrised diagram automatically
+  accommodates the imposed symmetry.  In practice what happens is
+  immediately visually apparent: one sees beautiful symmetric nodes
+  joined by Bezier curves that do not posess the intended topology.
+  The diagram is a disorganised ugly tangled mess, clearly not
+  reflecting the intended configuration.  In practice, one glance at
+  the diagram simultaneously reveals the problem and its solution
+  (just reconnect the dots in the correct order).\\ \\}
 

### 2023-01-31 inst/jma_rebuttal_2023_01_30.tex
fixes #86

@@
 is clear why this would not be practical in the current system, where
-the optimization of a pretzel knot "may take a few weeks."  But this
+the optimization of a pretzel knot ``may take a few weeks."  But this
 also suggests that there is much room for improvement!  Perhaps some
 badness functions are not really needed or are badly implemented.
+\textcolor{blue}{\\ This is an excellent suggestion and indeed many of
+  the badness functions may well turn out to be superfluous, or at
+  least susceptible to faster heuristics [although I would argue that
+    the functions all adhere to efficient vectorised R coding
+    standards].  If it turns out that some of the badness weightings
+  can be set to zero without degrading the resulting visual
+  attractiveness, this would certainly be worth pursuing.\\ \\}
 

### 2023-02-01 inst/jma_rebuttal_2023_01_30.tex
further points for issue #100
fixes #94 also refers issue #89

@@
   proceeds over $\mathbb{R}^n$ where $n$ is the number of parameters
-  needed to specify the path; typically $n\simeq 50$ for the knots
+  needed to specify the path; typically $n\gtrsim 50$ for the knots
   presented here".\\ \\}
@@
 Page 8, Line 60: "Why 64 real variables"?  What are they?  How does
-this agree with "R50" on page 6 ?
+this agree with "R50" on page 6?\textcolor{blue}{\\ The 64 variables are just the parameters required to describe the knot path, it agrees with the (reworded) coment about $\mathbb{R}^{50}$.\\ \\}
 
@@
 detailed description where the source can be found.
+\textcolor{blue}{\\Done. \\ \\}
 

### 2023-02-01 inst/jma_knot.tex
fixes #82
fixes #100
fixes #94 also refers issue #89

@@
 \usepackage{amsmath}
+\usepackage{url}
 %
@@
 specify the path of a knot projection; here cubic Bezier curves are
-used.  Bezier curves are a natural choice for working with
+used.  This gives users a convenient, direct control of where these
+curve segments will appear; compare G2-continuous B-splines, which
+only approximate their control points.
+
+Bezier curves are a natural choice for working with
 mathematical knots for several reasons: they are familiar to many
@@
 optimization proceeds over $\mathbb{R}^n$ where $n$ is the number of
-parameters needed to specify the path; typically $n\simeq 50$ for the
+parameters needed to specify the path; typically $n\gtrsim 50$ for the
 knots presented here.  In the context of the package, the methods

### 2023-02-01 inst/jma_rebuttal_2023_01_30.tex
fixes #85

@@
 
-RE: A hierarchical way of dealing with symmetry: "This is exactly what
-I have done!  The referee gives an accurate summary of the code's
+RE: A hierarchical way of dealing with symmetry: ``This is exactly
+what I have done!  The referee gives an accurate summary of the code's
 methodology." - Place a corresponding, extended description into the
-main text.
+main text.\textcolor{blue}{\\Done.  New text added in section 3.2:
+  From a numerical perspective, the software takes advantage of a
+  hierarchical representation of symmetrical knots: the curve is
+  specified for one unique fragment of the knot, and the complete knot
+  is reconstructed by replicating that fragment with mirroring and
+  rotation as required.  The optimization is effectively carried out
+  for this unique fragment (with proper end-conditions).\\ \\}
 

### 2023-02-01 inst/jma_knot.tex
fixes #85

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{url}
@@
 
+From a numerical perspective, the software takes advantage of a
+hierarchical representation of symmetrical knots: the curve is
+specified for one unique fragment of the knot, and the complete knot
+is reconstructed by replicating that fragment with mirroring and
+rotation as required.  The optimization is effectively carried out for
+this unique fragment (with proper end-conditions).
+
 \subsection{Subjective \label{section_cca} choice for weighting}

### 2023-02-01 inst/jma_rebuttal_2023_01_30.tex
fixes #92
fixes #99
fixes #83

@@
 \usepackage{amssymb}
+\usepackage{graphicx}
 \begin{document}
@@
 of such a package.
+\textcolor{blue}{\\ This scheme certainly sounds achievable and I have
+  alluded to it in the manuscript at the two trefoil knot
+  discussion.\\ \\}
 
@@
 Figures 3 and 5: Try running these figures with only 50 points (and
-circles) for each Bezier segment.
+circles) for each Bezier segment. \textcolor{blue}{\\This is easily
+  achievable in the package but the result fails to provide any
+  informative or enlightening information.  It simply highlights the
+  fact that $n$ was too small.  I show the requested figure below}
+
+\includegraphics[width=13cm]{k7_6_n05.pdf}
+% created by knotplot2(k7_6,n=7)
 
@@
 Figure 14: All the knots are nicer than the ones in Ref [24].  Good
-results!
+results!\textcolor{blue}{\\Thank you!}
 

### 2023-02-01 inst/jma_rebuttal_2023_01_30.tex
further work on issue #92

@@
 \usepackage{graphicx}
+\usepackage{fancyvrb}
 \begin{document}
@@
   informative or enlightening information.  It simply highlights the
-  fact that $n$ was too small.  I show the requested figure below}
+  fact that $n$ was too small.  Using $n=50$ is (almost)
+  indistinguishable from $n=100$.  I found that $n=100$ represented
+  essentially the best visual representation while being
+  computationally not too onerous.  I show the requested figure at the
+  end (along with $n=7$ and $n=100$ for comparison),
+  figures~\ref{7_6_n7}, \ref{7_6_n50}, and \ref{7_6_n100}.  These were
+  created with}
 
-\includegraphics[width=13cm]{k7_6_n05.pdf}
-% created by knotplot2(k7_6,n=7)
+\begin{Verbatim}[commandchars=\\\{\}]
+\textcolor{blue}{R> pdf(file="k7_6_n07.pdf") ; knotplot2(k7_6,n=7)   ; dev.off()}
+\end{Verbatim}
+\begin{Verbatim}[commandchars=\\\{\}]
+\textcolor{blue}{R> pdf(file="k7_6_n50.pdf") ; knotplot2(k7_6,n=50)  ; dev.off()}
+\end{Verbatim}
+\begin{Verbatim}[commandchars=\\\{\}]
+\textcolor{blue}{R> pdf(file="k7_6_n100.pdf") ; knotplot2(k7_6,n=100) ; dev.off()}
+\end{Verbatim}
 
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=10cm]{"k7_6_n07.pdf"}
+  \caption{Knot \label{7_6_n7} $7_6$ with $n=7$ nodes per segment}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=10cm]{"k7_6_n50.pdf"}
+  \caption{Knot \label{7_6_n50} $7_6$ with $n=50$ nodes per segment}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=10cm]{"k7_6_n100.pdf"}
+  \caption{Knot \label{7_6_n100} $7_6$ with $n=100$ nodes per segment}
+\end{figure}
+
 \end{document}

### 2023-02-01 inst/jma_knot.tex
issue #81

@@
 The knots supplied in the package minimize a weighted sum of these and
-other badnesses\footnote{Function {\tt badness()} includes various
+other badnesses.  Function {\tt badness()} includes various
 ``housekeeping'' badnesses which are used to make sure that the
 minimum found by {\tt nlm()} is topologically identical to the
-starting configuration (a list is given in the Appendix).  Function
-{\tt non\_crossing\_strand\_close\_approach\_badness()}, for example,
+starting configuration.  Function {\tt
+  non\_crossing\_strand\_close\_approach\_badness()}, for example,
 makes non-crossing strands ``repel'' one another so as not to
-introduce spurious intersections.}.  The weightings for the various
-badnesses are, of course, subjective; but the system discussed here
-allows the user to change the weightings used and compare results.  I
-present a short discussion in section~\ref{section_cca} below.
-Numerically, the badnesses are evaluated by function {\tt badness()}:
+introduce spurious intersections.
+
+There follows a more detailed description and motivation for various
+metrics used to evaluate knot diagrams in the package.
+
+\begin{description}
+\item[{\tt total\_crossing\_potential\_energy()}] The idea is to make
+  the crossing points attract one another as if they are
+  gravitationally active objects; metric measured by their mutual
+  potential energy.  If crossing point $i$ is at position ${\mathbf
+    p}_i$ we have $\sum_{i\neq j}\left|{\mathbf p}_i-{\mathbf
+    p}_j\right|^{2}$.  The idea is to encourage more compact
+  configurations rather than long skinny ones.
+\item[{\tt node\_crossing\_badness()}] Here we make the crossing
+  points {\em repel} one another, as though they are all charged
+  objects (with the same sign).  The idea is to spread out the nodes
+  so no two nodes are too close together.  If crossing point $i$ is at
+  position ${\mathbf p}_i$ we have $\sum_{i\neq j}\left|{\mathbf
+    p}_i-{\mathbf p}_j\right|^{-1}$.  The length scale for this
+  repulsion is smaller than that of the attractive term above.  The
+  overall potential thus resembles that sometimes seen in atomic
+  physics when considering the hydrogen molecule $H_2$: the two
+  protons attract one another when distant, but repel when closer.  In
+  the package, the intent is for this optimal distance to be a shallow
+  minimum and the optimization routine is free to adjust within a wide
+  range.
+\item[{\tt total\_crossing\_angles()}] Here we wish to punish
+  configurations with strands that cross at acute angles.  If crossing
+  point $i$ has crossing angle $\theta_i$ the metric is
+  $\sum_i\left(\cos\theta_i\right)^2$.
+\item[{\tt total\_bending\_energy()}] The idea is for the strands to
+  preferentially follow elastica curves, as in elementary (nonlinear)
+  beam theory.  From differential geometry, the Euler elastica
+  potential energy is $\int\kappa(s)^2\,ds$.  In the package, this is
+  calculated numerically for each strand and the total returned.
+\item[{\tt total\_string\_length()}] The length is just $\int ds$, with a
+  nominal value of 5000 (which value is about right for a printed
+  copy).
+\item[{\tt midpoint\_badness()}] The idea is to suppress strands with
+  crossing points too close to the endpoint nodes.  For each crossing
+  point $c_{ij}$ of strand $i$ with strand $j$, we calculate the arc
+  length $\int_{s=0}^{c_{ij}}\,ds$ from the node to each crossing
+  point and normalize with the strand length $\int\,ds$.  We reward
+  configurations with normalized crossing point arc lengths that are
+  close to $1/2$.  Note that this particular metric is not independent
+  of the total crossing potential energy metric.
+\item[{\tt curvature\_switching\_badness()}] Here the problem is knots
+  which should have either all their strands curving to the left, or
+  all to the right; the signed curvature should be everywhere the same
+  sign (the canonical example would be $8_{18}$ which in earlier
+  versions of the software had ugly changes of curvature).  We impose
+  a fixed penalty for any knot with more than one sign.
+\item[{\tt curvature\_consecutive\_segment\_switching\_badness()}] We
+  wish to discourage configurations in which curvature undergoes a
+  sudden change at a node.  The value for a single (Bezier) node is
+  simply $\left(\delta\kappa\right)^2$, and the penalty for a given
+  configuration is the sum of the values for each node.  See
+  Figure~\ref{two_trefoil_knots} for motivation and discussion.
+\item[{\tt non\_crossing\_strand\_close\_approach\_badness()}] Here
+  the idea is to prevent non-intersecting strands from approaching one
+  another too closely.  Quite apart from aesthetic considerations,
+  this metric prevents the creation of new spurious nodes that would
+  disrupt the optimization routines.
+\end{description}
+
+\noindent The weightings for the various badnesses are, of course,
+subjective; but the system discussed here allows the user to change
+the weightings used and compare results.  I present a short discussion
+in section~\ref{section_cca} below.  Numerically, the badnesses are
+evaluated by function {\tt badness()}:
 
@@
 
-Further work might include adding functionality to deal with links,
-although this is not straightforward, and a brief discussion is given
-in Appendix B.
+Further work might include functionality to deal with links.  Much of
+the required functionality is (in principle) already present in the
+software.  Each component would have its own badness, and in addition
+there would be $n(n-1)/2$ inter-component interaction terms measuring
+features such as closeness between non-intersecting strands of
+different components.  However, the implementation of such
+functionality is not straightforward and several problems stand out.
+Firstly, dealing with inter-component intersection points is
+difficult.  Currently, crossing points are represented as a
+(symmetrical) Boolean matrix with rows and columns corresponding to
+strands, and entry $(i,j)$ being whether strand $i$ intersects with
+strand $j$.  The link generalization would be a symmetric matrix with
+entries that are themselves intersection matrices: an object with four
+indices and entry $(i,j,k,l)$ indicating whether strand $j$ of
+component $i$ intersects strand $l$ of component $k$.  Such objects
+are unwieldy to work with and are typically difficult to manipulate.
+Secondly, the imposition of symmetry is not straightforward in the
+case of links: one might desire that certain components of a link have
+a particular kind of symmetry and others to have a different kind (or
+no!) symmetry.  Further, one unexpectedly difficult problem was
+dealing with a subset of components that individually possess no
+symmetry but collectively possess mirror (or rotational) symmetry.
+Even the simplest links, as in Figure~\ref{thistlethwaite}, can be
+problematic.  Consider Thistlethwaite's L4a1, for example: one would
+expect the two components to be identical except for a 90-degree
+rotation, and implementing this in the context of the package does not
+seem to be at all easy.  In general, new badnesses would be needed.
+Further, in L7a1, the smaller component should be not only as circular
+as possible (?), but also one might desire the overall length to be
+smaller than that of the other component.  It would be reasonable to
+conclude that dealing with links might be possible in principle but,
+due to a number of technical and mathematical reasons, considerably
+harder than the single-component links considered here.
+
+\begin{figure}[htbp]
+  \begin{center}
+\includegraphics[width=7cm]{L4a1_L7a1}  % made by exporting L4a1_L7a1.svg as PDF
+\caption{Two links \label{thistlethwaite} taken from the
+  Thistlethwaite link table~\cite{hoste1998}.  These are produced
+  directly with inkscape and are not optimized}
+  \end{center}
+\end{figure}
 
@@
 
-\clearpage
-\section*{Appendix A: details and motivation for the badness functions}
-
-There follows a more detailed description and motivation for various
-metrics used to evaluate knot diagrams in the package.
-
-\begin{description}
-\item[{\tt total\_crossing\_potential\_energy()}] The idea is to make
-  the crossing points attract one another as if they are
-  gravitationally active objects; metric measured by their mutual
-  potential energy.  If crossing point $i$ is at position ${\mathbf
-    p}_i$ we have $\sum_{i\neq j}\left|{\mathbf p}_i-{\mathbf
-    p}_j\right|^{2}$.  The idea is to encourage more compact
-  configurations rather than long skinny ones.
-\item[{\tt node\_crossing\_badness()}] Here we make the crossing
-  points {\em repel} one another, as though they are all charged
-  objects (with the same sign).  The idea is to spread out the nodes
-  so no two nodes are too close together.  If crossing point $i$ is at
-  position ${\mathbf p}_i$ we have $\sum_{i\neq j}\left|{\mathbf
-    p}_i-{\mathbf p}_j\right|^{-1}$.  The length scale for this
-  repulsion is smaller than that of the attractive term above.  The
-  overall potential thus resembles that sometimes seen in atomic
-  physics when considering the hydrogen molecule $H_2$: the two
-  protons attract one another when distant, but repel when closer.  In
-  the package, the intent is for this optimal distance to be a shallow
-  minimum and the optimization routine is free to adjust within a wide
-  range.
-\item[{\tt total\_crossing\_angles()}] Here we wish to punish
-  configurations with strands that cross at acute angles.  If crossing
-  point $i$ has crossing angle $\theta_i$ the metric is
-  $\sum_i\left(\cos\theta_i\right)^2$.
-\item[{\tt total\_bending\_energy()}] The idea is for the strands to
-  preferentially follow elastica curves, as in elementary (nonlinear)
-  beam theory.  From differential geometry, the Euler elastica
-  potential energy is $\int\kappa(s)^2\,ds$.  In the package, this is
-  calculated numerically for each strand and the total returned.
-\item[{\tt total\_string\_length()}] The length is just $\int ds$, with a
-  nominal value of 5000 (which value is about right for a printed
-  copy).
-\item[{\tt midpoint\_badness()}] The idea is to suppress strands with
-  crossing points too close to the endpoint nodes.  For each crossing
-  point $c_{ij}$ of strand $i$ with strand $j$, we calculate the arc
-  length $\int_{s=0}^{c_{ij}}\,ds$ from the node to each crossing
-  point and normalize with the strand length $\int\,ds$.  We reward
-  configurations with normalized crossing point arc lengths that are
-  close to $1/2$.  Note that this particular metric is not independent
-  of the total crossing potential energy metric.
-\item[{\tt curvature\_switching\_badness()}] Here the problem is knots
-  which should have either all their strands curving to the left, or
-  all to the right; the signed curvature should be everywhere the same
-  sign (the canonical example would be $8_{18}$ which in earlier
-  versions of the software had ugly changes of curvature).  We impose
-  a fixed penalty for any knot with more than one sign.
-\item[{\tt curvature\_consecutive\_segment\_switching\_badness()}] We
-  wish to discourage configurations in which curvature undergoes a
-  sudden change at a node.  The value for a single (Bezier) node is
-  simply $\left(\delta\kappa\right)^2$, and the penalty for a given
-  configuration is the sum of the values for each node.  See
-  Figure~\ref{two_trefoil_knots} for motivation and discussion.
-\item[{\tt non\_crossing\_strand\_close\_approach\_badness()}] Here
-  the idea is to prevent non-intersecting strands from approaching one
-  another too closely.  Quite apart from aesthetic considerations,
-  this metric prevents the creation of new spurious nodes that would
-  disrupt the optimization routines.
-\end{description}
-
-\section*{Appendix B: some reflections on links}
-
-Further work might include functionality to deal with links.  Much of
-the required functionality is (in principle) already present in the
-software.  Each component would have its own badness, and in addition
-there would be $n(n-1)/2$ inter-component interaction terms measuring
-features such as closeness between non-intersecting strands of
-different components.  However, the implementation of such
-functionality is not straightforward and several problems stand out.
-Firstly, dealing with inter-component intersection points is
-difficult.  Currently, crossing points are represented as a
-(symmetrical) Boolean matrix with rows and columns corresponding to
-strands, and entry $(i,j)$ being whether strand $i$ intersects with
-strand $j$.  The link generalization would be a symmetric matrix with
-entries that are themselves intersection matrices: an object with four
-indices and entry $(i,j,k,l)$ indicating whether strand $j$ of
-component $i$ intersects strand $l$ of component $k$.  Such objects
-are unwieldy to work with and are typically difficult to manipulate.
-Secondly, the imposition of symmetry is not straightforward in the
-case of links: one might desire that certain components of a link have
-a particular kind of symmetry and others to have a different kind (or
-no!) symmetry.  Further, one unexpectedly difficult problem was
-dealing with a subset of components that individually possess no
-symmetry but collectively possess mirror (or rotational) symmetry.
-Even the simplest links, as in Figure~\ref{thistlethwaite}, can be
-problematic.  Consider Thistlethwaite's L4a1, for example: one would
-expect the two components to be identical except for a 90-degree
-rotation, and implementing this in the context of the package does not
-seem to be at all easy.  In general, new badnesses would be needed.
-Further, in L7a1, the smaller component should be not only as circular
-as possible (?), but also one might desire the overall length to be
-smaller than that of the other component.  It would be reasonable to
-conclude that dealing with links might be possible in principle but,
-due to a number of technical and mathematical reasons, considerably
-harder than the single-component links considered here.
-
-\begin{figure}[htbp]
-  \begin{center}
-\includegraphics[width=7cm]{L4a1_L7a1}  % made by exporting L4a1_L7a1.svg as PDF
-\caption{Two links \label{thistlethwaite} taken from the
-  Thistlethwaite link table~\cite{hoste1998}.  These are produced
-  directly with inkscape and are not optimized}
-  \end{center}
-\end{figure}
-
 \end{document}

### 2023-02-02 inst/jma_rebuttal_2023_01_30.tex
issue #81
fixes #87

@@
 directly to what I see as the main contribution of a JMA paper on this
-topic.
+topic.\textcolor{blue}{Done.}
 
@@
   immediately visually apparent: one sees beautiful symmetric nodes
-  joined by Bezier curves that do not posess the intended topology.
-  The diagram is a disorganised ugly tangled mess, clearly not
+  joined by Bezier curves that do not possess the intended topology.
+  The diagram is a disorganised and ugly tangled mess, clearly not
   reflecting the intended configuration.  In practice, one glance at
@@
 and should be replaced with some other functions of the length of
-curve segments between crossings.
+curve segments between crossings.  \textcolor{blue}{I have spent many
+  hours pondering this comment, in particular my original assertion
+  that each component would have its own badness.  I took this to be
+  axiomatic, but on [extensive] reflection the referee makes an
+  excellent point, specifically that the badness metrics are all {\em
+    local}.  This is a profound observation, one that eluded me over
+  several years of developing and using the software and thinking
+  about link diagrams.  I think I missed it because the badness
+  functions are largely geared towards preserving the {\em global}
+  topology of the diagram [including preventing Reidemeister moves]
+  and it seemed natural to consider a link diagram from a topological
+  perspective, that is, to consider the badness of each link
+  separately.  But the referee's comment is infuriatingly obvious once
+  it has been pointed out.  Much of the appendix B is irrelevant in
+  the light of this observation; I have shortened it considerably and
+  rephrased to reflect the new thinking.  One suggestion above was to
+  incorporate appendices A and B in to the main body of the text,
+  which I have done, but to accommodate this comment I have
+  restructured the text (of what was appendix B) to reflect this
+  insight.\\ \\}
 

### 2023-02-02 inst/jma_knot.tex
fixes #87

@@
 the required functionality is (in principle) already present in the
-software.  Each component would have its own badness, and in addition
-there would be $n(n-1)/2$ inter-component interaction terms measuring
-features such as closeness between non-intersecting strands of
-different components.  However, the implementation of such
-functionality is not straightforward and several problems stand out.
-Firstly, dealing with inter-component intersection points is
-difficult.  Currently, crossing points are represented as a
-(symmetrical) Boolean matrix with rows and columns corresponding to
-strands, and entry $(i,j)$ being whether strand $i$ intersects with
-strand $j$.  The link generalization would be a symmetric matrix with
-entries that are themselves intersection matrices: an object with four
-indices and entry $(i,j,k,l)$ indicating whether strand $j$ of
-component $i$ intersects strand $l$ of component $k$.  Such objects
-are unwieldy to work with and are typically difficult to manipulate.
-Secondly, the imposition of symmetry is not straightforward in the
-case of links: one might desire that certain components of a link have
-a particular kind of symmetry and others to have a different kind (or
-no!) symmetry.  Further, one unexpectedly difficult problem was
-dealing with a subset of components that individually possess no
-symmetry but collectively possess mirror (or rotational) symmetry.
-Even the simplest links, as in Figure~\ref{thistlethwaite}, can be
-problematic.  Consider Thistlethwaite's L4a1, for example: one would
-expect the two components to be identical except for a 90-degree
-rotation, and implementing this in the context of the package does not
-seem to be at all easy.  In general, new badnesses would be needed.
-Further, in L7a1, the smaller component should be not only as circular
-as possible (?), but also one might desire the overall length to be
-smaller than that of the other component.  It would be reasonable to
-conclude that dealing with links might be possible in principle but,
-due to a number of technical and mathematical reasons, considerably
-harder than the single-component links considered here.
+software.  Although it might be tempting to assign each link component
+its own badness, and to combine these in some way along with some form
+of inter-component interaction term, a perceptive referee observed
+that the badness components are all based purely on {\em local}
+geometry.  As such, many of the software and mathematical principles
+designed for knots will apply similarly to links with little
+modification.  Consider Thistlethwaite's L4a1
+(Figure~\ref{thistlethwaite}), for example: one would expect the two
+components to be identical except for a 90-degree rotation, and
+implementing this in the context of the package is straightforward in
+principle: the symmetry object does not depend on the overall linking
+topology of the diagram.  However, dealing with a subset of components
+that individually or collectively possess mirror (or rotational)
+symmetry, can be problematic.  Consider L7a1, in which only the
+smaller component should possess circular symmetry; such restrictions
+would require substantial work to implement.
 
@@
 
-\clearpage
 \section{Gallery}

### 2023-02-02 inst/jma_rebuttal_2023_01_30.tex
typos

@@
 
-For each crossing point, the user would see an orthogonal "cross-hair"
+For each crossing point, the user would see an orthogonal ``cross-hair"
 that can be shifted and rotated as a whole.  In addition, there are
-two parameters (handle-pairs) that control the "velocities" of the two
+two parameters (handle-pairs) that control the ``velocities" of the two
 strands through the crossing.  The optimization routine would
@@
 cross hairs would be placed at the vertices of a regular n-gon.  (All
-this is not such a "nightmare" if it is properly considered from the
+this is not such a ``nightmare" if it is properly considered from the
 beginning.)

### 2023-02-02 inst/jma_knot.tex
start work on #79

@@
 
-\begin{itemize}
-\item Curvature to be as smoothly changing as possible, with limited maximal curvature
-\item Strands to cross at right angles
-\item Crossing points to be separate from one another
-\item Any symmetry desired in the knot should be enforced exactly, and be visually apparent
-\end{itemize}
+\begin{description}
+\item[Path clarity] The projection should clearly show the path of the
+  knot, with no confusion or ambiguity in the visual representation.
+\item[Visual continuity] Strands to be as smoothly changing as
+  possible, with limited maximal curvature and no distracting features
+\item[Crossing identification] Strands to cross at right angles if
+  possible, as this admits a precise visual location and ready
+  differentiation between overstrand and understrand
+\item[Crossing spacing] The distance between crossings should be
+  visually balanced to create a harmonious projection
+\item[Symmetry] Any symmetry present in the knot (either rotational or
+  mirror) should be enforced exactly, and be visually apparent
+\end{description}
 

### 2023-02-02 inst/jma_rebuttal_2023_01_30.tex
typos

@@
 The content that is most fitting for a JMA paper is a discussion of
-what makes a knot diagram "beautiful"  and that discussion will
+what makes a knot diagram ``beautiful" and that discussion will
 clearly be biased by the personal view of the author (perhaps enhanced
-by a "user-study" among family and colleagues).  Thus lines 14 - 20 on
-page 4 are of prime importance and could be fleshed out more.
+by a ``user-study" among family and colleagues).  Thus lines 14 - 20
+on page 4 are of prime importance and could be fleshed out more.
 
@@
 directly to what I see as the main contribution of a JMA paper on this
-topic.\textcolor{blue}{Done.}
+topic.\textcolor{blue}{\\Done.\\ \\}
 

### 2023-02-03 inst/jma_rebuttal_2023_01_30.tex
make it clear that it is my reply that is more nuanced, issue #82

@@
   imply a symmetric diagram".\\ \\} \textcolor{blue}{A more nuanced
-  reply would be to observe that in the event of a mismatch between
-  diagram and symmetry specifications, the symmetry specifications
-  take priority and the resulting symmetrised diagram automatically
-  accommodates the imposed symmetry.  In practice what happens is
-  immediately visually apparent: one sees beautiful symmetric nodes
-  joined by Bezier curves that do not possess the intended topology.
-  The diagram is a disorganised and ugly tangled mess, clearly not
-  reflecting the intended configuration.  In practice, one glance at
-  the diagram simultaneously reveals the problem and its solution
-  (just reconnect the dots in the correct order).\\ \\}
+  reply would have been for me to observe that in the event of a
+  mismatch between diagram and symmetry specifications, the symmetry
+  specifications take priority and the resulting symmetrised diagram
+  automatically accommodates the imposed symmetry.  In practice what
+  happens is immediately visually apparent: one sees beautiful
+  symmetric nodes joined by Bezier curves that do not possess the
+  intended topology.  The diagram is a disorganised and ugly tangled
+  mess, clearly not reflecting the intended configuration.  In
+  practice, one glance at the diagram simultaneously reveals the
+  problem and its solution (just reconnect the dots in the correct
+  order).\\ \\}
 

### 2023-02-05 inst/jma_rebuttal_2023_01_30.tex
final closure of issue #81

@@
 tutorial on how to use package [27] to draw optimized knot diagrams,
-but JMA is not the right place to publish such a tutorial.
+but JMA is not the right place to publish such a
+tutorial.\textcolor{blue}{I have expanded the discussion of the
+  aesthetics of knot diagrams, but would point out that capturing the
+  goals}
 
@@
 future package ( -- which can also handle links).
+\textcolor{blue}{\\I have added discussions as suggested and also
+  included a modified analysis of functionality for links.\\ \\}
 

### 2023-02-05 inst/jma_rebuttal_2023_01_30.tex
fixes #78

--- 2023-02-03 inst/jma_rebuttal_2023_01_30.tex
@@
 
+\section*{Rebuttal to ``Visually pleasing knot projections" by
+  Robin K. S. Hankin (submitted to JMA, manuscript 209604076); review
+  30 January 2023}
+
 (Below, the reviewer's comments are in black, and my replies to the
@@
 
-\section*{Rebuttal to "Visually pleasing knot projections" by
-  Robin K. S. Hankin (submitted to JMA, manuscript 209604076); review
-  30 January 2023}
+\section*{\textcolor{blue}{Overall response to the review}}
 
-\textcolor{blue}{ Short story: }
+\textcolor{blue}{The comments have resulted in a much stronger
+  document.  I have carried out all the detailed suggestions, as
+  presented below.  My take on the comment reading ``much of this
+  material is perfect for an on-line tutorial" requires a more
+  subjective and personal response.  I would characterise the
+  submission, not as a online tutorial, but as a discursive yet
+  mathematical outlining of the conceptual and computational
+  difficulties of automated beautification of knot diagrams (not
+  forgetting to discuss the advantages!).  Automated beautification of
+  knot diagrams sounds easy, and I originally thought it would take me
+  a couple of hours.  I cannot be the first person to try to do this;
+  much of the submission could be viewed as documenting the
+  difficulties of the only known solution to this problem.\\ \\ I do
+  not regard the essence of the submission as being at all dependent
+  on inkscape or R.  Any vectorised graphics editor (VGE) that can
+  work with Bezier curves would be as good as inkscape.  Adobe
+  illustrator, or Google drawings etc etc could be substituted for
+  inkscape.  So in the submission, ``inkscape'' is short for ``any
+  VGE''.  The difficulty is that a manuscript needs a specfic VGE to
+  concretify the discussions.  The same reasoning applies to my choice
+  of R, which was again driven by my familiarity with it; matlab or
+  python would be just as good.  I believe that the resubmission gets
+  this point across.}
 
@@
 tutorial on how to use package [27] to draw optimized knot diagrams,
-but JMA is not the right place to publish such a tutorial.
+but JMA is not the right place to publish such a
+tutorial.\textcolor{blue}{\\I have expanded the discussion of the
+  aesthetics of knot diagrams, but would point out that capturing the
+  goals\\ \\}
 

### 2023-02-06 inst/jma_rebuttal_2023_01_30.tex
fixes #79
fixes #80

--- 2023-02-03 inst/jma_rebuttal_2023_01_30.tex
@@
 
-(Below, the reviewer's comments are in black, and my replies to the
-issues are in \textcolor{blue}{blue}.  I have indicated changes to the
-manuscript where appropriate.
-
-\section*{Rebuttal to "Visually pleasing knot projections" by
+\section*{Rebuttal to ``Visually pleasing knot projections" by
   Robin K. S. Hankin (submitted to JMA, manuscript 209604076); review
@@
 
-\textcolor{blue}{ Short story: }
+(Below, the reviewer's comments are in black, and my replies to the
+issues are in \textcolor{blue}{blue}.  I have indicated changes to the
+manuscript where appropriate).
+
+\section*{\textcolor{blue}{Overall response to the review}}
+
+\textcolor{blue}{The comments have resulted in a much stronger
+  document.  I have carried out all the detailed suggestions, as
+  presented below, and offer a commentary on some of the ideas.}
 
@@
 tutorial on how to use package [27] to draw optimized knot diagrams,
-but JMA is not the right place to publish such a tutorial.
+but JMA is not the right place to publish such a
+tutorial.\textcolor{blue}{ I would characterise the submission, not as
+  a online tutorial, but as a discursive yet mathematical outlining of
+  the conceptual and computational difficulties of automated
+  beautification of knot diagrams (not forgetting to discuss the
+  advantages!).  I have expanded the discussion of the aesthetics of
+  knot diagrams.  \\ \\ On the online tutorial comparison: I do not
+  regard the essence of the submission as being at all dependent on
+  inkscape or R.  Any vectorised graphics editor (VGE) that can work
+  with Bezier curves would be as good as inkscape.  Adobe illustrator,
+  or Google drawings etc etc could be substituted for inkscape.  So in
+  the submission, ``inkscape'' is short for ``any VGE''.  The
+  difficulty is that a manuscript needs a specific VGE to concretify
+  the discussions.  The same reasoning applies to my choice of R,
+  which was again driven by my familiarity with it; matlab or python
+  would be just as good.  I believe that the resubmission gets this
+  point across.}
 
@@
 on page 4 are of prime importance and could be fleshed out more.
+\textcolor{blue}{\\I have added a some discussion as suggested.}
 
@@
   specifications take priority and the resulting symmetrised diagram
-  automatically accommodates the imposed symmetry.  In practice what
-  happens is immediately visually apparent: one sees beautiful
+  has symmetric nodes but the connections are not as intended.  In
+  practice, this is immediately visually apparent: one sees beautiful
   symmetric nodes joined by Bezier curves that do not possess the
   intended topology.  The diagram is a disorganised and ugly tangled
-  mess, clearly not reflecting the intended configuration.  In
-  practice, one glance at the diagram simultaneously reveals the
-  problem and its solution (just reconnect the dots in the correct
-  order).\\ \\}
+  mess, clearly not reflecting the intended configuration.  Generally,
+  one glance at the diagram simultaneously reveals the problem and its
+  solution (just reconnect the dots in the correct order).\\ \\}
 
@@
 links should come in right at the beginning.  It is not at all clear
-why "Each component would have its own badness."
+why ``Each component would have its own badness."
 
@@
 
-Page 6, Line 42: What is "{\tt > b <}" ?  \textcolor{blue}{\\ This is
+Page 6, Line 42: What is ``{\tt > b <}" ?  \textcolor{blue}{\\ This is
   just a detail of standard R syntax.  The first ``{\tt > }" is just
@@
 
-Figure 4 caption: Much of the explanation of "7 3" versus "3 7" is
+Figure 4 caption: Much of the explanation of ``7 3" versus ``3 7" is
 already in the main text.  \textcolor{blue}{\\Fixed\\ \\}
 
-Page 8, Line 60: "Why 64 real variables"?  What are they?  How does
-this agree with "R50" on page 6?\textcolor{blue}{\\ The 64 variables are just the parameters required to describe the knot path, it agrees with the (reworded) coment about $\mathbb{R}^{50}$.\\ \\}
+Page 8, Line 60: ``Why 64 real variables"?  What are they?  How does
+this agree with ``R50" on page 6?\textcolor{blue}{\\ The 64 variables
+  are just the parameters required to describe the knot path, it
+  agrees with the (reworded) coment about $\mathbb{R}^{50}$.\\ \\}
 
@@
 pure cyclic rotational symmetry, rather than dihedral symmetry"
-  \textcolor{blue}{Done\\ \\}
+  \textcolor{blue}{\\Done\\ \\}
 

### 2023-02-06 inst/jma_knot.tex
pursuant to #15
typo

@@
 \item[{\tt total\_string\_length()}] The length is just $\int ds$, with a
-  nominal value of 5000 (which value is about right for a printed
-  copy).
+  nominal value of 5000 (which is about right for a printed copy).
 \item[{\tt midpoint\_badness()}] The idea is to suppress strands with
@@
 
+\section*{Acknowledgement}
+
+I would like to thank an anonymous referee whose insightful comments
+led to a much improved manuscript.
+
 \bibliography{knot}
